### 2021-07-08 papers/model/paper_model.tex
paper notes

%  LaTeX support: latex@mdpi.com 
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[journal,article,submit,moreauthors,pdftex]{Definitions/mdpi} 
\usepackage{natbib}
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, analytica, animals, antibiotics, antibodies, antioxidants, appliedchem, applmech, applmicrobiol, applnano, applsci, arts, asi, atmosphere, atoms, audiolres, automation, axioms, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomechanics, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, brainsci, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coatings, colloids, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, crops, cryptography, crystals, curroncol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entropy, environments, environsciproc, epidemiologia, epigenomes, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forensicsci, forests, fractalfract, fuels, futureinternet, futuretransp, futurepharmacol, futurephys, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, land, languages, laws, life, liquids, literature, livers, logistics, lubricants, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, metabolites, metals, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, minerals, mining, modelling, molbank, molecules, mps, mti, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photochem, photonics, physchem, physics, physiolsci, plants, plasma, pollutants, polymers, polysaccharides, proceedings, processes, prosthesis, proteomes, psych, psychiatryint, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, systems, taxonomy, technologies, telecom, textiles, thermo, tourismhosp, toxics, toxins, transplantology, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wevj, women, world

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
% MDPI internal commands
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2021}
\copyrightyear{2020}
%\externaleditor{Academic Editor: Firstname Lastname} % For journal Automation, please change Academic Editor to "Communicated by"
\datereceived{} 
\dateaccepted{} 
\datepublished{} 
\hreflink{https://doi.org/} % If needed use \linebreak
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{A modelling strategy to improve cacao quality and productivity }

% MDPI internal command: Title for citation in the left column
\TitleCitation{Title}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Angela RV $^{1,\dagger,\ddagger}$\orcidA{}, Adriana $^{1,\ddagger}$ and Anyela VCR$^{1,}$*}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad The John Bingham Laboratory, NIAB, 93 Lawrence Weaver Road, Cambridge CB3 0LE, UK 1; e-mail@e-mail.com\\
$^{2}$ \quad BIOS 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence:anyela.camargorodriguez@niab.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3} 
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{As.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia
%\encyclopediadef{Instead of the abstract}
%\entrylink{The Link to this entry published on the encyclopedia platform.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.

\section{Introduction}

A \cite{Woo2010}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

\begin{quote}
This is an example of a quote.
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Patents}

This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work~reported.}

\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''. Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}

\dataavailability{In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study. Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}. You might choose to exclude this statement if the study did not report any data.}

\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections. This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}

\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section. If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the~results''.}

%% Optional

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{paracol}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% To add notes in main text, please use \endnote{} and un-comment the codes below.
%\begin{adjustwidth}{-5.0cm}{0cm}
%\printendnotes[custom]
%\end{adjustwidth}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
\externalbibliography{yes}
\bibliography{bioinfo.bib}

%=====================================
% References, variant B: internal bibliography
%=====================================
%\begin{thebibliography}{999}
%% Reference 1
%\bibitem[Author1(year)]{ref-journal}
%Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
%% Reference 2
%\bibitem[Author2(year)]{ref-book1}
%Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
%% Reference 3
%\bibitem[Author3(year)]{ref-book2}
%Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
%% Reference 4
%\bibitem[Author4(year)]{ref-unpublish}
%Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} stage of publication (under review; accepted; in~press).
%% Reference 5
%\bibitem[Author5(year)]{ref-communication}
%Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
%% Reference 6
%\bibitem[Author6(year)]{ref-proceeding}
%Author 1, A.B.; Author 2, C.D.; Author 3, E.F. Title of Presentation. In Title of the Collected Work (if available), Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference; Editor 1, Editor 2, Eds. (if available); Publisher: City, Country, Year (if available); Abstract Number (optional), Pagination (optional).
%% Reference 7
%\bibitem[Author7(year)]{ref-thesis}
%Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
%% Reference 8
%\bibitem[Author8(year)]{ref-url}
%Title of Site. Available online: URL (accessed on Day Month Year).
%\end{thebibliography}

%\bibliographystyle{te}
%\bibliography{bioinfo}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

% The following MDPI journals use author-date citation: Admsci,  Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Jintelligence, JRFM, Languages, Laws, Literature, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2021-07-14 weather SIMPLE/climate model Sant/climate_model Sant.R
Add files via upload

@@ -1,0 +1,120 @@
+#' Anyela V Camargo R anyelavcamargo@gmail.com
+
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+#source('../aquacropr/R/Aqua_library.R')
+#change ARV
+
+#my_path <- "C:/Users/Angelita/Documents/aNIAB/Cacao/Analysis R"
+#setwd(my_path)
+
+
+transform_weather <- function(fname_in, fname_out){
+  
+  
+  fname_in <- 'POWER_SinglePoint_DailySantander.csv'
+  weather_raw <- read.csv(fname_in, skip = 17, header = TRUE)
+  
+  
+  DATE <- paste(paste(strsplit(as.character(weather_raw$YEAR), split ='')[[1]][3:4], 
+                      collapse = ''), sprintf("%03d", weather_raw$DOY), sep = '')
+  
+  weather_filter <- data.frame(cbind(DATE, SRAD = weather_raw[['ALLSKY_SFC_SW_DWN']],
+                                     TMAX = weather_raw[['T2M_MAX']],
+                                     TMIN = weather_raw[['T2M_MIN']],
+                                     RAIN = weather_raw[['PRECTOT']]))
+  
+ write.table(weather_filter, file = fname_out, sep=' ',
+              quote = FALSE, row.names = FALSE)
+  
+}
+
+fname_in <- 'weather_filter'
+fname_out <- 'dummy_weather.WTH'
+transform_weather(fname_in, fname_out)
+#-------------------------
+
+plot_data <- function(sdata){
+  
+  
+  
+  
+  #yy$Date <- droplevels(yy$Date)
+  # pdf('climate_plots.pdf', width=20)
+  # plot(y$Date, y$flowering, col='black', lwd=6, cex.axis=0.8, xaxt = "n", 
+  #      ylim = c(0, 110))
+  # lines(y$Date, y$rain,  col='red')
+  # lines(y$Date, y$RH,  col='blue')
+  # lines(y$Date, y$SRAD,  col='purple')
+  # lines(y$Date, y$WS2M*150,  col='brown')
+  # axis(1, at = seq(1,nrow(y), 15), labels= s$Date[seq(1, nrow(y), by=15)],
+  #      las=2, cex.axis=0.8)
+  # legend(1, 75, legend=c("Peso.cosechado.seco..kg.","rain","RH",
+  #                        "SRAD","WS2M"),
[…]

### 2021-07-15 papers/model/paper_model.tex
Fixed paper problem with references

@@
 %=================================================================
-\documentclass[journal,article,submit,moreauthors,pdftex]{Definitions/mdpi} 
+
+%\documentclass[genes]{Definitions/mdpi} 
+\documentclass[genes,article,submit,moreauthors,pdftex,10pt,a4paper]{Definitions/mdpi}
 \usepackage{natbib}
+
 % For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
@@
 % Keywords
-\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}
+\keyword{keyword 1; keyword 2; keyword 3}
 
-% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
-%\PACS{J0101}
-%\MSC{}
-%\JEL{}
+\begin{document}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Diversity
-%\LSID{\url{http://}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Applied Sciences:
-%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Data:
-%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
-If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal.
-In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
-
-%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Toxins
-%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Encyclopedia
-%\encyclopediadef{Instead of the abstract}
-%\entrylink{The Link to this entry published on the encyclopedia platform.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{section}{-1} %% Remove this when starting to work on the template.
+%\setcounter{section}{-1} %% Remove this when starting to work on the template.
 
@@
 \externalbibliography{yes}
-\bibliography{bioinfo.bib}
+\bibliography{bioinfo}
 

### 2021-07-15 weather SIMPLE/climate model Sant/climate_model Sant.R
Delete weather SIMPLE/climate model Sant directory

@@ -1,120 +1,0 @@
-#' Anyela V Camargo R anyelavcamargo@gmail.com
-
-library(ggplot2)
-library(dplyr)
-library(tidyr)
-#source('../aquacropr/R/Aqua_library.R')
-#change ARV
-
-#my_path <- "C:/Users/Angelita/Documents/aNIAB/Cacao/Analysis R"
-#setwd(my_path)
-
-
-transform_weather <- function(fname_in, fname_out){
-  
-  
-  fname_in <- 'POWER_SinglePoint_DailySantander.csv'
-  weather_raw <- read.csv(fname_in, skip = 17, header = TRUE)
-  
-  
-  DATE <- paste(paste(strsplit(as.character(weather_raw$YEAR), split ='')[[1]][3:4], 
-                      collapse = ''), sprintf("%03d", weather_raw$DOY), sep = '')
-  
-  weather_filter <- data.frame(cbind(DATE, SRAD = weather_raw[['ALLSKY_SFC_SW_DWN']],
-                                     TMAX = weather_raw[['T2M_MAX']],
-                                     TMIN = weather_raw[['T2M_MIN']],
-                                     RAIN = weather_raw[['PRECTOT']]))
-  
- write.table(weather_filter, file = fname_out, sep=' ',
-              quote = FALSE, row.names = FALSE)
-  
-}
-
-fname_in <- 'weather_filter'
-fname_out <- 'dummy_weather.WTH'
-transform_weather(fname_in, fname_out)
-#-------------------------
-
-plot_data <- function(sdata){
-  
-  
-  
-  
-  #yy$Date <- droplevels(yy$Date)
-  # pdf('climate_plots.pdf', width=20)
-  # plot(y$Date, y$flowering, col='black', lwd=6, cex.axis=0.8, xaxt = "n", 
-  #      ylim = c(0, 110))
-  # lines(y$Date, y$rain,  col='red')
-  # lines(y$Date, y$RH,  col='blue')
-  # lines(y$Date, y$SRAD,  col='purple')
-  # lines(y$Date, y$WS2M*150,  col='brown')
-  # axis(1, at = seq(1,nrow(y), 15), labels= s$Date[seq(1, nrow(y), by=15)],
-  #      las=2, cex.axis=0.8)
-  # legend(1, 75, legend=c("Peso.cosechado.seco..kg.","rain","RH",
-  #                        "SRAD","WS2M"),
[…]

### 2021-07-19 papers/model/paper_model.tex
first draft
I keep working on it. I am adding more information in the intro. This  is not the final first draft.

--- 2021-07-08 papers/model/paper_model.tex
@@
 %=================================================================
-\documentclass[journal,article,submit,moreauthors,pdftex]{Definitions/mdpi} 
+\documentclass[gene,journal,article,submit,moreauthors,pdftex]{Definitions/mdpi} 
 \usepackage{natbib}
@@
 % Keywords
-\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}
+\keyword{ICS95; CCN51; thermal time (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{section}{-1} %% Remove this when starting to work on the template.
+%\setcounter{section}{-1} %% Remove this when starting to work on the template.
 
@@
 
-A \cite{Woo2010}.
+Cocoa ( \textit{Theobroma cacao }L.
+) is a perennial tropical crop endemic to the South American rainforests \citep{zuidema2005, motamayor2002, argout2011, Rodriguez2019}.
+Only the 5\% of the world cocoa yield is desalinated for Fine-cocoa production due to the low productivity of  the traditional crop management \citep{argout2011}.
+Cacao plant member of the Malvaceae (formerly Sterculiaceae)  botanical family, is grown for its fruits, known as cacao pods.
+\citep{ Niemenak2010, suarez2021}.
+
+Cocoa has  a  long  tradition  in  Colombia  where  cocoa  is  widely  consumed  as  a  beverage.
+It is one of the crops promoted by the Colombian government in the social and agricultural development  programs aimed at favouring peace in post-conflict regions \citep{Rodriguez2019} (Abbott et al.
+2018).
+
+In this work we used the SIMPLE crop model \citep{Zao2019simple} for three reasons: 1: That it is very comprehensively described in the original paper.
+2: That the code was available in R for initial trials and 3: That it had already been successfully fitted to perennial crops in south America.
+Overall, the model simulates crop development, growth and yield, and predict the maturation day when the friut is ready to harvest.
+It includes 13 parameters to specify a crop type, with four of these for cultivar characteristics.
+Commonly available inputs that are required for the model include daily weather data, crop management, and soil water holding parameters and key dates.
+Malvaceae such as cotton \textit{ Gossypium hirstium} \citep{Nix2017cotton} wich is modeled in SIMPLE model \citep{Zao2019simple}.
+ The aim of this study is predict the best time for harvest cocoa fruits.
 
@@
 \section{Materials and Methods}
+\subsection{Phenology of Cocoa in Colombia}
+\begin{figure}[h]
+	\centering
+	\caption{\footnotesize {Phenology of cocoa in Colombia for crop modelling}} 
+	\includegraphics[scale=0.4]{images/phenology.png}\\
+	\footnotesize{Credits: Taken from Download from Dreamstime.com, 2021}
+	\label{fig:pheno}
+\end{figure}.
 
-\begin{quote}
-This is an example of a quote.
-\end{quote}
+\subsection{Crop model acquisition}
+
+A zip file of the R codes and input files for the SIMPLE model was kindly provided by Dr Chuang Zhao of the Agricultural and Biological Engineering Department, University of Florida.
+The main directory contains R codes, the “Parameter definition” Excel ® file, where the units required in each input file for the model are described, instructions and directories containing other input files.
+Examples of all input files are provided in the Zip files together with a range of data sets that can be used to check model operation after installation.
+Before running the SimpleB.R program initially line 72 was edited, with RStudio, to provide the full path for the installation.
+Also, the single year/single experiment mode was selected by setting the GridsimulationSwitch in line 58 to “off” by entering [1] at the end of line 58.
+After checking the first column in the Input/Simulation Management.csv file is all zeros apart from 1 on line alongside cocoa experiment, then “Source” at the top of the RStudio editing window was clicked.
+
+In the SIMPLE model dummy files are provided for adding new crop data and weather data, and files 2 to 6 in the list below can be edited to define new cultivars or experiments etc.
+Then modifying the simulation management file will cause the new files to be read when the program is run.
+\begin{enumerate}
+	\item Input/Simulation Management.csv
+	\item Input/Species parameter.csv
+	\item Input/Cultivar.csv
+    \item Input/Treatment.csv	
+    \item Input/Irrigation.csv
+    \item Input/Soil.csv
+    \item Observation/Obsdummy crop Exp name.csv	
+    \item Weather/dummy weather.WTH
+\end{enumerate}
+
+\subsection{Applying the SIMPLE model for the Colombia cocoa crop }
+
+In outline the procedure used was a sequential process of modifying or adding the appropriate files, changing the parameters in the simulation management file, running the program and inspecting the results.
+This cycle was then repeated until the dry matter yields for Colombia were reasonably reproduced in the results.
+The first step was add the new experiment for cocoa in the dummy crop and experiment names were replaced with our crop and treatment names in the input files and flowering and harvest dates were changed to harvest at 6 months after flowering to check that our modifications could be read.
+Initially, the new experiment name “KOKOlatl” for the cultivar ICS 95 was used in the species parameter file with other parameters and data founded in papers \textcolor{red}{debo completar eso con la bibliografia de donde saque datos de fisio}, including irrigation, soil features and radiation use efficiency.
+An appropriate treatment label was added (Figure 7), but all other options pointed to the test data from Brazil.
+
+\subsection{Input data acquisition } 
+Input Weather data were sourced from the POWER Data Access Viewer \citep{nasapower}, which allows automatic download.
+These were used in as inputs for SIMPLE model.
+Weather data had to be transformed to .WHT file.
 
@@
 \section{Results}
+\begin{figure}[h]
+	\centering
+	\caption{\footnotesize {SIMPLE model output for cocoa in Colombia }} 
+	\includegraphics[scale=0.4]{images/averageS.png}\\
+	\label{fig:weather}
+\end{figure}.
+
+\begin{figure}[h]
+	\centering
+	\caption{\footnotesize {SIMPLE model output for cocoa in Colombia }} 
+	\includegraphics[scale=0.4]{images/outmodel.jpg}\\
+	\label{fig:m1}
+\end{figure}.
 
@@
 \externalbibliography{yes}
-\bibliography{bioinfo.bib}
+\bibliography{kokobib}
 

### 2021-07-21 climate_model.R
modified climate_model.R

@@ -4,66 +4,68 @@ library(ggplot2)
 library(dplyr)
 library(tidyr)
 #source('../aquacropr/R/Aqua_library.R')
+#change ARV
 
+my_path <- "C:/Users/Angelita/Documents/R/Cacao"
+setwd(my_path)
 
 
 transform_weather <- function(fname_in, fname_out){
   
   
+  fname_in <- 'POWER_SinglePoint_DailySantander.csv'
+  weather_raw <- read.csv(fname_in, skip = 17, header = TRUE)
   
-  #fname_in <- 'weather SIMPLE/Original POWER_SinglePoint_Daily_20180101_.csv'
-  weather_raw <- read.csv(fname_in, skip = 20, header = TRUE)
+  #dat <- data.frame(sprintf('%02d', weather_raw$YEAR  %% 100))                
   
- 
-  DATE <- paste(paste(strsplit(as.character(weather_raw$YEAR), split ='')[[1]][3:4], 
-                      collapse = ''), sprintf("%03d", weather_raw$DOY), sep = '')
   
-  weather_filter <- data.frame(cbind(DATE, sRAD = weather_raw[['ALLSKY_SFC_SW_DWN']],
+  DATE <- paste(paste(sprintf('%02d', weather_raw$YEAR  %% 100), 
+                      sprintf("%03d", weather_raw$DOY), sep = ''))
+  
+  weather_filter <- data.frame(cbind(DATE, SRAD = weather_raw[['ALLSKY_SFC_SW_DWN']],
                                      TMAX = weather_raw[['T2M_MAX']],
                                      TMIN = weather_raw[['T2M_MIN']],
                                      RAIN = weather_raw[['PRECTOT']]))
   
-  write.table(weather_filter, file = fname_out, sep=' ',
+  
+  
+  write.table(weather_filter, file = fname_out, sep='  ',
               quote = FALSE, row.names = FALSE)
   
-  
-  
-  
 }
 
-
-plot_data <- function(sdata){
-  
-  
-  
-  
-  #yy$Date <- droplevels(yy$Date)
-  # pdf('climate_plots.pdf', width=20)
-  # plot(y$Date, y$flowering, col='black', lwd=6, cex.axis=0.8, xaxt = "n", 
-  #      ylim = c(0, 110))
-  # lines(y$Date, y$rain,  col='red')
[…]

### 2021-07-21 SIMPLEcocoa/Obsfunction.R
Add files via upload
SIMPLE model calibration for cocoa Colombia

@@ -1,0 +1,37 @@
+ObsInput=function(i){
+  Obsdata <-list()
+  ###treatment filter 
+  treat<-treatment[i,];print(paste(treat$Exp[1],treat$Trt[1]))
+  #########
+  obs<-read.table(paste("./Observation/Obs_",treat$Species,"_",treat$Exp,".csv",sep=""),header=TRUE,sep=",",stringsAsFactors=FALSE)
+  obs<-obs[obs$Trt. %in% (treat$Trt.),]
+  obs$Exp=treat$Exp.
+  obs$Label=treat$Label
+  obs$Trt=obs$Trt.  
+  #source("Mainfunction.R")
+  sowingDate=DOYtoDate(treat$SowingDate)
+  obs$Date=sowingDate+obs$DAP
+  Summary=obs[nrow(obs),]
+  obs$Crop=treat$Species.
+  
+  Obsdata$Biomass=obs[complete.cases(obs$Biomass),c("Date","DAP","Biomass","Trt","Exp","Crop","Label")]
+  
+  Obsdata$FSolar=obs[complete.cases(obs$FSolar),c("Date","DAP","FSolar","Trt","Exp","Crop","Label")]
+  
+  Obsdata$Yield=obs[complete.cases(obs$Yield),c("Date","DAP","Yield","Trt","Exp","Crop","Label")]
+  
+  Obsdata$Summary=Summary[,c("Date","DAP","Biomass","FSolar","Yield","Trt","Exp","Label")]
+  
+  return(Obsdata)
+}
+
+DOYtoDate=function(DATE)
+{
+  DATE=sprintf("%05d",as.numeric(DATE))
+  YEAR<-as.numeric(substring(DATE,1,2))
+  YEAR[YEAR>20]<-YEAR[YEAR>20]+1900;YEAR[YEAR<=20]<-YEAR[YEAR<=20]+2000
+  DOY=as.numeric(substring(DATE,3,5))
+  DATE=as.Date(DOY-1, origin = paste0(YEAR,"-01-01"))
+  return(DATE)
+}
+

### 2021-07-21 SIMPLEcocoa/SimpleB.R

@@ -1,0 +1,230 @@
+#########################SIMPLE MODEL v1.1##########################################################
+##############  20170308-water stress                                                     ##########
+#######  By Bing Liu, ABE, University of Florida ------  2017-04-27                       ##########
+#######  Updated by Liujun Xiao,ABE, University of Florida ------  2017-06-08             ##########
+#######  Adding PET function by Kwang Soo Kim, Seoul National University ------2017-06-08  ##########
+#######  Recoding by Liujun Xiao,ABE, University of Florida ------ 2018-03-09             ##########
+####################################################################################################
+
+####################################################################################################
+#######		Six Steps for adding a new crop       			             ###############
+#######		                         					     ###############
+#######   1. Prepare the model inputs in Treatment.csv 
+#######     Crop,Exp,Trt,weather,lat,Elev,CO2,sowingDate,
+#######	    irrigation=irrigation ID in irrigation table: must be identical to Trt in irrigation.csv table, if no irri=0
+#######	    AWC=DUL-LL (e.g. from DSSAT soil)
+#######     RCN=runoff number (DSSAT:SLRO)
+#######  	DDC=deep drainage coeff (DSSAT:SLDR)
+####### 	WUC=water uptake coeff (use default=0.096)
+#######  	RZD=root depth in mm)
+#######  	Water=switch (yes=use rainfall and irri table; no for no water stress)
+#######  	set following parameter to overwrite in R code: Tsum,	HI,	I50A,	I50B,
+#######  	NOTE-to leave some notes
+####### 
+#######  2. prepare irrigation table
+#######  	file: irrigation.csv        IRVAL=irrgation in mm per specific date
+#######
+#######  3. Prepare weather file
+#######   	if from DSSAT
+#######          *WEATHER : AUCB
+#######          @ INSI      LAT     LONG  ELEV   TAV   AMP REFHT WNDHT
+#######          AUCB   -35.00   149.00     0  17.0   7.0   -99   -99
[…]

### 2021-07-30 papers/model/paper_model.tex
introduction cocoa models in progress

@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{As.}
+\abstract{ The aim of this study is predict the best time for harvest cocoa fruits.}
 
@@
 
+Crop models represent a quantitative assumption of plant growth depending on sunlight interception efficiency values and climate data supported by a large amount of empirical and ground data.
+\citep{Reynolds2018}.
+Physiological crop models have shown to be very useful tool for provided agronomical advices and improvements of the cropping systems of annual crops mainly.
+Recently crop modelling studies are focusing on  perennial crops  production \citep{zuidema2005, Zao2019simple, Bai2020, Romero2021}.
+However, the information reported  is be less than for annual crops due to the lack of field data available , relatively high research costs and the difficulties of accumulated errors in long-term simulations \citep{zuidema2005}.
+For cacao there  approaches  to predict yield mainly  using algorithms of machine learning \citep{lamos2020} and just one simulate cocoa performance as physiological production model "SUCROS-cocoa" \citep{zuidema2005}.
+
+Regarding cocoa crops, the yield prediction has been approached from different perspectives.
+Corrales et al.
+[11] predicted the cocoa yield level in Santander.
+The authors evaluated the daily average temperature, daily relative humidity, and total
+daily precipitations rate, using ten different algorithms implemented in WEKA software.
+For them, Random Forest was the algorithm that generates the best model in order to classify cocoa yield levels.
+Other studies [12-13] evaluate the yield using linear regression models, ANOVA and mechanistic models like SUCROS, finding that climatic conditions (such as temperature, radiation and rainfall) are the most
+critical in the cocoa productivity \citep{lamos2020}.
+
 Cocoa ( \textit{Theobroma cacao }L.
-) is a perennial tropical crop endemic to the South American rainforests \citep{zuidema2005, motamayor2002, argout2011, Rodriguez2019}.
-Only the 5\% of the world cocoa yield is desalinated for Fine-cocoa production due to the low productivity of  the traditional crop management \citep{argout2011}.
+) is an important worldwide perennial tropical crop endemic to the South American rainforests \citep{zuidema2005, motamayor2002, argout2011, Rodriguez2019}.
 Cacao plant member of the Malvaceae (formerly Sterculiaceae)  botanical family, is grown for its fruits, known as cacao pods.
 \citep{ Niemenak2010, suarez2021}.
+Only the 5\% of the world cocoa yield is desalinated for Fine-cocoa production due to the low productivity of  the traditional crop management \citep{argout2011}.
+In Colombia, cocoa  is  traditionally  consumed  as  a  beverage.
+It is one of the crops promoted by the Colombian government in the social and agricultural development  programs aimed at favouring peace in post-conflict regions \citep{Rodriguez2019, Abbott2019} as cocoa is grown by approximately 52.000 \citep{Gutierrez2020} and 98\% of production being carried out by small and medium-sized producers \citep{Garcia2014, Escobar2020}.
+Colombia registered an increase of 3.750 tons in production in 2020 compared to the previous year \citep{lamos2020}.
 
-Cocoa has  a  long  tradition  in  Colombia  where  cocoa  is  widely  consumed  as  a  beverage.
-It is one of the crops promoted by the Colombian government in the social and agricultural development  programs aimed at favouring peace in post-conflict regions \citep{Rodriguez2019} (Abbott et al.
-2018).
+Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}.
+However, it is still not widely produced as the lack of adoptions of technologies by the traditional farmers.
+They empirically harvest after 5 or 6 months after flowering date, hence they ferment cocoa beans without considering the  quality of the seeds at the harvest time.
+This produce heterogeneous characteristics between each fermentation batch diminishing the quality of cocoa final product \citep{Escobar2021}.
+To identify the best moment to harvest is important to consider physiological responses affected by climate variables such as rain, solar radiation and wind.
+Crop simulation models (CSM) are mathematical calculations include in software tools that simulated the behaviour of crop systems \citep{Oteng-Darko2013}.
+For cocoa in Colombia, physiological simulation models may be valuable to identify the best moment to harvest cocoa considering variable weather conditions, soil types and cropping systems.
 
-In this work we used the SIMPLE crop model \citep{Zao2019simple} for three reasons: 1: That it is very comprehensively described in the original paper.
+In this paper, we present a physiological parametrization of SIMPLE crop model for cocoa to predict the best harvest time  and yield production.
+We used the SIMPLE crop model \citep{Zao2019simple} for three reasons: 1: That it is very comprehensively described in the original paper.
 2: That the code was available in R for initial trials and 3: That it had already been successfully fitted to perennial crops in south America.
-Overall, the model simulates crop development, growth and yield, and predict the maturation day when the friut is ready to harvest.
+Overall, the model simulates crop development, growth and yield, and predict the maturation day when the fruit is ready to harvest.
 It includes 13 parameters to specify a crop type, with four of these for cultivar characteristics.
@@
 Malvaceae such as cotton \textit{ Gossypium hirstium} \citep{Nix2017cotton} wich is modeled in SIMPLE model \citep{Zao2019simple}.
- The aim of this study is predict the best time for harvest cocoa fruits.
 
@@
 Future research directions may also be highlighted.
+
+Evaluating the level of knowledge of producers regarding cocoa crop management, the harvest was in the group of activities that presented the lowest level of knowledge on the part of the producers according to the general averages \citep{Gutierrez2020}.
 

### 2021-07-31 papers/model/paper_model.tex
Update paper_model.tex

@@
 
-Crop models represent a quantitative assumption of plant growth depending on sunlight interception efficiency values and climate data supported by a large amount of empirical and ground data.
-\citep{Reynolds2018}.
-Physiological crop models have shown to be very useful tool for provided agronomical advices and improvements of the cropping systems of annual crops mainly.
-Recently crop modelling studies are focusing on  perennial crops  production \citep{zuidema2005, Zao2019simple, Bai2020, Romero2021}.
-However, the information reported  is be less than for annual crops due to the lack of field data available , relatively high research costs and the difficulties of accumulated errors in long-term simulations \citep{zuidema2005}.
-For cacao there  approaches  to predict yield mainly  using algorithms of machine learning \citep{lamos2020} and just one simulate cocoa performance as physiological production model "SUCROS-cocoa" \citep{zuidema2005}.
-
-Regarding cocoa crops, the yield prediction has been approached from different perspectives.
-Corrales et al.
-[11] predicted the cocoa yield level in Santander.
-The authors evaluated the daily average temperature, daily relative humidity, and total
-daily precipitations rate, using ten different algorithms implemented in WEKA software.
-For them, Random Forest was the algorithm that generates the best model in order to classify cocoa yield levels.
-Other studies [12-13] evaluate the yield using linear regression models, ANOVA and mechanistic models like SUCROS, finding that climatic conditions (such as temperature, radiation and rainfall) are the most
-critical in the cocoa productivity \citep{lamos2020}.
-
 Cocoa ( \textit{Theobroma cacao }L.
 ) is an important worldwide perennial tropical crop endemic to the South American rainforests \citep{zuidema2005, motamayor2002, argout2011, Rodriguez2019}.
-Cacao plant member of the Malvaceae (formerly Sterculiaceae)  botanical family, is grown for its fruits, known as cacao pods.
+Cacao plant member of the Malvaceae (formerly Sterculiaceae)  botanical family such as  cotton \textit{ Gossypium hirstium} \citep{Nix2017cotton} wich is modeled in SIMPLE model \citep{Zao2019simple}.
+Cocoa is grown for its fruits, known as cacao pods.
 \citep{ Niemenak2010, suarez2021}.
@@
 To identify the best moment to harvest is important to consider physiological responses affected by climate variables such as rain, solar radiation and wind.
-Crop simulation models (CSM) are mathematical calculations include in software tools that simulated the behaviour of crop systems \citep{Oteng-Darko2013}.
-For cocoa in Colombia, physiological simulation models may be valuable to identify the best moment to harvest cocoa considering variable weather conditions, soil types and cropping systems.
+Thus,  for cocoa in Colombia, physiological simulation models may be valuable to identify the best moment to harvest cocoa considering variable weather conditions, soil types and cultivar specifications.
+
+Crop models represent a quantitative assumption of plant growth depending on sunlight interception efficiency values and climate data supported by a large amount of empirical and ground data.
+\citep{Reynolds2018}.
+Physiological crop models have shown to be very useful tool for provided agronomical advices and improvements of the cropping systems of annual crops mainly.
+Recently crop modelling studies are focusing on  perennial crops  production \citep{zuidema2005, Zao2019simple, Bai2020, Romero2021}.
+However, the information reported  is be less than for annual crops due to the lack of field data available , relatively high research costs and the difficulties of accumulated errors in long-term simulations \citep{zuidema2005}.
+For cacao there  approaches  to predict yield mainly  using algorithms of machine learning \citep{lamos2020} and just one mechanistic model simulates physiological cocoa performance  "SUCROS-cocoa" \citep{zuidema2005}.
+This crop model calculates light interception, photosynthesis, maintenance respiration,
+evapotranspiration, biomass production and cocoa yield.
+It can be parametrised having data on cocoa physiology and morphology \citep{zuidema2005}.
+However,  there is not specific cocoa physiology data available from small and medium-sized producers.
+Thus, we adapted the simple generic crop model (SIMPLE) that could be easily modified for any crop to simulate development, crop growth and yield using few parameters such as weather and cultivar specification \citep{Zao2019simple}.
 
@@
 Overall, the model simulates crop development, growth and yield, and predict the maturation day when the fruit is ready to harvest.
-It includes 13 parameters to specify a crop type, with four of these for cultivar characteristics.
-Commonly available inputs that are required for the model include daily weather data, crop management, and soil water holding parameters and key dates.
-Malvaceae such as cotton \textit{ Gossypium hirstium} \citep{Nix2017cotton} wich is modeled in SIMPLE model \citep{Zao2019simple}.
+It includes 13 parameters ( daily weather data, irrigation, and soil and key dates) to specify a crop type, with four of these for cultivar characteristics easily available from farmer.
+Thus, this could be used as  tool for small farmers with the aim to improve the quality of cocoa to become more competitive in Fine-cocoa market.
 

### 2021-07-31 papers/model/paper_model.tex
Update paper_model.tex

@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{ The aim of this study is predict the best time for harvest cocoa fruits.}
+\abstract{ Crop modelling can support agronomical decisions of crop production under a range of scenarios improving competitiveness.
+Cocoa production systems in Latin America has a high importance over social and economic development, facing  the fight against hunger and poverty.
+Cocoa fruits development can be considered as the result of a number of physiological and morphological processes that can be described by mathematical relationships even under uncontrolled environments.
+We parametrize the SIMPLE crop model \citep{Zao2019simple} to predict the best time for harvest cocoa fruits in Colombia.
+The aim of this study is develop a practical tool for farmers using flowering date and weather variables (Solar radiation, rain and temperature) easily available for crop modelling to understand and predict the phenology of cacao trees which is key to produce high quality cacao beans.}
 

### 2021-07-31 papers/model/paper_model.tex
Update paper_model.tex

@@
 Cocoa fruits development can be considered as the result of a number of physiological and morphological processes that can be described by mathematical relationships even under uncontrolled environments.
-We parametrize the SIMPLE crop model \citep{Zao2019simple} to predict the best time for harvest cocoa fruits in Colombia.
+We parametrized the SIMPLE crop model \citep{Zao2019simple} to predict the best time for harvest cocoa fruits in Colombia.
 The aim of this study is develop a practical tool for farmers using flowering date and weather variables (Solar radiation, rain and temperature) easily available for crop modelling to understand and predict the phenology of cacao trees which is key to produce high quality cacao beans.}
@@
 % Keywords
-\keyword{ICS95; CCN51; thermal time (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}
+\keyword{ICS95; CCN51; thermal time, flowering date}
 

### 2021-08-03 papers/model/paper_model.tex
Update paper_model.tex

@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{ Crop modelling can support agronomical decisions of crop production under a range of scenarios improving competitiveness.
+\abstract{Crop modelling can support agronomical decisions of crop production under a range of scenarios improving competitiveness.
 Cocoa production systems in Latin America has a high importance over social and economic development, facing  the fight against hunger and poverty.
@@
 We parametrized the SIMPLE crop model \citep{Zao2019simple} to predict the best time for harvest cocoa fruits in Colombia.
-The aim of this study is develop a practical tool for farmers using flowering date and weather variables (Solar radiation, rain and temperature) easily available for crop modelling to understand and predict the phenology of cacao trees which is key to produce high quality cacao beans.}
+The aim of this study is develop a practical tool for farmers using flowering date and weather variables (Solar radiation, rain and temperature) easily available for crop modelling to understand and predict the phenology of cacao trees which is key to produce high quality cacao beans.
+}
 
 % Keywords
-\keyword{ICS95; CCN51; thermal time, flowering date}
+\keyword{ICS95; CCN51; thermal time, flowering date }
 
@@
 \section{Materials and Methods}
-\subsection{Phenology of Cocoa in Colombia}
+\subsection{Floral Phenology of Cocoa }
+In the SIMPLE model we simulate the floral phenology from the date of inflorescence emergence (BBCH scale 5) to predict the date of ripening of fruit and seed (BBCH 8) \citep{Niemenak2010} (See figure \ref{fig:pheno}).
+For modelling purposes the crop cycle of cocoa as perennial plant does not start at the plantation date such as annual crops where usually the planting date as start point of the crop cycle to model.
+The growth period of the fruit can varied from  110 to 150  daa (days after anthesis) \citep{lopez2018} when cacao fruits reaches the physiological maturity, but if can be harvested at 170 days daa \citep{Niemenak2010}.
+
 \begin{figure}[h]
 	\centering
-	\caption{\footnotesize {Phenology of cocoa in Colombia for crop modelling}} 
-	\includegraphics[scale=0.4]{images/phenology.png}\\
-	\footnotesize{Credits: Taken from Download from Dreamstime.com, 2021}
+	\caption{\footnotesize {Phenology of cocoa in Colombia for crop modelling\\}} 
+	\includegraphics[scale=0.5]{images/phenology.png}\\
+	\footnotesize{Credits: Taken from phys.org \citep{toledo2021} and \cite{lopez2018}}
 	\label{fig:pheno}
-\end{figure}.
+\end{figure}
+\newpage
 
-\subsection{Crop model acquisition}
+\subsection{Input data acquisition for SIMPLE crop model calibration}
+Input variables required to run SIMPLE model for cocoa include the flowering date and daily weather of solar radiation (SRAD), maximum and minimum temperature (TMAX, TMIN) and rain.
+Weather data as csv file was  downloaded from the POWER Data Access Viewer \citep{nasapower} from January 01 2018 to December 31 2020.
+The csv file had to be transformed to .WHT file using R 1.4 version \citep{Rstudio2020}.
+
+\subsection{SIMPLE crop model calibration}
+
+The procedure for the model calibration used was a sequential process of modifying or adding the appropriate files, changing the parameters in the simulation management file, running the program and inspecting the results.
+This cycle was then repeated until the  simulated yields for cocoa were reasonably close to the observed yield.
 
 A zip file of the R codes and input files for the SIMPLE model was kindly provided by Dr Chuang Zhao of the Agricultural and Biological Engineering Department, University of Florida.
-The main directory contains R codes, the “Parameter definition” Excel ® file, where the units required in each input file for the model are described, instructions and directories containing other input files.
+The main directory contains R codes, the \emph{“Parameter definition”} Excel ® file, where the units required in each input file for the model are described, instructions and directories containing other input files.
 Examples of all input files are provided in the Zip files together with a range of data sets that can be used to check model operation after installation.
 Before running the SimpleB.R program initially line 72 was edited, with RStudio, to provide the full path for the installation.
-Also, the single year/single experiment mode was selected by setting the GridsimulationSwitch in line 58 to “off” by entering [1] at the end of line 58.
-After checking the first column in the Input/Simulation Management.csv file is all zeros apart from 1 on line alongside cocoa experiment, then “Source” at the top of the RStudio editing window was clicked.
+Also, the single year/single experiment mode was selected by setting the Grid-simulation switch in line 58 to “off” by entering [1] at the end of line 58.
+After checking the first column in the Input/Simulation Management.csv file is all zeros apart from 1 on line alongside cocoa experiment, then the model was run.
 
-In the SIMPLE model dummy files are provided for adding new crop data and weather data, and files 2 to 6 in the list below can be edited to define new cultivars or experiments etc.
+In the SIMPLE model dummy files are provided for adding new cocoa crop data and weather data, and files 2 to 6 in the list below can be edited to define new cultivars or experiments.
 Then modifying the simulation management file will cause the new files to be read when the program is run.
@@
 
-\subsection{Applying the SIMPLE model for the Colombia cocoa crop }
+The first step was add the new experiment for cocoa in the dummy crop and experiment names were replaced with our crop and treatment names in the input files and flowering date was changed.
+Thermal time values required for Input/Cultivar.csv  had to be defined for ICS 95 and CCN51 cultivars using field data from Fedecacao plot reports in Santander and Caldas, Colombia.
+Thus, the date for harvest is predicted when the model calculations reach the thermal time  for  each specific cultivar ICS 95 and CCN51.
+The new experiment name “KOKOlatl” for was used in the species parameter file with the respective values of parameters for cocoa including irrigation, soil features and radiation use efficiency \citep{zuidema2005} \textcolor{red}{add table parameters values will change through the calibration}.
 
-In outline the procedure used was a sequential process of modifying or adding the appropriate files, changing the parameters in the simulation management file, running the program and inspecting the results.
-This cycle was then repeated until the dry matter yields for Colombia were reasonably reproduced in the results.
-The first step was add the new experiment for cocoa in the dummy crop and experiment names were replaced with our crop and treatment names in the input files and flowering and harvest dates were changed to harvest at 6 months after flowering to check that our modifications could be read.
-Initially, the new experiment name “KOKOlatl” for the cultivar ICS 95 was used in the species parameter file with other parameters and data founded in papers \textcolor{red}{debo completar eso con la bibliografia de donde saque datos de fisio}, including irrigation, soil features and radiation use efficiency.
-An appropriate treatment label was added (Figure 7), but all other options pointed to the test data from Brazil.
+\begin{table}[h!]	
+\caption {\footnotesize {Cocoa crop parameter values used in Species-parameterfile.}} \label{tab:Speparams} 
+	\centering
+	\begin{small}
+		\begin{tabular}{{c c c c c c c c c }}
+			\hline
+			{\bf Tbase }& {\bf Topt } & {\bf RUE} & {\bf I50maxH}& {\bf I50maxW}& {\bf MaxT}& {\bf ExtremeT} & {\bf CO2-RUE}& {\bf S-water}\\
+			\hline
+			105& 26 & 0.8 & 100 & 5& 35&45&0.09&1.5 \\
+			\hline
+		\end{tabular}  
+	\end{small}
+\end{table}
 
-\subsection{Input data acquisition } 
-Input Weather data were sourced from the POWER Data Access Viewer \citep{nasapower}, which allows automatic download.
-These were used in as inputs for SIMPLE model.
-Weather data had to be transformed to .WHT file.
+\begin{table}[h!]	
+	\caption {\footnotesize {Cocoa crop parameter values used in Cultivar file.}} \label{tab:Culparam} 
+	\centering
+	\begin{small}
+		\begin{tabular}{{c c c c c}}
+			\hline
+			{\bf Cultivar }& {\bf Tsum } & {\bf HI} & {\bf I50A}& {\bf I50B}\\
+			\hline
+			ICS 95& 3300 & 0.4 & 680 & 200 \\
+			CCN51  & 3600 & 0.4 & 680 & 200 \\
+			\hline
+		\end{tabular}  
+	\end{small}
+\end{table}
+
+\begin{table}[h!]	
+	\caption {\footnotesize {Cocoa crop parameter values used in Treatment and Observation file.}} \label{tab:Treaparam} 
+	\centering
+	\begin{small}
+		\begin{tabular}{{c c }}
+			\hline
+			{\bf Variable name }&{\bf Value}\\
+			\hline
+			Exp&KOKOlatl\\
+			Label& Colombia\\
+			Weather & KOKOS (.WTH file name)\\
+			CO2 & 400\\
+			SowingDate &flowering date\\
+			SoilName & Loamy sand4\\
+			Cultivar& ICS 95 or CCN51\\
+			MaxIntercept& 0.95\\
+			InitialBio & 1\\
+			InitialTT & 0\\
+			InitialFsolar & 0.01 \\
+			LAI & 2 \\
+			\hline
+		\end{tabular}  
+	\end{small}
+\end{table}
+
+\subsection{Evaluation of model performance}
+
+The SIMPLE model performance was evaluated by comparing simulated values cocoa yield with those reported by Fedecacao from cocoa plantations, using statistical indices of relative-RMSE (RRMSE) \citep{Zao2019simple} and coefficient of determination (R$^{2}$) \citep{Zao2019simple, Bai2020, Camargo2019aquacropr}.
+The harvest date predicted was supported by \textcolor{red}{Adriana results}
 
@@
 	\label{fig:weather}
+\end{figure}
+
+Studying the weather data effects over the flowering date and their final yield.
+We found that the number of successful flowers pollinated to produce final yield can be affected mostly by the rain, TMAX and wind (Figure \ref{fig:correlation}).
+The later can affect the availability of tiny flies pollinators from Diptera order and from the families of of the biting midges \textit{Ceratopogonidae},  genus  \textit{Forcipomyia} \citep{Saunders1959, kaufmann1975, sotomayor2020} to reach the cocoa flowers.
+However, wind is not included in the SIMPLE model as an input as cocoa crop has not been simulated with this model before.
+Flower opening is very well synchronised between the cohorts of mature flowers opening each night.
+The flowers open at almost exactly the same time and rate, irrespective of their position on the trunk.
+Unfertilised flowers abscise from the trunk approximately 1 day after flower opening  \citep{Niemenak2010}.
+\begin{figure}[h]
+	\centering
+	\caption{\footnotesize {Phenology of cocoa in Colombia for crop modelling\\}} 
+	\includegraphics[scale=0.7]{images/correlation.png}\\
+	\label{fig:correlation}
 \end{figure}.
@@
 Future research directions may also be highlighted.
+\textcolor{red}{this section on going to organize}
 

### 2021-08-04 SIMPLEcocoa/Mainfunction.R
fixed bug in DB references

@@ -359,6 +359,7 @@ ARIDFunction<-function(weather,soil,Lat,Elev)
   dat<-data.frame(SRAD=dat$SRAD,TMAX=dat$TMAX,TMIN=dat$TMIN,DEWPOINT=dat$TMIN,RAIN=dat$RAIN,
                   WINDSPEED=rep(1,nrow(dat)),DOY=as.numeric(substring(dat$DATE,3,5)),YEAR=as.numeric(format(dat$IDATE,"%Y")));
   
+ 
   rain <- dat[,5]                                                   # The 5th column of the data is PRECIPITATION
   
   ######## 3. set up Constants for ARID
@@ -402,7 +403,7 @@ ARIDFunction<-function(weather,soil,Lat,Elev)
     SD <- 0.409*sin(2*pi*dat[i,7]/days-1.39)
     SSA <- acos(-tan(lat)*tan(SD))
     extra <- 24*60*0.082/pi*IRDES*(SSA*sin(lat)*sin(SD)+cos(lat)*cos(SD)*sin(SSA))
-    CSR <- (0.75+2*10^-5*elevation)*extra
+    CSR <- (0.75+2*10^-5*elevation)*extra  # (Eq. 37 FAO )
     RRAD <- dat[i,1]/CSR
     ea <- 0.6108*exp(17.27*dat[i,4]/(dat[i,4]+237.3))
     LWR <- 4.903*10^-9*((dat[i,2]+273.16)^4+(dat[i,3]+273.16)^4)/2*(0.34-0.14*sqrt(ea))*(1.35*RRAD-0.35)

### 2021-08-04 papers/model/paper_model.tex
Update paper_model.tex
some adds to paper

@@
 Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}.
-However, it is still not widely produced as the lack of adoptions of technologies by the traditional farmers.
+It is still not widely produced as the lack of adoptions of technologies by the traditional farmers.
 They empirically harvest after 5 or 6 months after flowering date, hence they ferment cocoa beans without considering the  quality of the seeds at the harvest time.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Patents}
-
-This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace{6pt}
 
@@
 
-\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
-Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section.
-If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the~results''.}
+\conflictsofinterest{The authors declare no conflict of interest.}
 

### 2021-09-01 papers/model/paper_model.tex
some adds to paper
Update paper_model.tex

@@
-%  LaTeX support: latex@mdpi.com 
+%  %  LaTeX support: latex@mdpi.com 
 %  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.
@@
 \usepackage{natbib}
+\usepackage[colorinlistoftodos]{todonotes}
+\newcounter{mycomment}
+\newcommand{\mycomment}[2][]{%
+% initials of the author (optional) + note in the margin
+\refstepcounter{mycomment}%
+{%
+\setstretch{0.7}% spacing
+\todo[color={red!100!green!33},size=\small]{%
+\textbf{Comment [\uppercase{#1}\themycomment]:}~#2}%
+}}
 % For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
@@
 
-Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}.
-It is still not widely produced as the lack of adoptions of technologies by the traditional farmers.
+Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}, it is still not widely produced as the lack of adoptions of technologies by the traditional farmers.
 They empirically harvest after 5 or 6 months after flowering date, hence they ferment cocoa beans without considering the  quality of the seeds at the harvest time.
@@
 \subsection{Floral Phenology of Cocoa }
-In the SIMPLE model we simulate the floral phenology from the date of inflorescence emergence (BBCH scale 5) to predict the date of ripening of fruit and seed (BBCH 8) \citep{Niemenak2010} (See figure \ref{fig:pheno}).
-For modelling purposes the crop cycle of cocoa as perennial plant does not start at the plantation date such as annual crops where usually the planting date as start point of the crop cycle to model.
-The growth period of the fruit can varied from  110 to 150  daa (days after anthesis) \citep{lopez2018} when cacao fruits reaches the physiological maturity, but if can be harvested at 170 days daa \citep{Niemenak2010}.
+Normally the phenological stages in cocoa are  divided in two main phases: vegetative and reproductive.
+In the SIMPLE model we simulate reproductive phase described by the floral phenology from the date of inflorescence emergence (BBCH scale 5) to predict the date of ripening of fruit and seed (BBCH 8) \citep{Niemenak2010} (See figure \ref{fig:pheno}).
+In the Andean region the reproductive phase is cyclically fulfilled during two annual cycles passing by the following phases: inflorescence emergence, flowering, pollination, fruit development and harvest.
+Therefore, for modelling parametrization the crop cycle of cocoa as perennial plant does not start at the plantation date such as annual crops systems.
+Instead  the start point of the cocoa crop cycle to model is the inflorescence emergence date.
+Consequently , the growth period of the fruit can varied from  110 to 150  daa (days after anthesis) \citep{lopez2018} when cacao fruits reaches the physiological maturity, but it can be harvested at 170 days daa \citep{Niemenak2010} for quality purposes.
+
+Cocoa is a cauliflorous plant, which means that flowers grow on the trunk and branches.
+In Colombia cocoa trees usually produce flowers throughout the year, however the biggest flowering occurs in September and January to harvest in March and July (Fig.\ref{fig:yield}).
+Cocoa trees produces with up to 10000 flowers per tree each year.
+which the 50 \%  do not develop into ripe fruits according to Fedecacao reports.
+The flower takes 30 days passing by 12 micro-stages from meristem development (stages 1–6) to the fully developed flower(Stages 7–12) \citep{swanson2005} when it is ready to be pollinated.
+2005).
+The opening of flowers or anthesis  occurs over a 12-hour period during the night and it is synchronised between the groups of mature flowers \citep{Niemenak2010}.
+However, the live of a flower can last approximately 1 day after the opening falling form the trunk if it is unfertilised \citep{cheesman1927, Niemenak2010}.
+
+Subsequentially, after anthesis the fruit growths by approximately 150 days until the maturation, mucilage.
+Therefore, the complete maturation process of the fruit, from the pollination to fully mature fruit, takes 160–210 days \cite{berry1994}.
+The accumulation of lipids, storage proteins and anthocyanin starts about 85 days after pollination when fruits have an active metabolism and seeds   moisture content decreases up to 30\% \cite{Lehrian1980, Niemenak2010}.
+During this phase the quality of cocoa seeds is defined.\\
+
+\begin{figure}[h!]
+	\centering
+	\caption{\footnotesize {Phenology of cocoa in Colombia for crop modelling\\}} 
+	\includegraphics[scale=0.3]{images/phenology.png}\\
+	\footnotesize{Credits: Taken from from Dreamstime.com, phys.org \citep{toledo2021} and \cite{lopez2018}}
+	\label{fig:pheno}
+\end{figure}
 
@@
 	\centering
-	\caption{\footnotesize {Phenology of cocoa in Colombia for crop modelling\\}} 
-	\includegraphics[scale=0.5]{images/phenology.png}\\
-	\footnotesize{Credits: Taken from phys.org \citep{toledo2021} and \cite{lopez2018}}
-	\label{fig:pheno}
+	\caption{\footnotesize {Thermal time characterization of cocoa for four locations\\}}
+	\includegraphics[scale=0.35]{images/yieldmonth.png}
+	\label{fig:yield}
+\end{figure}
+\newpage
+
+\subsection{Test Site}
+Cocoa field are located in Saravena (Arauca), Rionegro (Santander), Cali (Valle del Cauca), Apartado (Antioquia) and Manizalez (Caldas).
+We considered 112 parcels data from Fedecacao reports.Each parcel data contained age of crop, density of planting, yield, number of fruits harvested, flowering date, harvest date.
+
+\subsection{Climatological Conditions}
+The temperature ranges from 16 to 28 $^\circ$C and it is relatively constant during the year.
+Saravena presented the maximum variability with hotter months during the first half of the year.
+Apartado was found as the hottest with 26 $^\circ$C and Caldas as the coolest site with 18 $^\circ$C compared with other regions tested.
+Cali and Santander have similar temperature but contrasting PAR conditions.
+Santander and Caldas have the biggest variability and the maximum of solar radiation values over 20 MJ m$^{2}$day$^{-1}$.
+In contrast , Cali and Saravena presented the lowest values of PAR below 5 MJ m$^{2}$day$^{-1}$.
+(Fig.\ref{fig:temp})
+
+Precipitation in Colombia is presented in two seasons per year from February to April and from October to November (Fig.\ref{fig:rain}). While the relative humidity
+
+\begin{figure}[h!]
+	\centering
+	\caption{\footnotesize {Daily average temperature and available photosynthetic solar radiation (PAR) conditions. \\}}
+	\includegraphics[scale=0.25]{images/tempe.png}
+	\includegraphics[scale=0.25]{images/SRAD.png}
+	\label{fig:temp}
+\end{figure}
+
+\begin{figure}[h!]
+	\centering
+	\caption{\footnotesize {Precipitation and relative humidity per month in Colombia \\}}
+	\includegraphics[scale=0.4]{images/rainmonth.png}
+	\label{fig:rain}
 \end{figure}
@@
 Input variables required to run SIMPLE model for cocoa include the flowering date and daily weather of solar radiation (SRAD), maximum and minimum temperature (TMAX, TMIN) and rain.
-Weather data as csv file was  downloaded from the POWER Data Access Viewer \citep{nasapower} from January 01 2018 to December 31 2020.
+Weather data as csv file was  downloaded from the POWER Data Access Viewer \citep{nasapower} from January 01 2018 to December 31 2020 for four locations in Colombia (Cali, Rionegro -Santander, Apartado - Antioquia, Saravena Arauca).
 The csv file had to be transformed to .WHT file using R 1.4 version \citep{Rstudio2020}.
@@
 
-The procedure for the model calibration used was a sequential process of modifying or adding the appropriate files, changing the parameters in the simulation management file, running the program and inspecting the results.
+The procedure for the SIMPLE model \citep{Zao2019simple}calibration used was a sequential process of modifying or adding the appropriate files, changing the parameters in the simulation management file, running the program and inspecting the results.
 This cycle was then repeated until the  simulated yields for cocoa were reasonably close to the observed yield.
-
-A zip file of the R codes and input files for the SIMPLE model was kindly provided by Dr Chuang Zhao of the Agricultural and Biological Engineering Department, University of Florida.
-The main directory contains R codes, the \emph{“Parameter definition”} Excel ® file, where the units required in each input file for the model are described, instructions and directories containing other input files.
-Examples of all input files are provided in the Zip files together with a range of data sets that can be used to check model operation after installation.
-Before running the SimpleB.R program initially line 72 was edited, with RStudio, to provide the full path for the installation.
-Also, the single year/single experiment mode was selected by setting the Grid-simulation switch in line 58 to “off” by entering [1] at the end of line 58.
-After checking the first column in the Input/Simulation Management.csv file is all zeros apart from 1 on line alongside cocoa experiment, then the model was run.
 
@@
 			\hline
-			105& 26 & 0.8 & 100 & 5& 35&45&0.09&1.5 \\
+			10& 26 & 0.8 & 100 & 5& 35&45&0.09&1.5 \\
 			\hline
@@
 \begin{table}[h!]	
-	\caption {\footnotesize {Cocoa crop parameter values used in Cultivar file.}} \label{tab:Culparam} 
+\caption {\footnotesize {Cocoa crop parameter values used in Cultivar file.}}
+\label{tab:Culparam} 
 	\centering
 	\begin{small}
-		\begin{tabular}{{c c c c c}}
+		\begin{tabular}{c c c c c}
 			\hline
-			{\bf Cultivar }& {\bf Tsum } & {\bf HI} & {\bf I50A}& {\bf I50B}\\
+			{\bf Cultivar }&{\bf Tsum }&{\bf HI}&{\bf I50A}&{\bf I50B}\\
 			\hline
-			ICS 95& 3300 & 0.4 & 680 & 200 \\
+			ICS 95 & 3300 & 0.4 & 680 & 200 \\
 			CCN51  & 3600 & 0.4 & 680 & 200 \\
@@
 The harvest date predicted was supported by \textcolor{red}{Adriana results}
+Yield reports from 20190 to 2020  were shared by Federación Nacional de Cacaoteros - Fedecacao.
+
+\subsection{Outputs}
 
@@
 	\centering
-	\caption{\footnotesize {SIMPLE model output for cocoa in Colombia }} 
+	\caption{\footnotesize {Weather variables during flowering times}} 
 	\includegraphics[scale=0.4]{images/averageS.png}\\
@@
 Unfertilised flowers abscise from the trunk approximately 1 day after flower opening  \citep{Niemenak2010}.
-\begin{figure}[h]
+\begin{figure}[h!]
 	\centering
-	\caption{\footnotesize {Phenology of cocoa in Colombia for crop modelling\\}} 
+	\caption{\footnotesize {Pearson correlation among weather variables and flowering dates \\}} 
 	\includegraphics[scale=0.7]{images/correlation.png}\\
@@
 	\centering
+	\caption{\footnotesize {Thermal time characterization of cocoa for four locations\\}} 
+	\includegraphics[scale=0.5]{images/ttbbox.png}
+	\label{fig:ttbox}
+\end{figure}.
+\newpage
+
+\begin{figure}[h!]
+	\centering
 	\caption{\footnotesize {SIMPLE model output for cocoa in Colombia }} 
-	\includegraphics[scale=0.4]{images/outmodel.jpg}\\
+	\includegraphics[scale=0.3]{images/outmodel.png}\\
 	\label{fig:m1}

### 2021-09-02 papers/model/paper_model.tex
equations

@@
 
-The SIMPLE model performance was evaluated by comparing simulated values cocoa yield with those reported by Fedecacao from cocoa plantations, using statistical indices of relative-RMSE (RRMSE) \citep{Zao2019simple} and coefficient of determination (R$^{2}$) \citep{Zao2019simple, Bai2020, Camargo2019aquacropr}.
+The SIMPLE model performance was evaluated by comparing simulated values cocoa yield with those reported by Fedecacao from cocoa plantations, using statistical indices of relative-RMSE (RRMSE) (Equ.
+\ref{equ:RMSE}) \citep{Zao2019simple} and coefficient of determination (R$^{2}$) (Equ.
+\ref{equ:R2}) \citep{Zao2019simple, Bai2020, Camargo2019aquacropr}.
 The harvest date predicted was supported by \textcolor{red}{Adriana results}
@@
 
-\subsection{Outputs}
+\begin{equation}
+RMSE= \sqrt{\frac{1}{n}  \sum_{i=1}^{n} (Y_{i}-X_{i})^{2} } 
+\label{equ:RMSE}
+\end{equation}
+
+\begin{equation}
+R^{2}= 1- \frac{\sum_{i=1}^{n} (Y_{i}-X_{i})^{2}}{{\sum_{i=1}^{n}(Y_{i}-0)}^{2}}
+\label{equ:R2}
+\end{equation}
 

### 2021-10-07 papers/model/paper_model.tex
results in progress

@@
-%  %  LaTeX support: latex@mdpi.com 
+%  LaTeX support: latex@mdpi.com 
 %  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.
@@
 %----------
-% Choose between the following MDPI journals:
-% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, analytica, animals, antibiotics, antibodies, antioxidants, appliedchem, applmech, applmicrobiol, applnano, applsci, arts, asi, atmosphere, atoms, audiolres, automation, axioms, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomechanics, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, brainsci, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coatings, colloids, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, crops, cryptography, crystals, curroncol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entropy, environments, environsciproc, epidemiologia, epigenomes, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forensicsci, forests, fractalfract, fuels, futureinternet, futuretransp, futurepharmacol, futurephys, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, land, languages, laws, life, liquids, literature, livers, logistics, lubricants, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, metabolites, metals, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, minerals, mining, modelling, molbank, molecules, mps, mti, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photochem, photonics, physchem, physics, physiolsci, plants, plasma, pollutants, polymers, polysaccharides, proceedings, processes, prosthesis, proteomes, psych, psychiatryint, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, systems, taxonomy, technologies, telecom, textiles, thermo, tourismhosp, toxics, toxins, transplantology, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wevj, women, world
 
@@
 % Authors, for the paper (add full first names)
-\Author{Angela RV $^{1,\dagger,\ddagger}$\orcidA{}, Adriana $^{1,\ddagger}$ and Anyela VCR$^{1,}$*}
+\Author{Angela Romero V $^{1,\dagger,\ddagger}$\orcidA{}, Adriana Gallego $^{1,\ddagger}$ and Anyela V. Camargo R.$^{1,}$*}
 
@@
 % MDPI internal command: Authors, for citation in the left column
-\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
+\AuthorCitation{Romero, A.; Gallego, A.; Camargo, A.}
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
@@
 The aim of this study is develop a practical tool for farmers using flowering date and weather variables (Solar radiation, rain and temperature) easily available for crop modelling to understand and predict the phenology of cacao trees which is key to produce high quality cacao beans.
+\textcolor{blue}{need to complete}.
 }
@@
 
-% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
-%\PACS{J0101}
-%\MSC{}
-%\JEL{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Diversity
-%\LSID{\url{http://}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Applied Sciences:
-%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Data:
-%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
-If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal.
-In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
-
-%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Toxins
-%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Encyclopedia
-%\encyclopediadef{Instead of the abstract}
-%\entrylink{The Link to this entry published on the encyclopedia platform.}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 which the 50 \%  do not develop into ripe fruits according to Fedecacao reports.
-The flower takes 30 days passing by 12 micro-stages from meristem development (stages 1–6) to the fully developed flower(Stages 7–12) \citep{swanson2005} when it is ready to be pollinated.
+The flower takes 30 days passing by 12 micro-stages from meristem development (stages 1–6) to the fully developed flower (Stages 7–12) \citep{swanson2005} when it is ready to be pollinated.
 2005).
@@
 	\centering
-	\caption{\footnotesize {Phenology of cocoa in Colombia for crop modelling\\}} 
+	\caption{\footnotesize {Phenology of cocoa in Colombia for crop modelling.\\}} 
 	\includegraphics[scale=0.3]{images/phenology.png}\\
@@
 	\centering
-	\caption{\footnotesize {Thermal time characterization of cocoa for four locations\\}}
-	\includegraphics[scale=0.35]{images/yieldmonth.png}
+	\caption{\footnotesize {Average of cocoa production characterization of cocoa for four locations\\}}
+	\includegraphics[scale=0.35]{images/yieldmonth.png}\\
+	{\footnotesize Flowering occurs in September and January to harvest six months latter in March and July respectively.}
 	\label{fig:yield}
@@
 
+\subsection{Input data acquisition for SIMPLE crop model calibration}
+Input variables required to run SIMPLE model for cocoa include the flowering date and daily weather of solar radiation (SRAD), maximum and minimum temperature (TMAX, TMIN) and rain.
+Weather data as csv file was  downloaded from the POWER Data Access Viewer \citep{nasapower} from January 01 2018 to December 31 2020 for four locations in Colombia (Cali, Rionegro -Santander, Apartado - Antioquia, Saravena Arauca).
+The csv file had to be transformed to .WHT file using R 1.4 version \citep{Rstudio2020}.
+
 \subsection{Climatological Conditions}
@@
 In contrast , Cali and Saravena presented the lowest values of PAR below 5 MJ m$^{2}$day$^{-1}$.
-(Fig.\ref{fig:temp})
-
-Precipitation in Colombia is presented in two seasons per year from February to April and from October to November (Fig.\ref{fig:rain}). While the relative humidity
+(Fig.\ref{fig:temp}).
+Precipitation in Colombia is presented in two seasons per year from February to April and from October to November, hile the relative humidity remain constant (Fig.\ref{fig:rain}).
 
@@
 
-\subsection{Input data acquisition for SIMPLE crop model calibration}
-Input variables required to run SIMPLE model for cocoa include the flowering date and daily weather of solar radiation (SRAD), maximum and minimum temperature (TMAX, TMIN) and rain.
-Weather data as csv file was  downloaded from the POWER Data Access Viewer \citep{nasapower} from January 01 2018 to December 31 2020 for four locations in Colombia (Cali, Rionegro -Santander, Apartado - Antioquia, Saravena Arauca).
-The csv file had to be transformed to .WHT file using R 1.4 version \citep{Rstudio2020}.
+\subsection{Thermal time for pod harvest date identification}
+Thermal time (degree days) was calculated as cumulative sum of temperature from the flowering date to the pod harvest.
+It is required to model crop growth  considering the cocoa base temperature (T$_{b}$), at which the plant development stops \citep{Ritchie1991}.
+Thus a cocoa T$_{b}$ of 10$^\circ$C was considered according to \cite{lahive2019}.
+This characterization was conducted for the five regions counted 180 days (6 months) and 210 days (7 months just for Cladas) after flowering, as farmers used to harvest by calendar days.
+Characterizing the thermal time the models predicts the maturation day to harvest the cocoa can vary depending on temperature variations.
 
@@
 \begin{table}[h!]	
-\caption {\footnotesize {Cocoa crop parameter values used in Species-parameterfile.}} \label{tab:Speparams} 
+\caption {\footnotesize {Cocoa crop parameter values used in Species-parameter file.}} \label{tab:Speparams} 
 	\centering
@@
 			\hline
-			10& 26 & 0.8 & 100 & 5& 35&45&0.09&1.5 \\
+			10& 26 & 0.6 & 100 & 5& 35&40&0.09&0 \\
 			\hline
 		\end{tabular}  
+	{\footnotesize S-water is associated drought stress evaluations ranging from 0 (no water shortage) to 1 (extreme water shortage) \cite{Zao2019simple} }
 	\end{small}
@@
 			\hline
-			ICS 95 & 3300 & 0.4 & 680 & 200 \\
-			CCN51  & 3600 & 0.4 & 680 & 200 \\
+			ICS 95 & 1912 & 0.3 & 680 & 200 \\
+			CCN51  & 1912 & 0.3 & 680 & 200 \\
 			\hline
@@
 			InitialFsolar & 0.01 \\
-			LAI & 2 \\
+			LAI & 1.8 \\
 			\hline
@@
 The harvest date predicted was supported by \textcolor{red}{Adriana results}
-Yield reports from 20190 to 2020  were shared by Federación Nacional de Cacaoteros - Fedecacao.
+
+Yield reports from 20190 to 2020  were shared by Federación Nacional de Cacaoteros Fedecacao.
 
@@
 \section{Results}
-\begin{figure}[h]
+
+\subsection{Thermal time per region}
+The thermal time characterization showed differences per region as was expected.
+Figure \ref{fig:ttbox} shows degree days ($^\circ$C) of cocoa growth cycles.
+Caldas had the lowest values and Apartado had the the highest results.
+this disttribucion is conincident with the temperature (Fig.\ref{fig:temp}).
+Meanwhile, Cali and Santander presented similar thermal time arround 2016 $^\circ$C.
+
+\begin{figure}[h!]
 	\centering
-	\caption{\footnotesize {Weather variables during flowering times}} 
-	\includegraphics[scale=0.4]{images/averageS.png}\\
-	\label{fig:weather}
+	\caption{\footnotesize {Thermal time characterization of cocoa for five locations.\\}} 
+	\includegraphics[scale=0.3]{images/ttbbox.png}
+	\label{fig:ttbox}
 \end{figure}
 
+\subsection{Weather effects}
 Studying the weather data effects over the flowering date and their final yield.
-We found that the number of successful flowers pollinated to produce final yield can be affected mostly by the rain, TMAX and wind (Figure \ref{fig:correlation}).
-The later can affect the availability of tiny flies pollinators from Diptera order and from the families of of the biting midges \textit{Ceratopogonidae},  genus  \textit{Forcipomyia} \citep{Saunders1959, kaufmann1975, sotomayor2020} to reach the cocoa flowers.
+In a preliminary analysis of Santanter data, We found that the number of successful flowers pollinated to produce final yield can be affected mostly by the rain, TMAX and wind (Figure \ref{fig:correlation}).
+Farmers
+
+\begin{figure}[h!]
+	\centering
+	\caption{\footnotesize {SIMPLE model output for cocoa in Colombia.\\ }} 
+	\includegraphics[scale=0.35]{images/outmodel.png}
+	\label{fig:m1}
+\end{figure}
+
+\subsection{Day of harvest prediction}
+\begin{figure}[h!]
+	\centering
+	\caption{\footnotesize {Day of harvest cocoa simulation\\ }} 
+	\includegraphics[scale=0.4]{images/RegionHarvest.png}
+	\label{fig:dayH}
+\end{figure}
+%\begin{figure}[h!]
+%	\centering
+%	\caption{\footnotesize {Weather variables during flowering times.\\}} 
+%	\includegraphics[scale=0.4]{images/averageS.png}
+%	\label{fig:weather}
+%\end{figure}
+
+\begin{figure}[h!]
+	\centering
+	\caption{\footnotesize {Pearson correlation among weather variables and flowering dates all locations. \\}} 
+	\includegraphics[scale=0.5]{images/heatm.png}\\
+	\label{fig:heat}
+	{\footnotesize thermalT= thermal time without Tb. ttb= thermal time with Tb. WS2M= Wind Speed at 2 Meters (m/s). T2MDEW = Dew/Frost Point at 2 Meters ($^\circ$C) }
+\end{figure}
+\newpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Discussion}
+Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses.
+The findings and their implications should be discussed in the broadest context possible.
+Future research directions may also be highlighted.
+
+The optimum temperature for photosynthesis in cacao has
+been reported to be between 31 °C and 33 °C (Balasimha et al.
+1991) and 33 °C–35 °C (Yapp 1992). Sena Gomes and
+Kozlowski (1987) reported a decrease in gs over the range
+18.7 °C to 27.2 °C and an increase at temperatures above this,
+most likely to enhance cooling at higher temperatures.
+However, the opposite effect was described by Raja Harun
+and Hardwick (1988a). In this study, gs and transpiration (E)
+increased within the range of 20 °C–30 °C. In field-grown
+cacao in India, photosynthetic rate declined as mean monthly
+temperature increased above 34 °C during the dry season
+(Balasimha et al. 1991). \cite{lahive2019}
+\textcolor{red}{this section on going to organize}
+
+Evaluating the level of knowledge of producers regarding cocoa crop management, the harvest was in the group of activities that presented the lowest level of knowledge on the part of the producers according to the general averages \citep{Gutierrez2020}.
+
+\subsection{Weather Effects}
+Wind can affect the availability of tiny flies pollinators from Diptera order and from the families of of the biting midges \textit{Ceratopogonidae},  genus  \textit{Forcipomyia} \citep{Saunders1959, kaufmann1975, sotomayor2020} to reach the cocoa flowers.
 However, wind is not included in the SIMPLE model as an input as cocoa crop has not been simulated with this model before.
@@
 Unfertilised flowers abscise from the trunk approximately 1 day after flower opening  \citep{Niemenak2010}.
-\begin{figure}[h!]
-	\centering
-	\caption{\footnotesize {Pearson correlation among weather variables and flowering dates \\}} 
-	\includegraphics[scale=0.7]{images/correlation.png}\\
-	\label{fig:correlation}
-\end{figure}.
-
-\begin{figure}[h]
-	\centering
-	\caption{\footnotesize {Thermal time characterization of cocoa for four locations\\}} 
-	\includegraphics[scale=0.5]{images/ttbbox.png}
-	\label{fig:ttbox}
-\end{figure}.
-\newpage
-
-\begin{figure}[h!]
-	\centering
-	\caption{\footnotesize {SIMPLE model output for cocoa in Colombia }} 
-	\includegraphics[scale=0.3]{images/outmodel.png}\\
-	\label{fig:m1}
-\end{figure}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-
-Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses.
-The findings and their implications should be discussed in the broadest context possible.
-Future research directions may also be highlighted.
-\textcolor{red}{this section on going to organize}
-
-Evaluating the level of knowledge of producers regarding cocoa crop management, the harvest was in the group of activities that presented the lowest level of knowledge on the part of the producers according to the general averages \citep{Gutierrez2020}.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-10-07 papers/model/paper_model.tex
last changes parameters tables

@@
 This cycle was then repeated until the  simulated yields for cocoa were reasonably close to the observed yield.
-
 In the SIMPLE model dummy files are provided for adding new cocoa crop data and weather data, and files 2 to 6 in the list below can be edited to define new cultivars or experiments.
@@
 
-The first step was add the new experiment for cocoa in the dummy crop and experiment names were replaced with our crop and treatment names in the input files and flowering date was changed.
-Thermal time values required for Input/Cultivar.csv  had to be defined for ICS 95 and CCN51 cultivars using field data from Fedecacao plot reports in Santander and Caldas, Colombia.
-Thus, the date for harvest is predicted when the model calculations reach the thermal time  for  each specific cultivar ICS 95 and CCN51.
-The new experiment name “KOKOlatl” for was used in the species parameter file with the respective values of parameters for cocoa including irrigation, soil features and radiation use efficiency \citep{zuidema2005} \textcolor{red}{add table parameters values will change through the calibration}.
+\subsection{Parameters}
+There were three parameters varied by region (table \ref{tab:reparam}) thermal time required for pod harvest after the flowering date (Tsum) , the Radiation use efficiency (RUE) and yield observed on field.
+Physiological parameters in table \ref{tab:Treaparam} are  common  for all the regions studied.
+These parameter were calibrate for cultivars ICS95 and CCN51 considering a range of time of 200 days (DAP) from flowering date to harvest day, even thought farmers collect the pod at 180 DAP.
+Heat and water stress parameters were not considered.
 
 \begin{table}[h!]	
-\caption {\footnotesize {Cocoa crop parameter values used in Species-parameter file.}} \label{tab:Speparams} 
+	\caption {\footnotesize {Cocoa crop parameter values used per region.}}
+	\label{tab:reparam} 
 	\centering
 	\begin{small}
-		\begin{tabular}{{c c c c c c c c c }}
+		\begin{tabular}{l c c c }
 			\hline
-			{\bf Tbase }& {\bf Topt } & {\bf RUE} & {\bf I50maxH}& {\bf I50maxW}& {\bf MaxT}& {\bf ExtremeT} & {\bf CO2-RUE}& {\bf S-water}\\
+			{\bf Cultivar }&{\bf Tsum }&{\bf RUE}&{\bf Yield$^{*}$}\\
 			\hline
-			10& 26 & 0.6 & 100 & 5& 35&40&0.09&0 \\
+			Santander & 2016 &0.6 & 2687 \\
+			Apartado   & 2906 & 0.6 & 3378  \\
+			Arauca   & 1912 & 0.7 & 3981  \\
+			Cali   & 2764 & 0.5 & 1900  \\
+			Caldas   & 1192 & 0.6 & 740  \\
 			\hline
-		\end{tabular}  
-	{\footnotesize S-water is associated drought stress evaluations ranging from 0 (no water shortage) to 1 (extreme water shortage) \cite{Zao2019simple} }
+		\end{tabular} \\
+		{\footnotesize $^{*}$ Yield observed kg ha$^{-1}$ per year} 
 	\end{small}
@@
 \begin{table}[h!]	
-\caption {\footnotesize {Cocoa crop parameter values used in Cultivar file.}}
-\label{tab:Culparam} 
+	\caption {\footnotesize {Parameter values used to run SIMPLEcocoa model.}}
 	\centering
+     \label{tab:Treaparam} 
 	\begin{small}
-		\begin{tabular}{c c c c c}
+		\begin{tabular}{{l l l}}
 			\hline
-			{\bf Cultivar }&{\bf Tsum }&{\bf HI}&{\bf I50A}&{\bf I50B}\\
+            {\bf File }&{\bf Variable name }&{\bf Value}\\
 			\hline
-			ICS 95 & 1912 & 0.3 & 680 & 200 \\
-			CCN51  & 1912 & 0.3 & 680 & 200 \\
+			&SoilName & Loamy sand4\\
+			&InitialFsolar & 0.01\\
+			Treatment&Weather & KOKO (.WTH file name)\\
+			&CO$_{2}$ & 400\\
+			&SowingDate &Flowering date\\
 			\hline
-		\end{tabular}  
+			&Crop cycle DAP & 200 days\\
+			&LAI & 1.8 \\
+			Observation&FSolar& 0.70\\
+			&Biomass & 40kg dry mass per plant\\
+			\hline
+			&Harvest index & 0.3\\
+			Cultivar&150A & 680 $^\circ$C day \\
+			&150B & 680 $^\circ$C day \\
+			\hline
+			&Tbase & 10$^\circ$C\\
+			&Topti & 26$^\circ$C \\
+			Species&MaxT & 35$^\circ$C \\
+			&ExtremeT & 40$^\circ$C  \\
+			&CO$_{2}$RUE & 0.09$^\circ$C  \\			
+			&S-water & 0 ARID index \\
+			\hline			
+		\end{tabular} \\ 
 	\end{small}
-\end{table}
-
-\begin{table}[h!]	
-	\caption {\footnotesize {Cocoa crop parameter values used in Treatment and Observation file.}} \label{tab:Treaparam} 
-	\centering
-	\begin{small}
-		\begin{tabular}{{c c }}
-			\hline
-			{\bf Variable name }&{\bf Value}\\
-			\hline
-			Exp&KOKOlatl\\
-			Label& Colombia\\
-			Weather & KOKOS (.WTH file name)\\
-			CO2 & 400\\
-			SowingDate &flowering date\\
-			SoilName & Loamy sand4\\
-			Cultivar& ICS 95 or CCN51\\
-			MaxIntercept& 0.95\\
-			InitialBio & 1\\
-			InitialTT & 0\\
-			InitialFsolar & 0.01 \\
-			LAI & 1.8 \\
-			\hline
-		\end{tabular}  
-	\end{small}
+	{\footnotesize S-water is associated drought stress evaluations ranging from 0 (no water shortage) to 1 (extreme water shortage) \cite{Zao2019simple} }
 \end{table}
@@
 Studying the weather data effects over the flowering date and their final yield.
-In a preliminary analysis of Santanter data, We found that the number of successful flowers pollinated to produce final yield can be affected mostly by the rain, TMAX and wind (Figure \ref{fig:correlation}).
+In a preliminary analysis of Santanter data, We found that the number of successful flowers pollinated to produce final yield can be affected mostly by the rain, TMAX and wind (Figure \ref{fig:heat}).
 Farmers
@@
 	\caption{\footnotesize {Pearson correlation among weather variables and flowering dates all locations. \\}} 
-	\includegraphics[scale=0.5]{images/heatm.png}\\
+	\includegraphics[scale=0.4]{images/heatm.png}\\
 	\label{fig:heat}

### 2021-10-07 papers/model/paper_model.tex
Update paper_model.tex

@@
 % Authors, for the paper (add full first names)
-\Author{Angela Romero V $^{1,\dagger,\ddagger}$\orcidA{}, Adriana Gallego $^{1,\ddagger}$ and Anyela V. Camargo R.$^{1,}$*}
+\Author{Angela Romero V $^{1,\dagger,\ddagger}$*\orcidA{}, Adriana Gallego $^{2}$ and Anyela V. Camargo R.$^{1, \ddagger}$}
 
@@
 \address{%
-$^{1}$ \quad The John Bingham Laboratory, NIAB, 93 Lawrence Weaver Road, Cambridge CB3 0LE, UK 1; e-mail@e-mail.com\\
-$^{2}$ \quad BIOS 2; e-mail@e-mail.com}
+$^{1}$ \quad The John Bingham Laboratory, NIAB, 93 Lawrence Weaver Road, Cambridge CB3 0LE, UK 1}\\ %e-mail@e-mail.com\\
+$^{2}$ \quad BIOS Colombia}%; e-mail@e-mail.com}
 
 % Contact information of the corresponding author
-\corres{Correspondence:anyela.camargorodriguez@niab.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}
+\corres{Correspondence: angela.romerovergel@niab.com} %; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}
 
 % Current address and/or shared authorship
-\firstnote{Current address: Affiliation 3} 
+%\firstnote{Current address: Affiliation 3} 
 \secondnote{These authors contributed equally to this work.}

### 2021-10-12 papers/model/paper_model.tex
graficas modificadas

@@
 % The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
-
 %=================================================================
@@
 % Authors, for the paper (add full first names)
-\Author{Angela Romero V $^{1,\dagger,\ddagger}$*\orcidA{}, Adriana Gallego $^{2}$ and Anyela V. Camargo R.$^{1, \ddagger}$}
+\Author{Angela Romero V $^{1,\ddagger}$*\orcidA{}, Adriana Gallego $^{2}$ and Anyela V. Camargo R.$^{1, \ddagger}$}
 
@@
 % Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address{%
-$^{1}$ \quad The John Bingham Laboratory, NIAB, 93 Lawrence Weaver Road, Cambridge CB3 0LE, UK 1}\\ %e-mail@e-mail.com\\
-$^{2}$ \quad BIOS Colombia}%; e-mail@e-mail.com}
+\address{$^{1}$ \quad The John Bingham Laboratory, NIAB, 93 Lawrence Weaver Road, Cambridge CB3 0LE, UK 1\\$^{2}$ \quad BIOS Colombia}%; e-mail@e-mail.com}
 
@@
 \section{Introduction}
-
 Cocoa ( \textit{Theobroma cacao }L.
@@
 	\caption{\footnotesize {Phenology of cocoa in Colombia for crop modelling.\\}} 
-	\includegraphics[scale=0.3]{images/phenology.png}\\
+	\includegraphics[scale=0.4]{images/phenology.png}\\
 	\footnotesize{Credits: Taken from from Dreamstime.com, phys.org \citep{toledo2021} and \cite{lopez2018}}
@@
 	\centering
+	\includegraphics[scale=0.5]{images/clima.png}
 	\caption{\footnotesize {Daily average temperature and available photosynthetic solar radiation (PAR) conditions. \\}}	
-	\includegraphics[scale=0.25]{images/tempe.png}
-	\includegraphics[scale=0.25]{images/SRAD.png}
 	\label{fig:temp}
@@
 	\centering
+	\includegraphics[scale=0.4]{images/rainmonth.png}
 	\caption{\footnotesize {Precipitation and relative humidity per month in Colombia \\}}
-	\includegraphics[scale=0.4]{images/rainmonth.png}
 	\label{fig:rain}
 \end{figure}
-\newpage
 
@@
 	\centering
+	\includegraphics[scale=0.25]{images/ttbbox.png}
 	\caption{\footnotesize {Thermal time characterization of cocoa for five locations.\\}} 
-	\includegraphics[scale=0.3]{images/ttbbox.png}
 	\label{fig:ttbox}
@@
 	\centering
+	\includegraphics[scale=0.35]{images/outmodel.png}
 	\caption{\footnotesize {SIMPLE model output for cocoa in Colombia.\\ }} 
-	\includegraphics[scale=0.35]{images/outmodel.png}
 	\label{fig:m1}
@@
 	\centering
+	\includegraphics[scale=0.4]{images/RegionHarvest2.png}
 	\caption{\footnotesize {Day of harvest cocoa simulation\\ }} 
-	\includegraphics[scale=0.4]{images/RegionHarvest.png}
 	\label{fig:dayH}
@@
 	\centering
-	\caption{\footnotesize {Pearson correlation among weather variables and flowering dates all locations. \\}} 
 	\includegraphics[scale=0.4]{images/heatm.png}\\
 	\label{fig:heat}
+	\caption{\footnotesize {Pearson correlation among weather variables and flowering dates all locations. \\}} 
 	{\footnotesize thermalT= thermal time without Tb. ttb= thermal time with Tb. WS2M= Wind Speed at 2 Meters (m/s). T2MDEW = Dew/Frost Point at 2 Meters ($^\circ$C) }

### 2021-10-14 SIMPLEcocoa_generic/Plot.R
Update Plot.R

@@ -1,149 +1,140 @@
-#########################SIMPLE MODEL v1.1###########################
-
-#######  KOKOlatm Lanitoamenrican cocoa to predict the optimal harvest time 
-#Updated by Angela Romero Vergel,NIAB,Cocoa project ------  2021-08-11 
-##################################################################
-
-#rm(list=ls())   #### cleans memory - needs to saty here
-library(ggplot2)
-library(plyr)
-library(parallel)
-#setwd("C:/Users/Angelita/Documents/SIMPLEcocoa_Generic")
-source("Plot.R")
-source("Mainfunction.R")
-source("Obsfunction.R")
-
-RunModel=function(i){
-  source("Mainfunction.R")
-  paras=ParaInput(i)
-  res<-tryCatch({SIMPLE(para=paras[c(1:3)],weather=paras$weather,
-                        ARID=paras$ARID)},error=function(e){cat("ERROR :", conditionMessage(e),"\n")})
-  return(res)
-}
-
-organise_data <- function(GridsimulationSwitch){
-  
-  if(GridsimulationSwitch=='OFF'){
-    management<-read.table("./Input/Simulation Management.csv",header=TRUE,sep=","
-                           ,stringsAsFactors=FALSE)
-    treatment<-read.table("./Input/Treatment.csv",header=TRUE,sep=","
-                          ,stringsAsFactors=FALSE);treatment$Species.=tolower(treatment$Species.)
-    
-    cultivar<-read.table("./Input/Cultivar.csv",header=TRUE,sep=","
-                         ,stringsAsFactors=FALSE);cultivar$Species.=tolower(cultivar$Species.)
-    
-    irri<-read.table("./Input/Irrigation.csv",header=TRUE,sep=","
-                     ,stringsAsFactors=FALSE);irri$Species.=tolower(irri$Species.)
-    soil<-read.table("./Input/Soil.csv",header=TRUE,sep=",",stringsAsFactors=FALSE)  
-    para_spec<-read.table("./Input/Species parameter.csv",header=TRUE,sep=","
-                          ,stringsAsFactors=FALSE);para_spec$Species.=tolower(para_spec$Species.)
-    
-    ####match experiment
-    management<-management[management$ON_Off==1,]
-    treatment<-merge(treatment,management,by=c("Species.","Exp.","Trt.")
[…]

### 2021-10-14 SIMPLEcocoa_generic/SimpleBcocoa.R
arrows changed

@@ -4,7 +4,7 @@
 #Updated by Angela Romero Vergel,NIAB,Cocoa project ------  2021-08-11 
 ##################################################################
 
-#rm(list=ls())   #### cleans memory - needs to saty here
+rm(list=ls())   #### cleans memory - needs to saty here
 library(ggplot2)
 library(plyr)
 library(parallel)
@@ -13,51 +13,51 @@ source("Plot.R")
 source("Mainfunction.R")
 source("Obsfunction.R")
 
-RunModel=function(i){
+RunModel  <-  function(i){
   source("Mainfunction.R")
-  paras=ParaInput(i)
-  res<-tryCatch({SIMPLE(para=paras[c(1:3)],weather=paras$weather,
-                        ARID=paras$ARID)},error=function(e){cat("ERROR :", conditionMessage(e),"\n")})
+  paras  <-  ParaInput(i)
+  res  <-  tryCatch({SIMPLE(para=paras[c(1:3)],weather=paras$weather,
+    ARID=paras$ARID)},error=function(e){cat("ERROR :", conditionMessage(e),"\n")})
   return(res)
 }
 
 organise_data <- function(GridsimulationSwitch){
   
   if(GridsimulationSwitch=='OFF'){
-    management<-read.table("./Input/Simulation Management.csv",header=TRUE,sep=","
-                           ,stringsAsFactors=FALSE)
-    treatment<-read.table("./Input/Treatment.csv",header=TRUE,sep=","
-                          ,stringsAsFactors=FALSE);treatment$Species.=tolower(treatment$Species.)
+    management <- read.table("./Input/Simulation Management.csv",header=TRUE,sep=","
+                  ,stringsAsFactors=FALSE)
+    treatment <- read.table("./Input/Treatment.csv",header=TRUE,sep=","
+                 ,stringsAsFactors=FALSE);treatment$Species.=tolower(treatment$Species.)
     
-    cultivar<-read.table("./Input/Cultivar.csv",header=TRUE,sep=","
-                         ,stringsAsFactors=FALSE);cultivar$Species.=tolower(cultivar$Species.)
+    cultivar <- read.table("./Input/Cultivar.csv",header=TRUE,sep=","
+             ,stringsAsFactors=FALSE);cultivar$Species.=tolower(cultivar$Species.)
     
[…]

### 2021-10-15 SIMPLEcocoa_generic/Mainfunction.R
Add files via upload

@@ -8,7 +8,8 @@
 
 SIMPLE<-function(para,weather,ARID)
 {
-
+  #para=paras[c(1:3)];weather=paras$weather;ARID=paras$ARID
+#
   #############  read crop and cv parameter
   Trt=para$Species$Trt.
   Tbase<-para$Species$Tbase
@@ -35,10 +36,12 @@ SIMPLE<-function(para,weather,ARID)
   FHarv<-400
   
   
-  if(is.na(harvestDate)){DUA=NA}else{
+  if(is.na(harvestDate)){
+    DUA=NA
+  }else {
     DUA<-as.numeric(harvestDate-sowingDate)+1
     weather<-weather[1:DUA,]
-    ARID<-ARID[1:DUA,]
+    ARID1<-ARID[1:DUA,]
   }
   
   stopday<-ifelse(nrow(weather)>FHarv,FHarv,nrow(weather))
@@ -47,9 +50,11 @@ SIMPLE<-function(para,weather,ARID)
   MaturityDay2<-stopday
   
   
-  Result=data.frame(Day = c(1:stopday),DATE=weather$IDATE[1:stopday],ETO=rep(0,stopday), TT=rep(0,stopday),Biomass=rep(0,stopday),
-                    Tmax=weather$TMAX[1:stopday],Tmin=weather$TMIN[1:stopday],Radiation=weather$SRAD[1:stopday],
-                    HI=rep(HIp,stopday),ARID=ARID[1:stopday,2],Trt=rep(Trt,stopday))
+  Result=data.frame(Day = c(1:stopday),DATE=weather$IDATE[1:stopday],
+                    ETO=rep(0,stopday), TT=rep(0,stopday),Biomass=rep(0,stopday),
+                    Tmax=weather$TMAX[1:stopday],Tmin=weather$TMIN[1:stopday],
+                    Radiation=weather$SRAD[1:stopday], HI=rep(HIp,stopday), 
+                    ARID=ARID[1:stopday,2], Trt=rep(Trt,stopday))
   
   Result$F_Water=mapply(Water_Response,ARID$ARID,S_Water);if(!water){Result$F_Water=1}
   Result$dETO<-ARID$ETO
@@ -203,11 +208,14 @@ ParaInput=function(i)
 {
   para <-list()
   ###treatment filter 
-  treat<-treatment[i,];print(paste(treat$Exp[1],treat$Trt[1]))
+  #treat <- i$treatment[i,];
[…]

### 2021-10-15 SIMPLEcocoa_generic/Plot.R

@@ -81,19 +81,21 @@ gplot=function(Res_Daily,Res_Summary,Obs_Biomass,Obs_FSolar){
     #   theme(legend.position = "none",
     #         plot.title=element_text(size=14, hjust=0, vjust=0))
     
-    cosecha=Res_daily$MaturityDay[1]
+    cosecha=Res_daily$MaturityDay[1]# FIXME Angela
     R_summary<-Res_Summary[,c(-6,-7)] 
     R_summary<-R_summary[order(R_summary$Crop),]
-    print(R_summary)
+    R_summary2<-Res_Summary[,c(-5,-6,-7)] # FIXME Angela
+    print(R_summary2)# FIXME Angela
     
     ##### calculate RMSE and RRMSE for yield
     R_summary<-R_summary[!is.na(R_summary$Obs_Yield),]
     RMSE_yield<-sqrt(mean((R_summary$Sim_Yield-R_summary$Obs_Yield)^2))
     RRMSE_yield<-RMSE_yield/mean(R_summary$Obs_Yield)
     
-    print(paste("RMSE for yield is: ",round(RMSE_yield,0),"kg.ha-1"))
-    print(paste("RRMSE for yield is: ",100*round(RRMSE_yield,3),"%",sep=""))
+    #print(paste("RMSE for yield is: ",round(RMSE_yield,0),"kg.ha-1"))
+    #print(paste("Error RRMSE for yield is: ",100*round(RRMSE_yield,3),"%",sep=""))
     print(paste("Dia de cosecha:", cosecha))
+    print(paste("Rendimiento potencial:", Res_summary$Yield,"kg.ha-1 año"))
     
     Obs_Biomass$LableCrop=paste0(Obs_Biomass$Crop,"_",Obs_Biomass$Label)
     Obs_FSolar$LableCrop=paste0(Obs_FSolar$Crop,"_",Obs_FSolar$Label)
@@ -137,4 +139,5 @@ gplot=function(Res_Daily,Res_Summary,Obs_Biomass,Obs_FSolar){
     # windows(width=16, height=8)
     # multiplot(P5, P3, P4, cols=3)
 }
+
 

### 2021-10-15 SIMPLEcocoa_generic/SimpleBcocoa.R

@@ -1,42 +1,52 @@
+
 #########################SIMPLE MODEL v1.1###########################
 
 #######  KOKOlatm Lanitoamenrican cocoa to predict the optimal harvest time 
 #Updated by Angela Romero Vergel,NIAB,Cocoa project ------  2021-08-11 
 ##################################################################
 
-rm(list=ls())   #### cleans memory - needs to saty here
+#rm(list=ls())   #### cleans memory - needs to saty here
 library(ggplot2)
 library(plyr)
 library(parallel)
-#setwd("C:/Users/Angelita/Documents/SIMPLEcocoa_Generic")
+#setwd("C:/Users/Angelita/Documents/SIMPLEcocoa_NOirri")
 source("Plot.R")
 source("Mainfunction.R")
 source("Obsfunction.R")
 
-RunModel  <-  function(i){
+RunModel = function(i){
   source("Mainfunction.R")
-  paras  <-  ParaInput(i)
-  res  <-  tryCatch({SIMPLE(para=paras[c(1:3)],weather=paras$weather,
-    ARID=paras$ARID)},error=function(e){cat("ERROR :", conditionMessage(e),"\n")})
+  paras <- ParaInput(i)
+  res<-tryCatch({SIMPLE(para=paras[c(1:3)],weather=paras$weather,
+                          ARID=paras$ARID)},error=function(e){cat("ERROR :", conditionMessage(e),"\n")})
   return(res)
 }
+
+#FIXME Angela
+#pass date input to DOY
+fecha_floracion <- as.Date("2020-06-23") 
+date <- format(fecha_floracion, format="%Y-%b-%d")
+year <- (format(fecha_floracion, format="%y"))
+day = (format(fecha_floracion, format="%j"))
+doy=as.numeric(paste(year, day, sep = ''))
 
 organise_data <- function(GridsimulationSwitch){
   
   if(GridsimulationSwitch=='OFF'){
     management <- read.table("./Input/Simulation Management.csv",header=TRUE,sep=","
-                  ,stringsAsFactors=FALSE)
+                           ,stringsAsFactors=FALSE)
     treatment <- read.table("./Input/Treatment.csv",header=TRUE,sep=","
-                 ,stringsAsFactors=FALSE);treatment$Species.=tolower(treatment$Species.)
-    
+                          ,stringsAsFactors=FALSE);treatment$Species.=tolower(treatment$Species.)
+    treatment$SowingDate<-doy
[…]

### 2021-10-15 SIMPLEcocoa_generic/SimpleBcocoa.R
Update SimpleBcocoa.R

@@ -24,7 +24,7 @@ RunModel = function(i){
 
 #FIXME Angela
 #pass date input to DOY
-fecha_floracion <- as.Date("2020-06-23") 
+fecha_floracion <- as.Date("2020-06-23") #enter here date of flowering
 date <- format(fecha_floracion, format="%Y-%b-%d")
 year <- (format(fecha_floracion, format="%y"))
 day = (format(fecha_floracion, format="%j"))

### 2021-10-15 SIMPLEcocoa_generic/SimpleBcocoa.R
Update SimpleBcocoa.R

@@ -24,7 +24,8 @@ RunModel = function(i){
 
 #FIXME Angela
 #pass date input to DOY
-fecha_floracion <- as.Date("2020-06-23") #enter here date of flowering
+#enter here date of flowering year-moth-day
+fecha_floracion <- as.Date("2020-06-23") 
 date <- format(fecha_floracion, format="%Y-%b-%d")
 year <- (format(fecha_floracion, format="%y"))
 day = (format(fecha_floracion, format="%j"))

### 2021-10-15 SIMPLEcocoa_generic/SimpleBcocoa.R
Update SimpleBcocoa.R

@@ -24,7 +24,7 @@ RunModel = function(i){
 
 #FIXME Angela
 #pass date input to DOY
-#enter here date of flowering year-moth-day
+#enter here date of flowering year-month-day
 fecha_floracion <- as.Date("2020-06-23") 
 date <- format(fecha_floracion, format="%Y-%b-%d")
 year <- (format(fecha_floracion, format="%y"))

### 2021-10-15 SIMPLEcocoa_generic/SimpleBcocoa.R
Update SimpleBcocoa.R

@@ -27,8 +27,8 @@ RunModel = function(i){
 #enter here date of flowering year-month-day
 fecha_floracion <- as.Date("2020-06-23") 
 date <- format(fecha_floracion, format="%Y-%b-%d")
-year <- (format(fecha_floracion, format="%y"))
-day = (format(fecha_floracion, format="%j"))
+year <- format(fecha_floracion, format="%y")
+day = format(fecha_floracion, format="%j")
 doy=as.numeric(paste(year, day, sep = ''))
 
 organise_data <- function(GridsimulationSwitch){

### 2021-10-15 SIMPLEcocoa_generic/SimpleBcocoa.R
Update SimpleBcocoa.R

@@ -26,7 +26,6 @@ RunModel = function(i){
 #pass date input to DOY
 #enter here date of flowering year-month-day
 fecha_floracion <- as.Date("2020-06-23") 
-date <- format(fecha_floracion, format="%Y-%b-%d")
 year <- format(fecha_floracion, format="%y")
 day = format(fecha_floracion, format="%j")
 doy=as.numeric(paste(year, day, sep = ''))

### 2021-10-15 SIMPLEcocoa_generic/SimpleBcocoa.R
Update SimpleBcocoa.R

@@ -27,8 +27,8 @@ RunModel = function(i){
 #enter here date of flowering year-month-day
 fecha_floracion <- as.Date("2020-06-23") 
 year <- format(fecha_floracion, format="%y")
-day = format(fecha_floracion, format="%j")
-doy=as.numeric(paste(year, day, sep = ''))
+day <- format(fecha_floracion, format="%j")
+doy <- as.numeric(paste(year, day, sep = ''))
 
 organise_data <- function(GridsimulationSwitch){
   
@@ -37,7 +37,7 @@ organise_data <- function(GridsimulationSwitch){
                            ,stringsAsFactors=FALSE)
     treatment <- read.table("./Input/Treatment.csv",header=TRUE,sep=","
                           ,stringsAsFactors=FALSE);treatment$Species.=tolower(treatment$Species.)
-    treatment$SowingDate<-doy
+    treatment$SowingDate <- doy
     cultivar <- read.table("./Input/Cultivar.csv",header=TRUE,sep=","
                          ,stringsAsFactors=FALSE);cultivar$Species.=tolower(cultivar$Species.)
     

### 2021-11-03 SIMPLEcocoa_generic/Mainfunction.R
Delete SIMPLEcocoa_generic directory

@@ -1,458 +1,0 @@
-#########################  Main model  ###################################################################################################################
-# ********************* Reference for a developer *****************************************
-# ET model can be selected using "petmodel". 
-# The value of petmodel is assigned by PET_OPTION, which is defined in the top section.
-# no change is needed here to selet a ET model. 
-##################################################################################################################
-
-
-SIMPLE<-function(para,weather,ARID)
-{
-  #para=paras[c(1:3)];weather=paras$weather;ARID=paras$ARID
-#
-  #############  read crop and cv parameter
-  Trt=para$Species$Trt.
-  Tbase<-para$Species$Tbase
-  Topt<-para$Species$Topt
-  RUE<-para$Species$RUE
-  I50maxH<-para$Species$I50maxH
-  I50maxW<-para$Species$I50maxW
-  maxT<-para$Species$MaxT
-  extremeT<-para$Species$ExtremeT
-  S_Water<-para$Species$S_Water 
-  CO2_RUE<-para$Species$CO2_RUE
-  
-  Tsum<-para$Cultivar$Tsum
-  HIp<-para$Cultivar$HI
-  I50A<-para$Cultivar$I50A
-  I50B<-para$Cultivar$I50B
-  
-  inifsolar<-para$treatment$InitialFsolar
-  Fsolarmax<-ifelse(is.na(para$treatment$MaxIntercept),0.95,para$treatment$MaxIntercept)
-  sowingDate<-para$treatment$SowingDate
-  harvestDate<-para$treatment$HarvestDate
-  CO2<-para$treatment$CO2
-  water<-para$treatment$Water
-  FHarv<-400
-  
-  
-  if(is.na(harvestDate)){
-    DUA=NA
-  }else {
-    DUA<-as.numeric(harvestDate-sowingDate)+1
-    weather<-weather[1:DUA,]
-    ARID1<-ARID[1:DUA,]
-  }
-  
-  stopday<-ifelse(nrow(weather)>FHarv,FHarv,nrow(weather))
-  weather<-weather[1:stopday,]
-  ARID<-ARID[1:stopday,]
-  MaturityDay2<-stopday
[…]

### 2021-11-03 SIMPLEcocoa_generic/MapPlot.R

@@ -1,78 +1,0 @@
-list.of.packages <- c("RColorBrewer", "viridis","raster","RColorBrewer","rasterVis")
-new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
-if(length(new.packages)) install.packages(new.packages)
-
-library(RColorBrewer)
-library(viridis)  
-library(raster)
-library(RColorBrewer)
-library(rasterVis)
-
-
-Transfer_to_raster=function(SummaryData,index){
-  newnarry=array(NA,dim = c(360,720))
-  newnarry=as.matrix(newnarry)
-  dim(newnarry)
-  
-  for(i in 1:nrow(SummaryData))
-  {
-    rows=SummaryData$row[i]
-    cols=SummaryData$col[i]
-    eval(parse(text=paste0("newnarry[rows,cols]<-SummaryData$",index,"[i]")))
-  }
-  
-  newras=raster(newnarry,xmn=-180,xmx=180,ymn=-90,ymx=90)
-  proj4string(newras)=CRS("+proj=longlat +datum=WGS84")
-  return(newras)
-}
-
-MapPlot=function(yieldirra=Res_Summary,index="Yield",WMap,MapExtention,Title="Yield",Unit){
-  
-  ras=Transfer_to_raster(yieldirra,index = index)
-  
-  if(!MapExtention=="World"){
-    ras <- mask(x=ras, mask=WMap)
-    ras<-crop(ras,WMap)}
-  b2p1 <- colorRampPalette(c("green","sky blue","blue","purple","yellow","Orange Red","red"))
-  #b2p1 <- colorRampPalette(c("grey90","grey40","black"))
-  pal8 <-  b2p1(100)
-  
-  if(MapExtention==c("China")){
-    load("H:/NC/China.RData")
-    Wapad=WMap
-    WMap=boundary
-    
-  }
-  
-  
-  #######Simply map
-  library(rgeos)
-  WMap_sim2 <- gSimplify(WMap, tol = .1, 
-                               topologyPreserve = TRUE)
-  
-  mapTheme <- rasterTheme(region=pal8)
-  windows(width=16, height=8)
-  par(mar=c(4,10,4,10))
-  plot(WMap_sim2,axes = TRUE,main=Title,cex.axis=1.3)
-  plot(ras,col=b2p1(100),#labels=pretty(0:14000),#main="Yield", axes=FALSE,
[…]

### 2021-11-03 SIMPLEcocoa_generic/Obsfunction.R

@@ -1,37 +1,0 @@
-ObsInput=function(i){
-  Obsdata <-list()
-  ###treatment filter 
-  treat<-treatment[i,];print(paste(treat$Exp[1],treat$Trt[1]))
-  #########
-  obs<-read.table(paste("./Observation/Obs_",treat$Species,"_",treat$Exp,".csv",sep=""),header=TRUE,sep=",",stringsAsFactors=FALSE)
-  obs<-obs[obs$Trt. %in% (treat$Trt.),]
-  obs$Exp=treat$Exp.
-  obs$Label=treat$Label
-  obs$Trt=obs$Trt.  
-  #source("Mainfunction.R")
-  sowingDate=DOYtoDate(treat$SowingDate)
-  obs$Date=sowingDate+obs$DAP
-  Summary=obs[nrow(obs),]
-  obs$Crop=treat$Species.
-  
-  Obsdata$Biomass=obs[complete.cases(obs$Biomass),c("Date","DAP","Biomass","Trt","Exp","Crop","Label")]
-  
-  Obsdata$FSolar=obs[complete.cases(obs$FSolar),c("Date","DAP","FSolar","Trt","Exp","Crop","Label")]
-  
-  Obsdata$Yield=obs[complete.cases(obs$Yield),c("Date","DAP","Yield","Trt","Exp","Crop","Label")]
-  
-  Obsdata$Summary=Summary[,c("Date","DAP","Biomass","FSolar","Yield","Trt","Exp","Label")]
-  
-  return(Obsdata)
-}
-
-DOYtoDate=function(DATE)
-{
-  DATE=sprintf("%05d",as.numeric(DATE))
-  YEAR<-as.numeric(substring(DATE,1,2))
-  YEAR[YEAR>20]<-YEAR[YEAR>20]+1900;YEAR[YEAR<=20]<-YEAR[YEAR<=20]+2000
-  DOY=as.numeric(substring(DATE,3,5))
-  DATE=as.Date(DOY-1, origin = paste0(YEAR,"-01-01"))
-  return(DATE)
-}
-

### 2021-11-03 SIMPLEcocoa_generic/Plot.R

@@ -1,143 +1,0 @@
-#####Plot modified by A Romero
-
-
-Plot_theme <-function(..., bg='transparent')
-{
-  require(grid)
-  theme_bw(...) +
-    theme( rect=element_rect(fill=bg),
-           legend.position="right",  
-           legend.text=element_text(size=12,colour="black"), 
-           legend.key=element_blank(), 
-           plot.title=element_text(size=12, colour="black",face="bold",hjust=0.5, vjust=0), 
-           plot.margin=unit(rep(0.5,4), 'lines'), 
-           panel.background=element_rect(fill='transparent', color='transparent'),
-           panel.grid=element_blank(),
-           axis.title=element_text(size=15,colour="black",  face="bold",vjust=0.1),
-           axis.line = element_blank(),
-           axis.ticks = element_line(colour = "black", size = 1, linetype = 1),
-           axis.text=element_text(size=12,colour="black"), 
-           axis.text.x = element_text(margin=margin(15,15,10,5,"pt"),face="bold",angle=0), 
-           axis.text.y = element_text(margin=margin(15,15,10,5,"pt"),face="bold"),
-           strip.text=element_text(size=12,colour="black"), 
-           strip.background=element_blank()   
-    )
-  
-}
-theme_set(Plot_theme()) #####set theme by Liujun
-
-####mutiplot 
-multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
-  library(grid)
-  
-  # Make a list from the ... arguments and plotlist
-  plots <- c(list(...), plotlist)
-  
-  numPlots = length(plots)
-  
-  # If layout is NULL, then use 'cols' to determine layout
-  if (is.null(layout)) {
-    # Make the panel
-    # ncol: Number of columns of plots
-    # nrow: Number of rows needed, calculated from # of cols
-    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
-                     ncol = cols, nrow = ceiling(numPlots/cols))
-  }
-  
-  if (numPlots==1) {
-    print(plots[[1]])
-    
-  } else {
-    # Set up the page
-    grid.newpage()
[…]

### 2021-11-03 SIMPLEcocoa_generic/SimpleBcocoa.R

@@ -1,168 +1,0 @@
-
-#########################SIMPLE MODEL v1.1###########################
-
-#######  KOKOlatm Lanitoamenrican cocoa to predict the optimal harvest time 
-#Updated by Angela Romero Vergel,NIAB,Cocoa project ------  2021-08-11 
-##################################################################
-
-#rm(list=ls())   #### cleans memory - needs to saty here
-library(ggplot2)
-library(plyr)
-library(parallel)
-#setwd("C:/Users/Angelita/Documents/SIMPLEcocoa_NOirri")
-source("Plot.R")
-source("Mainfunction.R")
-source("Obsfunction.R")
-
-RunModel = function(i){
-  source("Mainfunction.R")
-  paras <- ParaInput(i)
-  res<-tryCatch({SIMPLE(para=paras[c(1:3)],weather=paras$weather,
-                          ARID=paras$ARID)},error=function(e){cat("ERROR :", conditionMessage(e),"\n")})
-  return(res)
-}
-
-#FIXME Angela
-#pass date input to DOY
-#enter here date of flowering year-month-day
-fecha_floracion <- as.Date("2020-06-23") 
-year <- format(fecha_floracion, format="%y")
-day <- format(fecha_floracion, format="%j")
-doy <- as.numeric(paste(year, day, sep = ''))
-
-organise_data <- function(GridsimulationSwitch){
-  
-  if(GridsimulationSwitch=='OFF'){
-    management <- read.table("./Input/Simulation Management.csv",header=TRUE,sep=","
-                           ,stringsAsFactors=FALSE)
-    treatment <- read.table("./Input/Treatment.csv",header=TRUE,sep=","
-                          ,stringsAsFactors=FALSE);treatment$Species.=tolower(treatment$Species.)
-    treatment$SowingDate <- doy
-    cultivar <- read.table("./Input/Cultivar.csv",header=TRUE,sep=","
-                         ,stringsAsFactors=FALSE);cultivar$Species.=tolower(cultivar$Species.)
-    
-   # irri<-read.table("./Input/Irrigation.csv",header=TRUE,sep=","
-    #                 ,stringsAsFactors=FALSE);irri$Species.=tolower(irri$Species.)
-    soil <- read.table("./Input/Soil.csv",header=TRUE,sep=",",stringsAsFactors=FALSE)  
-    para_spec <- read.table("./Input/Species parameter.csv",header=TRUE,sep=","
[…]

### 2021-11-03 SIMPLEcocoa_genericAPP/AppCocoa.R
Add files via upload

@@ -1,0 +1,14 @@
+
+#######  KOKOlatm Lanitoamenrican cocoa to predict the optimal harvest time 
+# Updated by Angela Romero Vergel,Specialist Crop modeller 
+# NIAB,Cocoa Latam project  ------  2021-08-11 
+##################################################################
+# para correr debe seleccionar todo y dar click en RUN
+# Regiones disponibles:Santander, Apartado, Arauca, Cali o Caldas
+# si el cultivo esta en otra region diferente a las 5 anteriores:
+# escribir "otro"
+region <- "Caldas" 
+#enter here date of flowering year-month-day
+fecha_floracion <- as.Date("2020-06-23") 
+
+source("SimpleB_cocoa.R")

### 2021-11-03 SIMPLEcocoa_genericAPP/Mainfunction.R

@@ -1,0 +1,459 @@
+#########################  Main model  ###################################################################################################################
+# ********************* Reference for a developer *****************************************
+# ET model can be selected using "petmodel". 
+# The value of petmodel is assigned by PET_OPTION, which is defined in the top section.
+# no change is needed here to selet a ET model. 
+##################################################################################################################
+
+
+SIMPLE<-function(para,weather,ARID)
+{
+  #para=paras[c(1:3)];weather=paras$weather;ARID=paras$ARID
+#
+  #############  read crop and cv parameter
+  Trt=para$Species$Trt.
+  Tbase<-para$Species$Tbase
+  Topt<-para$Species$Topt
+  RUE<-para$Species$RUE
+  I50maxH<-para$Species$I50maxH
+  I50maxW<-para$Species$I50maxW
+  maxT<-para$Species$MaxT
+  extremeT<-para$Species$ExtremeT
+  S_Water<-para$Species$S_Water 
+  CO2_RUE<-para$Species$CO2_RUE
+  
+  Tsum<-para$Cultivar$Tsum
+  HIp<-para$Cultivar$HI
+  I50A<-para$Cultivar$I50A
+  I50B<-para$Cultivar$I50B
+  
+  inifsolar<-para$treatment$InitialFsolar
+  Fsolarmax<-ifelse(is.na(para$treatment$MaxIntercept),0.95,para$treatment$MaxIntercept)
+  sowingDate<-para$treatment$SowingDate
+  harvestDate<-para$treatment$HarvestDate
+  CO2<-para$treatment$CO2
+  water<-para$treatment$Water
+  FHarv<-400
+  
+  
+  if(is.na(harvestDate)){
+    DUA=NA
+  }else {
+    DUA<-as.numeric(harvestDate-sowingDate)+1
+    weather<-weather[1:DUA,]
+    ARID1<-ARID[1:DUA,]
+  }
+  
+  stopday<-ifelse(nrow(weather)>FHarv,FHarv,nrow(weather))
+  weather<-weather[1:stopday,]
+  ARID<-ARID[1:stopday,]
[…]

### 2021-11-03 SIMPLEcocoa_genericAPP/Obsfunction.R

@@ -1,0 +1,40 @@
+ObsInput=function(i){
+  i=1
+  Obsdata <-list()
+  ###treatment filter 
+  treat<-treatment[i,];#print(paste(treat$Exp[1],treat$Trt[1]))
+  #########
+  obs<-read.table(paste("./Observation/Obs_",treat$Species,"_",treat$Exp,".csv",sep=""),header=TRUE,sep=",",stringsAsFactors=FALSE)
+  obs<-obs[obs$Trt. %in% (treat$Trt.),]
+  obs$Exp=treat$Exp.
+  obs$Label=treat$Label
+  obs$Trt=obs$Trt.  
+  #source("Mainfunction.R")
+  sowingDate=DOYtoDate(treat$SowingDate)
+  obs$Date=sowingDate+obs$DAP
+  Summary=obs[nrow(obs),]
+  obs$Crop=treat$Species.
+  
+  Obsdata$Biomass=obs[complete.cases(obs$Biomass),c("Date","DAP","Biomass","Trt","Exp","Crop","Label")]
+  
+  Obsdata$FSolar=obs[complete.cases(obs$FSolar),c("Date","DAP","FSolar","Trt","Exp","Crop","Label")]
+  
+  Obsdata$Yield=obs[complete.cases(obs$Yield),c("Date","DAP","Yield","Trt","Exp","Crop","Label")]
+  
+  Obsdata$Summary=Summary[,c("Date","DAP","Biomass","FSolar","Yield","Trt","Exp","Label")]
+  
+  
+  return(Obsdata)
+ 
+}
+
+DOYtoDate=function(DATE)
+{
+  DATE=sprintf("%05d",as.numeric(DATE))
+  YEAR<-as.numeric(substring(DATE,1,2))
+  YEAR[YEAR>20]<-YEAR[YEAR>20]+1900;YEAR[YEAR<=20]<-YEAR[YEAR<=20]+2000
+  DOY=as.numeric(substring(DATE,3,5))
+  DATE=as.Date(DOY-1, origin = paste0(YEAR,"-01-01"))
+  return(DATE)
+}
+

### 2021-11-03 SIMPLEcocoa_genericAPP/Plot.R

@@ -1,0 +1,143 @@
+#####Plot modified by A Romero
+
+
+Plot_theme <-function(..., bg='transparent')
+{
+  require(grid)
+  theme_bw(...) +
+    theme( rect=element_rect(fill=bg),
+           legend.position="right",  
+           legend.text=element_text(size=12,colour="black"), 
+           legend.key=element_blank(), 
+           plot.title=element_text(size=12, colour="black",face="bold",hjust=0.5, vjust=0), 
+           plot.margin=unit(rep(0.5,4), 'lines'), 
+           panel.background=element_rect(fill='transparent', color='transparent'),
+           panel.grid=element_blank(),
+           axis.title=element_text(size=15,colour="black",  face="bold",vjust=0.1),
+           axis.line = element_blank(),
+           axis.ticks = element_line(colour = "black", size = 1, linetype = 1),
+           axis.text=element_text(size=12,colour="black"), 
+           axis.text.x = element_text(margin=margin(15,15,10,5,"pt"),face="bold",angle=0), 
+           axis.text.y = element_text(margin=margin(15,15,10,5,"pt"),face="bold"),
+           strip.text=element_text(size=12,colour="black"), 
+           strip.background=element_blank()   
+    )
+  
+}
+theme_set(Plot_theme()) #####set theme by Liujun
+
+####mutiplot 
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  library(grid)
+  
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+  
+  numPlots = length(plots)
+  
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                     ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+  
+  if (numPlots==1) {
+    print(plots[[1]])
+    
+  } else {
+    # Set up the page
+    grid.newpage()
[…]

### 2021-11-03 SIMPLEcocoa_genericAPP/SimpleB_cocoa.R

@@ -1,0 +1,111 @@
+
+#########################SIMPLE MODEL cocoa ###########################
+
+#######  KOKOlatm Lanitoamenrican cocoa to predict the optimal harvest time 
+#Updated by Angela Romero Vergel,NIAB,Cocoa project ------  2021-08-11 
+##################################################################
+
+library(ggplot2)
+library(plyr)
+library(grid)
+source("Plot.R")
+source("Mainfunction.R")
+source("Obsfunction.R")
+
+RunModel = function(i){
+  source("Mainfunction.R")
+  paras <- ParaInput(i)
+  res<-tryCatch({SIMPLE(para=paras[c(1:3)],weather=paras$weather,
+                          ARID=paras$ARID)},error=function(e){cat("ERROR :", conditionMessage(e),"\n")})
+  return(res)
+}
+
+
+#INPUTS from app
+# add here region:Santander, Apartado, Arauca, Cali o Caldas
+#region= "Caldas" 
+#pass date input to DOY
+#enter here date of flowering year-month-day
+#fecha_floracion <- as.Date("2020-06-23") 
+#date <- format(fecha_floracion, format="%Y-%b-%d")
+year <- format(fecha_floracion, format="%y")
+day <- format(fecha_floracion, format="%j")
+doy <- as.numeric(paste(year, day, sep = ''))
+
+
+organise_data <- function(GridsimulationSwitch){
+  
+  if(GridsimulationSwitch=='OFF'){
+    management<-read.table("./Input/Simulation Management.csv",header=TRUE,sep=","
+                           ,stringsAsFactors=FALSE)
+    treatment<-read.table("./Input/Treatment.csv",header=TRUE,sep=","
+                          ,stringsAsFactors=FALSE);treatment$Species.=tolower(treatment$Species.)
+    treatment$SowingDate <- doy #added Angela
+    cultivar<-read.table("./Input/Cultivar.csv",header=TRUE,sep=","
+                         ,stringsAsFactors=FALSE);cultivar$Species.=tolower(cultivar$Species.)
+    
+    soil<-read.table("./Input/Soil.csv",header=TRUE,sep=",",stringsAsFactors=FALSE)  
+    para_spec<-read.table("./Input/Species parameter.csv",header=TRUE,sep=","
+                          ,stringsAsFactors=FALSE);
+    para_spec$Species.=tolower(para_spec$Species.)
+    
+    ####match experiment
+    management<-management[management$Label==region,]
[…]

### 2021-11-03 papers/model/paperModel.tex
Add files via upload

--- 2021-10-12 papers/model/paper_model.tex
@@
 The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
-
+\usepackage{hyperref}
+\usepackage[utf8]{inputenc}
 %=================================================================
@@
 % Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address{$^{1}$ \quad The John Bingham Laboratory, NIAB, 93 Lawrence Weaver Road, Cambridge CB3 0LE, UK 1\\$^{2}$ \quad BIOS Colombia}%; e-mail@e-mail.com}
+\address{$^{1}$ \quad The John Bingham Laboratory, NIAB, 93 Lawrence Weaver Road, Cambridge CB3 0LE, UK 1\\$^{2}$ \quad Grupo BIOS, Centro de Bioinformática y Biología Computacional de Colombia - BIOS, Manizales, Caldas, Colombia, South America.
+}%; e-mail@e-mail.com}
 
@@
 Cocoa production systems in Latin America has a high importance over social and economic development, facing  the fight against hunger and poverty.
-Cocoa fruits development can be considered as the result of a number of physiological and morphological processes that can be described by mathematical relationships even under uncontrolled environments.
-We parametrized the SIMPLE crop model \citep{Zao2019simple} to predict the best time for harvest cocoa fruits in Colombia.
-The aim of this study is develop a practical tool for farmers using flowering date and weather variables (Solar radiation, rain and temperature) easily available for crop modelling to understand and predict the phenology of cacao trees which is key to produce high quality cacao beans.
-\textcolor{blue}{need to complete}.
+Although Colombian cocoa has the potential to be in the high value markets for fine flavour, it is still not widely produced as the lack of adoptions of technologies by the traditional farmers.
+They empirically harvest after 5 or 6 months after flowering date.
+However, cocoa fruits development can be considered as the result of a number of physiological and morphological processes that can be described by mathematical relationships even under uncontrolled environments.
+Thus, we parametrized the SIMPLE crop model \citep{Zao2019simple} to predict the best time for harvest cocoa fruits in Colombia.
+The results showed an RRMSE of 7.2\% for the yield prediction, while the simulated harvest date varied between +/- 2 to 20 days depending on the temperature variations of the year between regions.
+This crop model application contributed to understand and predict the phenology of cacao fruits the varieties ICS95 y CCN51, which is key to produce high quality cacao beans.
+The aim of this study was developed a practical tool for farmers on cocoa fields to predict the best moment to harvest, using easily available data for crop modelling such as flowering date and weather variables (solar radiation, rain, and temperature).
 }
@@
 In Colombia, cocoa  is  traditionally  consumed  as  a  beverage.
-It is one of the crops promoted by the Colombian government in the social and agricultural development  programs aimed at favouring peace in post-conflict regions \citep{Rodriguez2019, Abbott2019} as cocoa is grown by approximately 52.000 \citep{Gutierrez2020} and 98\% of production being carried out by small and medium-sized producers \citep{Garcia2014, Escobar2020}.
+It is one of the crops promoted by the Colombian government in the social and agricultural development  programs aimed at favouring peace in post-conflict regions \citep{Rodriguez2019, Abbott2019}.
+This crop is grown by approximately 52.000 families \citep{Gutierrez2020} and 98\% of production being carried out by small and medium-sized producers \citep{Garcia2014, Escobar2020}.
 Colombia registered an increase of 3.750 tons in production in 2020 compared to the previous year \citep{lamos2020}.
@@
 We used the SIMPLE crop model \citep{Zao2019simple} for three reasons: 1: That it is very comprehensively described in the original paper.
-2: That the code was available in R for initial trials and 3: That it had already been successfully fitted to perennial crops in south America.
+2: That the code was available in R for initial trials and 3: That it had already been successfully fitted to other perennial crops in south America.
 Overall, the model simulates crop development, growth and yield, and predict the maturation day when the fruit is ready to harvest.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Materials and Methods}
@@
 Normally the phenological stages in cocoa are  divided in two main phases: vegetative and reproductive.
-In the SIMPLE model we simulate reproductive phase described by the floral phenology from the date of inflorescence emergence (BBCH scale 5) to predict the date of ripening of fruit and seed (BBCH 8) \citep{Niemenak2010} (See figure \ref{fig:pheno}).
+In the SIMPLE model we simulate the reproductive phase (fig.
+\ref{fig:pheno}, a) described by the floral phenology from the date of inflorescence emergence (BBCH scale 5) (fig.
+\ref{fig:pheno},b ) to predict the date of ripening of fruit and seed (BBCH 8) (fig.
+\ref{fig:pheno},c ) \citep{Niemenak2010} .
 In the Andean region the reproductive phase is cyclically fulfilled during two annual cycles passing by the following phases: inflorescence emergence, flowering, pollination, fruit development and harvest.
 Therefore, for modelling parametrization the crop cycle of cocoa as perennial plant does not start at the plantation date such as annual crops systems.
-Instead  the start point of the cocoa crop cycle to model is the inflorescence emergence date.
+Instead  the start point of the cocoa crop cycle to model is the inflorescence emergence date (fig.
+\ref{fig:pheno},b).
 Consequently , the growth period of the fruit can varied from  110 to 150  daa (days after anthesis) \citep{lopez2018} when cacao fruits reaches the physiological maturity, but it can be harvested at 170 days daa \citep{Niemenak2010} for quality purposes.
@@
 Cocoa is a cauliflorous plant, which means that flowers grow on the trunk and branches.
-In Colombia cocoa trees usually produce flowers throughout the year, however the biggest flowering occurs in September and January to harvest in March and July (Fig.\ref{fig:yield}).
-Cocoa trees produces with up to 10000 flowers per tree each year.
-which the 50 \%  do not develop into ripe fruits according to Fedecacao reports.
-The flower takes 30 days passing by 12 micro-stages from meristem development (stages 1–6) to the fully developed flower (Stages 7–12) \citep{swanson2005} when it is ready to be pollinated.
-2005).
+In Colombia cocoa trees usually produce flowers throughout the year.
+Cocoa trees produces with up to 10000 flowers per tree each year, which the 50 \%  do not develop into ripe fruits according to Fedecacao reports.
+The flower takes 30 days passing by 12 micro-stages from meristem development (stages 1to 6) to the fully developed flower (Stages 7 to 12) \citep{swanson2005} when it is ready to be pollinated.
 The opening of flowers or anthesis  occurs over a 12-hour period during the night and it is synchronised between the groups of mature flowers \citep{Niemenak2010}.
@@
 Subsequentially, after anthesis the fruit growths by approximately 150 days until the maturation, mucilage.
-Therefore, the complete maturation process of the fruit, from the pollination to fully mature fruit, takes 160–210 days \cite{berry1994}.
-The accumulation of lipids, storage proteins and anthocyanin starts about 85 days after pollination when fruits have an active metabolism and seeds   moisture content decreases up to 30\% \cite{Lehrian1980, Niemenak2010}.
-During this phase the quality of cocoa seeds is defined.\\
+Therefore, the complete maturation process of the fruit, from the pollination to fully mature fruit, takes 160- 210 days \citep{berry1994}.
+The accumulation of lipids, storage proteins and anthocyanin starts about 85 days after pollination when fruits have an active metabolism and seeds   moisture content decreases up to 30\% \citep{Lehrian1980, Niemenak2010}.
+During this phase the quality of cocoa seeds is defined.
 
@@
 	\centering
+	\includegraphics[scale=0.4]{images/phenology.png}\\
 	\caption{\footnotesize {Phenology of cocoa in Colombia for crop modelling.\\}} 
-	\includegraphics[scale=0.4]{images/phenology.png}\\
 	\footnotesize{Credits: Taken from from Dreamstime.com, phys.org \citep{toledo2021} and \cite{lopez2018}}
@@
 
+\subsection{Test site and yield production }
+Cacao is cultivated in 30 states out of 32 from the total Colombian territory with about 147,000 ha \citep{Meza2021}.
+Cocoa fields studied were located in Saravena (Arauca), Rionegro (Santander), Cali (Valle del Cauca), Apartado (Antioquia) and Manizalez (Caldas) (Fig.\ref{fig:yield}, b).
+We considered 112 parcels data from Fedecacao reports.
+Each parcel data contained age of crop, density of planting, yield, number of fruits harvested, flowering date, harvest date.
+Acording to personal comunication from framers the biggest flowering occurs in September and January to harvest in March and July.
+However, the data reported showed that this flowering and pod productions are not constant for all the regions.
+Cali and Caldas presented the lowest production.
+However, the pod harvest in Caldas increased in May and from October to December, and decreased from January to March.
+Meanwhile, Arauca and Apartado reported the highest yield in the months January, July, November and December.
+Santander has picks of production in March, May and September (Fig.\ref{fig:yield}, a).
+
 \begin{figure}[h]
 	\centering
-	\caption{\footnotesize {Average of cocoa production characterization of cocoa for four locations\\}}
-	\includegraphics[scale=0.35]{images/yieldmonth.png}\\
-	{\footnotesize Flowering occurs in September and January to harvest six months latter in March and July respectively.}
+	\includegraphics[scale=0.3]{images/map.png}\\
+	\caption{\footnotesize {Cocoa production characterization of cocoa for five locations\\}}
 	\label{fig:yield}
@@
 
-\subsection{Test Site}
-Cocoa field are located in Saravena (Arauca), Rionegro (Santander), Cali (Valle del Cauca), Apartado (Antioquia) and Manizalez (Caldas).
-We considered 112 parcels data from Fedecacao reports.Each parcel data contained age of crop, density of planting, yield, number of fruits harvested, flowering date, harvest date.
-
-\subsection{Input data acquisition for SIMPLE crop model calibration}
-Input variables required to run SIMPLE model for cocoa include the flowering date and daily weather of solar radiation (SRAD), maximum and minimum temperature (TMAX, TMIN) and rain.
-Weather data as csv file was  downloaded from the POWER Data Access Viewer \citep{nasapower} from January 01 2018 to December 31 2020 for four locations in Colombia (Cali, Rionegro -Santander, Apartado - Antioquia, Saravena Arauca).
-The csv file had to be transformed to .WHT file using R 1.4 version \citep{Rstudio2020}.
-
-\subsection{Climatological Conditions}
-The temperature ranges from 16 to 28 $^\circ$C and it is relatively constant during the year.
-Saravena presented the maximum variability with hotter months during the first half of the year.
-Apartado was found as the hottest with 26 $^\circ$C and Caldas as the coolest site with 18 $^\circ$C compared with other regions tested.
-Cali and Santander have similar temperature but contrasting PAR conditions.
-Santander and Caldas have the biggest variability and the maximum of solar radiation values over 20 MJ m$^{2}$day$^{-1}$.
-In contrast , Cali and Saravena presented the lowest values of PAR below 5 MJ m$^{2}$day$^{-1}$.
-(Fig.\ref{fig:temp}).
-Precipitation in Colombia is presented in two seasons per year from February to April and from October to November, hile the relative humidity remain constant (Fig.\ref{fig:rain}).
-
+\subsection{Weather conditions}
 \begin{figure}[h!]
 	\centering
-	\includegraphics[scale=0.5]{images/clima.png}
-	\caption{\footnotesize {Daily average temperature and available photosynthetic solar radiation (PAR) conditions. \\}}	
+	\includegraphics[scale=0.4]{images/clima.png}
+	\caption{\footnotesize {Colombian weather conditions. \\}}	
 	\label{fig:temp}
+	{\footnotesize (a), available photosynthetic solar radiation (PAR). (b), Daily average temperature  (c), Precipitation and relative humidity per month}
 \end{figure}
+\newpage
 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[scale=0.4]{images/rainmonth.png}
-	\caption{\footnotesize {Precipitation and relative humidity per month in Colombia \\}}
-	\label{fig:rain}
-\end{figure}
+Solar radiation (PAR), temperature, precipitation ant relative humidity was studied from 2018 to 2020 (Fig.\ref{fig:temp}).
+Santander and Caldas had the biggest variability and the maximum of solar radiation values over 20 MJ m$^{2}$day$^{-1}$.
+In contrast , Cali and Arauca presented the lowest values of PAR below 5 MJ m$^{2}$day$^{-1}$.
+(Fig.\ref{fig:temp}, a).
+Even though, Cali and Santander had contrasting PAR conditions, they regions presented  have similar temperature during 2020 (Fig.\ref{fig:temp}, b).
+The temperature ranges from 16 to 28 $^\circ$C and it is relatively constant for each region.
+However, Arauca presented the biggest variability with hotter months during the first half of the year 2019 and 2020.
+Apartado was found as the hottest region studied with 26 $^\circ$C and Caldas as the coldest site with 18 $^\circ$C.
+Precipitation in Colombia is presented in two seasons per year from February to April and from October to November, while the relative humidity remain constant over 80\% (Fig.\ref{fig:temp},c).
+In general, the coldest regions tested (Caldas and Santander) had the maximun values of solar radiation available for photosynthesis.
 
 \subsection{Thermal time for pod harvest date identification}
-Thermal time (degree days) was calculated as cumulative sum of temperature from the flowering date to the pod harvest.
 It is required to model crop growth  considering the cocoa base temperature (T$_{b}$), at which the plant development stops \citep{Ritchie1991}.
-Thus a cocoa T$_{b}$ of 10$^\circ$C was considered according to \cite{lahive2019}.
-This characterization was conducted for the five regions counted 180 days (6 months) and 210 days (7 months just for Cladas) after flowering, as farmers used to harvest by calendar days.
+This characterization was conducted for the five regions counted 180 days (6 months) after flowering, as farmers used to harvest by calendar days.
 Characterizing the thermal time the models predicts the maturation day to harvest the cocoa can vary depending on temperature variations.
+Thermal time (degree days) was calculated  from the flowering date to the pod harvest  using the ecuation \ref{equ:tem}.
+Where tt is the cumulative sum of the daily temperature (T$_{i}$) and T$_{b}$ for cocoa is 10$^\circ$C according to \cite{lahive2019}.
 
-\subsection{SIMPLE crop model calibration}
+\begin{equation}
+Thermal~time~(tt) = \left\lbrace\begin{array}{c} \sum_{i=1}^{n} T_{i} - T_{b} \\
+\vspace{0.2cm}\\ 
+0,\hspace{0.2cm} Flowering~date\end{array}\right.
+\label{equ:tem}
+\end{equation}
 
-The procedure for the SIMPLE model \citep{Zao2019simple}calibration used was a sequential process of modifying or adding the appropriate files, changing the parameters in the simulation management file, running the program and inspecting the results.
-This cycle was then repeated until the  simulated yields for cocoa were reasonably close to the observed yield.
-In the SIMPLE model dummy files are provided for adding new cocoa crop data and weather data, and files 2 to 6 in the list below can be edited to define new cultivars or experiments.
-Then modifying the simulation management file will cause the new files to be read when the program is run.
+\subsection{Model Calibration}
+
+Calibration of crop models are conducted typically for particular cultivars and require site specific inputs for weather \citep{Crout20142}.
+The procedure for the SIMPLE model \citep{Zao2019simple} calibration  was a sequential process of modifying physiological variables specific for cocoa in the inputs files and  adding the appropriate files of weather for each region tested.
+In the SIMPLE model has seven input files where new cocoa crop data should be provided.
+Files in the list below can be edited to define the features of the new cultivars or experiments.
+Cocoa crop has not been simulated with SIMPLE model, hence consider previous cocoa studies, values such as leaf area index (LAI) \citep{Agele2016} and Harvest index (HI) \citep{Quintana2015} where modified.
+Radiation Use Efficiency (RUE) was calibrate according to the perennial crops  (banana and cotton) that had been calibrated previously in SIMPLE model \citep{Zao2019simple} with a RUE of 0.8 and 0.85 respectively.
+As cocoa trees are under shadow the RUE was lower with values between 0.7 and 0.5 g MJ$^{\mathbf{-1}}$ m$^{2}$ (table \ref{tab:reparam}).
+Once the physiological parameters were calibrated to the  simulated yields for cocoa were reasonably close to the observed yield.
+23 flowering dates from 12-July-2019 to 23-June-2020 were introduced in the treatment file, running the program and saving the results.
+
 \begin{enumerate}
@@
 	\item Input/Treatment.csv	
-    \item Input/Irrigation.csv
 	\item Input/Soil.csv
@@
 
-\subsection{Parameters}
+\subsection{Inputs and parameters}
+Input variables required to run SIMPLE model for cocoa include the flowering date and daily weather of solar radiation (SRAD), maximum and minimum temperature (TMAX, TMIN) and rain.
+Weather data as csv file was  downloaded from the POWER Data Access Viewer \citep{nasapower} from January 01 2018 to December 31 2020 for four locations in Colombia (Cali, Rionegro -Santander, Apartado - Antioquia, Saravena Arauca).
+The csv file had to be transformed to .WHT file using R 1.4 version \citep{Rstudio2020}.
+
 There were three parameters varied by region (table \ref{tab:reparam}) thermal time required for pod harvest after the flowering date (Tsum) , the Radiation use efficiency (RUE) and yield observed on field.
@@
 			\hline
-			{\bf Cultivar }&{\bf Tsum }&{\bf RUE}&{\bf Yield$^{*}$}\\
+			{\bf Region }&{\bf Tsum }&{\bf RUE}&{\bf Yield$^{*}$}\\
 			\hline
+			Apartado   & 2906 & 0.6 & 3378  \\
+			Arauca   & 2764 & 0.7 & 3981  \\
 			Santander & 2016 &0.6 & 2687 \\
-			Apartado   & 2906 & 0.6 & 3378  \\
-			Arauca   & 1912 & 0.7 & 3981  \\
-			Cali   & 2764 & 0.5 & 1900  \\
+			Cali   & 1912 & 0.5 & 1900  \\
 			Caldas   & 1192 & 0.6 & 740  \\
@@
 		\end{tabular} \\
-		{\footnotesize $^{*}$ Yield observed kg ha$^{-1}$ per year} 
+		{\footnotesize RUE Radiation use efficiency (above ground only and without respiration)g MJ$^{\mathbf{-1}}$ m$^{\mathbf{2}}$\\$^{*}$ Yield observed kg ha$^{\mathsf{-1}}$ per year. } 
 	\end{small}
@@
 			Treatment&Weather & KOKO (.WTH file name)\\
-			&CO$_{2}$ & 400\\
+			&CO$_{2}$ & 400 ppm\\
 			&SowingDate &Flowering date\\
@@
 \end{table}
+\newpage
 
@@
 
-The SIMPLE model performance was evaluated by comparing simulated values cocoa yield with those reported by Fedecacao from cocoa plantations, using statistical indices of relative-RMSE (RRMSE) (Equ.
-\ref{equ:RMSE}) \citep{Zao2019simple} and coefficient of determination (R$^{2}$) (Equ.
-\ref{equ:R2}) \citep{Zao2019simple, Bai2020, Camargo2019aquacropr}.
-The harvest date predicted was supported by \textcolor{red}{Adriana results}
-
-Yield reports from 20190 to 2020  were shared by Federación Nacional de Cacaoteros Fedecacao.
+The SIMPLE model performance was evaluated by comparing simulated values cocoa yield with those reported by Fedecacao from cocoa plantations, using the statistical indice of relative root mean square error (RRMSE) (Equ.
+\ref{equ:RMSE}) \citep{Zao2019simple, Bai2020}.
 
@@
 
-\begin{equation}
-R^{2}= 1- \frac{\sum_{i=1}^{n} (Y_{i}-X_{i})^{2}}{{\sum_{i=1}^{n}(Y_{i}-0)}^{2}}
-\label{equ:R2}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}
 
-\subsection{Thermal time per region}
-The thermal time characterization showed differences per region as was expected.
-Figure \ref{fig:ttbox} shows degree days ($^\circ$C) of cocoa growth cycles.
-Caldas had the lowest values and Apartado had the the highest results.
-this disttribucion is conincident with the temperature (Fig.\ref{fig:temp}).
-Meanwhile, Cali and Santander presented similar thermal time arround 2016 $^\circ$C.
+\subsection{Thermal time }
+Thermal time characterization is shown in figure \ref{fig:ttbox}.
+It showed differences per region as was expected following the tendencies of the temperature per region (fig.
+\ref{fig:temp},c).
+Apartado and Arauca had the highest temperatures hence, the highest thermal time values with 2909 and 2764 days$^\circ$C  respectively.
+Caldas had the lowest values with 1173 days$^\circ$C.
+Meanwhile, Cali and Santander presented similar thermal time around 2000 $^\circ$C (table \ref{tab:reparam}).
+The accumulated temperature during the pod development (fig.\ref{fig:ttbox}) depends on the region where cocoa is cultivated and the variety planted.
+Thermal time values are also proportional to the yield reported on field (table \ref{tab:reparam}).\\
 
@@
 	\centering
-	\includegraphics[scale=0.25]{images/ttbbox.png}
-	\caption{\footnotesize {Thermal time characterization of cocoa for five locations.\\}} 
+	\includegraphics[scale=0.3]{images/ttbbox.png}\\
+	\caption{\footnotesize {Coaoa yield and thermal time characterization.\\}}
 	\label{fig:ttbox}
@@
 
-\subsection{Weather effects}
-Studying the weather data effects over the flowering date and their final yield.
-In a preliminary analysis of Santanter data, We found that the number of successful flowers pollinated to produce final yield can be affected mostly by the rain, TMAX and wind (Figure \ref{fig:heat}).
-Farmers
+\subsection{Weather effects over flowering time }
+
+Figure \ref{fig:heat} shows the pearson correlation to study the weather data of the flowering time over the months of flowering  (monthF), month of harvest (monthH) and their final yield (fruit\_kg) .
+The results showed that thermal time with T$_{b}$ (ttb), temperature minimum and maximum (TMIN and TMAX) and Dew Frost Point at 2 meters (T2MDEW) are positive related.
+The wind (WS2M) is clear correlated with ttb.
+However, less clear correlations were found of monthF  with T2MDEW, relative humidity (RH), WS2M and rain.
 
@@
 	\centering
-	\includegraphics[scale=0.35]{images/outmodel.png}
+	\includegraphics[scale=0.35]{images/heatm.png}\\
+	\caption{\footnotesize {Pearson correlation among weather variables and flowering dates all locations. \\}}
+	\label{fig:heat} 
+\end{figure}
+
+\subsection{Model validation}
+The cocoa yield simulation (kg ha$^{-1}$ per year) was validated using observed data (table \ref{tab:reparam}) showing a final RRMSE of 7.2\% (fig.
+\ref{fig:m1}, a).
+Individual errors per region are presented in table \ref{tab:error}, where the best fit simulation model was for yield prediction in Apartado and the highest error was calculated for Caldas crops.
+The model responded to the variations of temperature and solar radiation.
+Therefore, the highest  yield values simulated were obtained for Arauca over 4000 kg ha$^{-1}$, followed by Apartado Santander with yields over 2000 kg ha$^{-1}$.
+The lowest yield was simulated for Caldas region with less of 1000 kg ha$^{-1}$ .
+Final yield in the model was is calculated as the product of biomass of aerial part and harvest index (HI) \citep{Zao2019simple, Amir1991}, where the HI is similar to the CropSyst \citep{STOCKLE2003} and AquaCrop \cite{Steduto2009}.
+
+Therefore, biomass production of the aerial part of the plant was simulated per region (fig.
+\ref{fig:m1}, b) as biomass-rate which is the daily biomass growth rate.
+As for cocoa simulation we did not consider water and heat stress variables, in this research biomass simulation is affected by the solar radiation, RUE, daily temperature, atmospheric CO$_{2}$ concentration (ppm) and the fraction of solar radiation intercepted by a tree of cocoa during the fruit development (fSolar) (fig.
+\ref{fig:m1}, c).
+Due to the lack of this kind of data from fields, the comparison was not possible with observed data of biomass production through the crop cycle.
+However, the biomass simulation also responded proportionally with the yield production.
+Biomass of cocoa fruits simulated include all the parts that compound  the pod when they are fresh.
+Therefore, these this information may be of interest to farmers to predict pod fresh mass.
+
+The figure \ref{fig:m1}, c the fSolar was simulated thought the fruit development resulting that crops in  all regions reached the maximun fSolar at 0.94, except Caldas where crops reached 0.76.
+Apartado and Arauca reached faster this fSolar-max.
+These high values of solar radiation intercepted for photosynthesis, lasted differently depending on the region and their solar radiation (Fig.\ref{fig:temp}, a) : Apartado 66 days  from  69 to 135 days after flowering (daf),  Arauca 61 days from 72 to 133 daf, Cali 38 days from 83 to 121 daf,  Santander 24 days from 92 to 116 daf and Caldas 2 days at 125 daf.
+fSolar declined in the interception of solar radiation until the pod harvest day.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[scale=0.4]{images/outmodel.png}
 	\caption{\footnotesize {SIMPLE model output for cocoa in Colombia.\\ }} 
@@
 
-\subsection{Day of harvest prediction}
+\begin{table}[h!]	
+	\caption {\footnotesize {Summary of relative root mean square error RMSE for the predicted cocoa yield using SIMPLE model.}}
+	\label{tab:error} 
+	\centering
+	\begin{small}
+		\begin{tabular}{l c c c c c c}
+			\hline
+			{\bf Region }&{\bf Apartado }&{\bf Arauca}&{\bf Santander}&{\bf Cali}&{\bf Caldas}&{\bf Overal}\\
+			\hline
+			RMMSE \%  & 3 & 6.05 & 10.06&8.5&14.90&7.2 \\
+			\hline
+		\end{tabular} \\
+	\end{small}
+\end{table}
+
+\subsection{Prediction of the pod harvest day }
+These result demonstrated that the traditional way to harvest always at 180 daf is not having account physiological and environmental conditions that can be affecting the pod development and maturation.
+Figure \ref{fig:dayH} present that observe day of harvest or maturity day is 180 daf independently of the region.
+All the regions except Santander, presented the earliest harvest when the flowering date was between December and February.
+The fruit can be ready to harvest depending of the enviromental condition per region.
+Thus, when the fruit development was simulated the maturity day in Cali and Caldas were from 3 to 12 and 4 to 23 days before 180 daf (Fig.
+\ref{fig:dayH}, c and f respectively).
+Apartado presented the most similar predicted dates of harvest to 180 daf with 170 to 182 daf  (Fig.
+\ref{fig:dayH}, d).
+The pod may be harvest in Arauca (Fig.\ref{fig:dayH}, b)  between 165 and 193 daf, Santander (Fig.\ref{fig:dayH}, a) between 165 and 183 daf, ten days less than in Arauca for the same months of flowering of July and August of 2019 .
+
 \begin{figure}[h!]
@@
 	\label{fig:dayH}
-\end{figure}
-%\begin{figure}[h!]
-%	\centering
-%	\caption{\footnotesize {Weather variables during flowering times.\\}} 
-%	\includegraphics[scale=0.4]{images/averageS.png}
-%	\label{fig:weather}
-%\end{figure}
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[scale=0.4]{images/heatm.png}\\
-	\label{fig:heat}
-	\caption{\footnotesize {Pearson correlation among weather variables and flowering dates all locations. \\}} 
-	{\footnotesize thermalT= thermal time without Tb. ttb= thermal time with Tb. WS2M= Wind Speed at 2 Meters (m/s). T2MDEW = Dew/Frost Point at 2 Meters ($^\circ$C) }
-\end{figure}
+\end{figure}.
 \newpage
@@
 Future research directions may also be highlighted.
+
+This study presents a new approach of SIMPLE model calibration for cocoa as tropical crop in South America for five environments.
 
@@
 Unfertilised flowers abscise from the trunk approximately 1 day after flower opening  \citep{Niemenak2010}.
+In a preliminary analysis of Santanter data, We found that the number of successful flowers pollinated to produce final yield can be affected mostly by the rain, TMAX and wind (fig.\ref{fig:heat}).
+
+\subsection{App development and future challenges }
+The original code of SIMPLE model of \citep{Zao2019simple} was modified to make easy the implementation of this cocoa model as and app to be used in smartphones and desktops by farmers in Colombia.
+Therefore, the new version for cocoa crops simulation will be used to predict yield, date of harvest and biomass production, inserting only the date of flowering and region.
+The app development is on charge of Grupo BIOS to be deliver to farmers in Caldas initially at the end of 2021.
+This research presented and initial crop calibration that can be improved with further studies, including effects over the pod production by diseases, nutritional diffidences and abiotic stresses.
+The future challenge will be that traditional farmers start to harvest more aware of the environmental effects over their crops.
+It will be necessary that they engage growers with adapt founding from scientific studies.
+Moreover, It will be required the help of entrepreneurs, researchers, academics and non-specialized communities to transfers the knowledge to cocoa growers.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusions}
+Based on principles of crop physiology this  generic crop model , which was developed with relatively few equations and parameters was fitted for cocoa fruit growth.
+The SIMPLE model worked well for cocoa yield prediction and to demonstrate that the moment to harvest the cocoa pod is affected by the weather variations over the physiology of the trees.
 
-This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}
-
-% Only for the journal Methods and Protocols:
-% If you wish to submit a video article, please do so with any other supplementary material.
-% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``Conceptualization, X.X.
-and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
-and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
-All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
-Authorship must be limited to those who have contributed substantially to the work~reported.}
-
-\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
-Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
-
-\dataavailability{In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study.
-Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.
-You might choose to exclude this statement if the study did not report any data.}
-
-\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
+\acknowledgments{The authors wish to thank to Grupo BIOS and Fedecacao for facilitated the data from cocoa fields.
+In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
 This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
@@
 \conflictsofinterest{The authors declare no conflict of interest.}
-
-%% Optional
 
@@
 \bibliography{kokobib}
-
-%=====================================
-% References, variant B: internal bibliography
-%=====================================
-%\begin{thebibliography}{999}
-%% Reference 1
-%\bibitem[Author1(year)]{ref-journal}
-%Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
-%% Reference 2
-%\bibitem[Author2(year)]{ref-book1}
-%Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
-%% Reference 3
-%\bibitem[Author3(year)]{ref-book2}
-%Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
-%% Reference 4
-%\bibitem[Author4(year)]{ref-unpublish}
-%Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} stage of publication (under review; accepted; in~press).
-%% Reference 5
-%\bibitem[Author5(year)]{ref-communication}
-%Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
-%% Reference 6
-%\bibitem[Author6(year)]{ref-proceeding}
-%Author 1, A.B.; Author 2, C.D.; Author 3, E.F.
-Title of Presentation.
-In Title of the Collected Work (if available), Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference; Editor 1, Editor 2, Eds.
-(if available); Publisher: City, Country, Year (if available); Abstract Number (optional), Pagination (optional).
-%% Reference 7
-%\bibitem[Author7(year)]{ref-thesis}
-%Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
-%% Reference 8
-%\bibitem[Author8(year)]{ref-url}
-%Title of Site. Available online: URL (accessed on Day Month Year).
-%\end{thebibliography}
-
-%\bibliographystyle{te}
-%\bibliography{bioinfo}
-
-% If authors have biography, please use the format below
-%\section*{Short Biography of Authors}
-%\bio
-%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
-%{\textbf{Firstname Lastname} Biography of first author}
-%
-%\bio
-%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
-%{\textbf{Firstname Lastname} Biography of second author}
-
-% The following MDPI journals use author-date citation: Admsci,  Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Jintelligence, JRFM, Languages, Laws, Literature, Religions, Risks, Social Sciences.
-For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328; \citeyear{ref-journal-3b}, p.475).
-This produces: Wong (1999, p.
-328; 2000, p.
-475)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% for journal Sci
-%\reviewreports{\\
-%Reviewer 1 comments and authors’ response\\
-%Reviewer 2 comments and authors’ response\\
-%Reviewer 3 comments and authors’ response
-%}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \end{document}

### 2021-11-05 paperModel/paperModel.tex
Add files via upload

--- 2021-11-03 papers/model/paperModel.tex
@@
 % Authors, for the paper (add full first names)
-\Author{Angela Romero V $^{1,\ddagger}$*\orcidA{}, Adriana Gallego $^{2}$ and Anyela V. Camargo R.$^{1, \ddagger}$}
+\Author{Angela Romero V $^{1,\ddagger}$*\orcidA{}, Anyela V. Camargo R. $^{1, \ddagger}$ and Adriana M. Gallego$^{2}$}
 
@@
 ) is an important worldwide perennial tropical crop endemic to the South American rainforests \citep{zuidema2005, motamayor2002, argout2011, Rodriguez2019}.
-Cacao plant member of the Malvaceae (formerly Sterculiaceae)  botanical family such as  cotton \textit{ Gossypium hirstium} \citep{Nix2017cotton} wich is modeled in SIMPLE model \citep{Zao2019simple}.
-Cocoa is grown for its fruits, known as cacao pods.
-\citep{ Niemenak2010, suarez2021}.
-Only the 5\% of the world cocoa yield is desalinated for Fine-cocoa production due to the low productivity of  the traditional crop management \citep{argout2011}.
+Cacao plant is a member of the Malvaceae (formerly Sterculiaceae)  botanical family such as  cotton \textit{ Gossypium hirstium} \citep{Nix2017cotton}.
+Coton has been is modeled in SIMPLE model \citep{Zao2019simple}.
+Cocoa is grown for its fruits, known as cacao pods \citep{ Niemenak2010, suarez2021}.
+Only the 5\% of the world cocoa yield is destined for Fine-cocoa production due to the low productivity through  the traditional crop management \citep{argout2011}.
 In Colombia, cocoa  is  traditionally  consumed  as  a  beverage.
@@
 Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}, it is still not widely produced as the lack of adoptions of technologies by the traditional farmers.
-They empirically harvest after 5 or 6 months after flowering date, hence they ferment cocoa beans without considering the  quality of the seeds at the harvest time.
+They empirically harvest at 5 or 6 months after flowering date, hence   a mix of quality of cocoa beans can be fermented.
 This produce heterogeneous characteristics between each fermentation batch diminishing the quality of cocoa final product \citep{Escobar2021}.
@@
 	\caption{\footnotesize {Phenology of cocoa in Colombia for crop modelling.\\}} 
-	\footnotesize{Credits: Taken from from Dreamstime.com, phys.org \citep{toledo2021} and \cite{lopez2018}}
+	\footnotesize{Source: Taken from from Dreamstime.com, phys.org \citep{toledo2021} and \cite{lopez2018}}
 	\label{fig:pheno}
@@
 	\label{fig:temp}
-	{\footnotesize (a), available photosynthetic solar radiation (PAR). (b), Daily average temperature  (c), Precipitation and relative humidity per month}
+	{\footnotesize (a), Available photosynthetic solar radiation (PAR).
+(b), Daily average temperature  (c), Monthly average of precipitation (bars) and relative humidity (dotted line) from 2018 to 2021 }
 \end{figure}
@@
 
-Calibration of crop models are conducted typically for particular cultivars and require site specific inputs for weather \citep{Crout20142}.
+Calibration of crop models are conducted typically for particular cultivars and require site specific inputs of weather \citep{Crout20142}.
 The procedure for the SIMPLE model \citep{Zao2019simple} calibration  was a sequential process of modifying physiological variables specific for cocoa in the inputs files and  adding the appropriate files of weather for each region tested.
@@
 
+\subsection{Weather effects over flowering time }
+
+Figure \ref{fig:heat} shows the pearson correlation to study the weather data of the flowering time over the months of flowering  (monthF), month of harvest (monthH) and their final yield (fruit\_kg) .
+The results showed that thermal time with T$_{b}$ (ttb), temperature minimum and maximum (TMIN and TMAX) and Dew Frost Point at 2 meters (T2MDEW) are positive related.
+The wind (WS2M) is clear correlated with ttb.
+However, less clear correlations were found of monthF  with T2MDEW, relative humidity (RH), WS2M and rain.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[scale=0.35]{images/heatm.png}\\
+	\caption{\footnotesize {Pearson correlation average weather variables and flowering dates for all regions. \\}}
+	\label{fig:heat} 
+\end{figure}
+\newpage
+
 \subsection{Thermal time }
-Thermal time characterization is shown in figure \ref{fig:ttbox}.
-It showed differences per region as was expected following the tendencies of the temperature per region (fig.
+Thermal time characterization  was made considering 180 days after flowering (daf) for each region.
+The boxplot in the figure \ref{fig:ttbox} shows the data distribution where boxes indicate the range of the central 50\% of the total data per region, the central line in the box is marking the median value and lines draw out from each box mean the range of the remaining data.
+Therefore, this boxplot shows differences between locations as was expected following the tendencies of the temperature  (fig.
 \ref{fig:temp},c).
@@
 	\includegraphics[scale=0.3]{images/ttbbox.png}\\
-	\caption{\footnotesize {Coaoa yield and thermal time characterization.\\}}
+	\caption{\footnotesize {Cocoa yield and thermal time characterization at 180 days after flowering.\\}}
 	\label{fig:ttbox}
@@
 
-\subsection{Weather effects over flowering time }
+\subsection{Model validation}
+Therefore, biomass production of the aerial part of the plant was simulated per region (fig.
+\ref{fig:m1}, a) as biomass-rate which is the daily biomass growth rate.
+As for cocoa simulation we did not consider water and heat stress variables, in this research biomass simulation is affected by the solar radiation, RUE, daily temperature, atmospheric CO$_{2}$ concentration (ppm) and the fraction of solar radiation intercepted by a tree of cocoa during the fruit development (fSolar) (fig.
+\ref{fig:m1}, b).
+Due to the lack of this kind of data from fields, the comparison was not possible with observed data of biomass production through the crop cycle.
+However, the biomass simulation also responded proportionally with the yield production.
+Biomass of cocoa fruits simulated include all the parts that compound  the pod when they are fresh.
+Therefore, these this information may be of interest to farmers to predict pod fresh mass.
 
-Figure \ref{fig:heat} shows the pearson correlation to study the weather data of the flowering time over the months of flowering  (monthF), month of harvest (monthH) and their final yield (fruit\_kg) .
-The results showed that thermal time with T$_{b}$ (ttb), temperature minimum and maximum (TMIN and TMAX) and Dew Frost Point at 2 meters (T2MDEW) are positive related.
-The wind (WS2M) is clear correlated with ttb.
-However, less clear correlations were found of monthF  with T2MDEW, relative humidity (RH), WS2M and rain.
+The figure \ref{fig:m1}, b the fSolar was simulated thought the fruit development resulting that crops in  all regions reached the maximun fSolar at 0.94, except Caldas where crops reached 0.76.
+Apartado and Arauca reached faster this fSolar-max.
+These high values of solar radiation intercepted for photosynthesis, lasted differently depending on the region and their solar radiation (Fig.\ref{fig:temp}, a) : Apartado 66 days  from  69 to 135 days after flowering (daf),  Arauca 61 days from 72 to 133 daf, Cali 38 days from 83 to 121 daf,  Santander 24 days from 92 to 116 daf and Caldas 2 days at 125 daf.
+fSolar declined in the interception of solar radiation until the pod harvest day.
 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[scale=0.35]{images/heatm.png}\\
-	\caption{\footnotesize {Pearson correlation among weather variables and flowering dates all locations. \\}}
-	\label{fig:heat} 
-\end{figure}
-
-\subsection{Model validation}
 The cocoa yield simulation (kg ha$^{-1}$ per year) was validated using observed data (table \ref{tab:reparam}) showing a final RRMSE of 7.2\% (fig.
-\ref{fig:m1}, a).
-Individual errors per region are presented in table \ref{tab:error}, where the best fit simulation model was for yield prediction in Apartado and the highest error was calculated for Caldas crops.
+\ref{fig:m1}, c).
+Individual errors per region are presented in table \ref{tab:error}, where the best fit of the calibration model  for yield prediction  was for crops in  Apartado and the highest error was calculated for Caldas crops.
 The model responded to the variations of temperature and solar radiation.
@@
 
-Therefore, biomass production of the aerial part of the plant was simulated per region (fig.
-\ref{fig:m1}, b) as biomass-rate which is the daily biomass growth rate.
-As for cocoa simulation we did not consider water and heat stress variables, in this research biomass simulation is affected by the solar radiation, RUE, daily temperature, atmospheric CO$_{2}$ concentration (ppm) and the fraction of solar radiation intercepted by a tree of cocoa during the fruit development (fSolar) (fig.
-\ref{fig:m1}, c).
-Due to the lack of this kind of data from fields, the comparison was not possible with observed data of biomass production through the crop cycle.
-However, the biomass simulation also responded proportionally with the yield production.
-Biomass of cocoa fruits simulated include all the parts that compound  the pod when they are fresh.
-Therefore, these this information may be of interest to farmers to predict pod fresh mass.
-
-The figure \ref{fig:m1}, c the fSolar was simulated thought the fruit development resulting that crops in  all regions reached the maximun fSolar at 0.94, except Caldas where crops reached 0.76.
-Apartado and Arauca reached faster this fSolar-max.
-These high values of solar radiation intercepted for photosynthesis, lasted differently depending on the region and their solar radiation (Fig.\ref{fig:temp}, a) : Apartado 66 days  from  69 to 135 days after flowering (daf),  Arauca 61 days from 72 to 133 daf, Cali 38 days from 83 to 121 daf,  Santander 24 days from 92 to 116 daf and Caldas 2 days at 125 daf.
-fSolar declined in the interception of solar radiation until the pod harvest day.
-
 \begin{figure}[h!]
@@
 	\includegraphics[scale=0.4]{images/outmodel.png}
-	\caption{\footnotesize {SIMPLE model output for cocoa in Colombia.\\ }} 
+	\caption{\footnotesize {Model predictions.
+(a) Biomass aerial part.
+(b) Interception of solar radiation.
+(c) Yield.
+Crop cycle close to 180 daf (vertical red line) base on figure  \ref{fig:ttbox}.
+\\ }}
 	\label{fig:m1}
@@
 The pod may be harvest in Arauca (Fig.\ref{fig:dayH}, b)  between 165 and 193 daf, Santander (Fig.\ref{fig:dayH}, a) between 165 and 183 daf, ten days less than in Arauca for the same months of flowering of July and August of 2019 .
+Only Arauca presented longer crop cycles when the flowering was between during that period of time.
+This means that Arauca had bigger variation of temperature between months (Fig.
+\ref{fig:temp}, b ).
 
@@
 	\includegraphics[scale=0.4]{images/RegionHarvest2.png}
-	\caption{\footnotesize {Day of harvest cocoa simulation\\ }} 
+	\caption{\footnotesize {Harvest day prediction from flowering date for (a) Santander, (b) Arauca, (c) Cali, (d) Apartado,(f) Caldas. \\ }} 
 	\label{fig:dayH}
@@
 \section{Discussion}
-Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses.
-The findings and their implications should be discussed in the broadest context possible.
-Future research directions may also be highlighted.
 
+\subsection{Weather Effects over Flower Stability and Pollination}
 This study presents a new approach of SIMPLE model calibration for cocoa as tropical crop in South America for five environments.
+The analysis of the weather conditions (Fig.
+\ref{fig:heat}) did not confirmed clearly that flower are affected by the wind and rain by mechanical damage.
+However, farmers stated that the number of flowers pollinated decrease by months where wind and rain are high.
+Moreover, wind can affect the availability of tiny flies pollinators from Diptera order and from the families of of the biting midges \textit{Ceratopogonidae},  genus  \textit{Forcipomyia} \citep{Saunders1959, kaufmann1975, sotomayor2020} to reach the cocoa flowers.
+However, the stability of cocoa flowers is influenced by seasonal wheather conditions (abiotic) and pollination (biotic) \citep{Frimpong2014}.
+Therefore, pollinator population shoud be coincidence with the phenology of the flowering cocoa trees \citep{Young1983, Young2012}.
+Flower opening is very well synchronised between the cohorts of mature flowers opening each night \citep{Niemenak2010}.
+The flowers open at almost exactly the same time and rate, irrespective of their position on the trunk.
+Thus, unfertilised flowers abscise from the trunk approximately 1 day after flower opening  \citep{Niemenak2010}.
+Hence more than 90\% of unpollinated flowers fall or abscised within 32 hours after  anthesis \citep{Aneja1999}.
+Abscission processed of  flowers are mainly controlled  by  three  hormones: auxin, ethylene, and abscisic acid (ABA) \citep{Aneja1999}.
+Ethylene generally promotes abscission because it may  inhibit the transport of auxin from the leaf blade, which allow the action of ABA to promove the fall of flowers \citep{Beyer1975}.
+In general, environmental conditions can also  stimulate   to  a  decrease  the  auxin/ethylene relationship \citep{Aneja1999}.
 
-The optimum temperature for photosynthesis in cacao has
-been reported to be between 31 °C and 33 °C (Balasimha et al.
-1991) and 33 °C–35 °C (Yapp 1992). Sena Gomes and
-Kozlowski (1987) reported a decrease in gs over the range
-18.7 °C to 27.2 °C and an increase at temperatures above this,
-most likely to enhance cooling at higher temperatures.
-However, the opposite effect was described by Raja Harun
-and Hardwick (1988a). In this study, gs and transpiration (E)
-increased within the range of 20 °C–30 °C. In field-grown
-cacao in India, photosynthetic rate declined as mean monthly
-temperature increased above 34 °C during the dry season
-(Balasimha et al. 1991). \cite{lahive2019}
-\textcolor{red}{this section on going to organize}
+In our preliminary analysis, Santanter data, showed that the number of successful flowers pollinated to produce final yield can be affected negatively mostly by the rain, TMAX and wind.
+Nevertheless, a better field data tracing flowers development is essential to understand if it is a mechanical or physiological effect.
+In contrast,  \cite{Frimpong2009, Frimpong2011} indicated that the numbers of cocoa pollinators were reduced during the dry season, but increased in the wet season.
+This, could be due to midges need a moist environment to develop \citep{Frimpong2014}, which is difficult during the dry season as cocoa leaves create a dried ground mat \citep{Frimpong2009, Frimpong2014}.
+Moreover, the lack of water during dry seasons may reduce the nutrients uptake provoking the massive flower drops \citep{Vaughton2017}.
+For future studies, wind could be included in the SIMPLE model as an input to simulated this mechanical effects over number of flowers.
+In general, field data regarding counting flowers pollinated by month should be better reported for the region here studied.
 
-Evaluating the level of knowledge of producers regarding cocoa crop management, the harvest was in the group of activities that presented the lowest level of knowledge on the part of the producers according to the general averages \citep{Gutierrez2020}.
+\subsection{Thermal Time for Harvest day predictions}
+We define the thermal time required to harvest cocoa pods for five Colombian regions as maturation of the fruit is related with temperature during the growth cycle \citep{lopez2018}.
+Previous studies have calculated thermal time for  different cocoa cultivar in Brazil and Ghana \cite{Daymond2008}.
+They also confirm that the fruit maturation time decrease in with an increase in temperature as was presented on others researches \citep{Alvim1974, End1991, Daymond2008}.
+the effects of temperature and solar radiation on fruit growth and development was previously studied by \cite{Daymond2008}, showing that crop under higher temperatures thought the crop cycle induce greater fruit losses because of physiological maturation (cherelle wilt).
+When the fruit is mature, seeds are able to germinate inside the pod before the harvest day \citep{lopez2018}.
 
-\subsection{Weather Effects}
-Wind can affect the availability of tiny flies pollinators from Diptera order and from the families of of the biting midges \textit{Ceratopogonidae},  genus  \textit{Forcipomyia} \citep{Saunders1959, kaufmann1975, sotomayor2020} to reach the cocoa flowers.
-However, wind is not included in the SIMPLE model as an input as cocoa crop has not been simulated with this model before.
-Flower opening is very well synchronised between the cohorts of mature flowers opening each night.
-The flowers open at almost exactly the same time and rate, irrespective of their position on the trunk.
-Unfertilised flowers abscise from the trunk approximately 1 day after flower opening  \citep{Niemenak2010}.
-In a preliminary analysis of Santanter data, We found that the number of successful flowers pollinated to produce final yield can be affected mostly by the rain, TMAX and wind (fig.\ref{fig:heat}).
+Our results showed that the hotter regions such as Apartado and Arauca presented higher thermal time values (Fig.
+\ref{fig:ttbox} and \ref{fig:temp}).
+Even though, Arauca had very low values of SAR but very high temperatures, this may be caused by clouds cover.
+The opposite can be seen for Caldas and Santander.
+These, extreme relations T/SAR can compensate the crop efficiency, for example in Santander (Fig.
+\ref{fig:m1}).
+The thermal time calculation was defined base on  180 daf because farmers cut the pods by calendar days.
+Previous studies stated that evaluating the level of knowledge of growers regarding cocoa crop management, showed that the harvest was in the group of activities that presented the lowest level of information by the farmers \citep{Gutierrez2020}.
+That is why, these results present important temperature boundaries to predict fruit maturation day.
+Therefore, may be other environmental factors that should be studied for further research.
+
+\subsection{Cocoa Crop Model Simulations}
+Although cocoa is a relevant crop and there is an extensive agronomic literature, there is only one physiological crop model specific for cocoa so far.
+The (SUCROS-Cocoa) developed by \cite{zuidema2005}.
+However, the code was not easy available for adaptations.
+In contrast, the SIMPLE model has an open code in R, which we could adapt such a model would be very useful to compare yields and predict harvest date in different climates.
+As the harvest day was predicted form the flowering date (Fig.
+\ref{fig:dayH}, consequently, biomass production and fSolar presented a crop cycle shorter than 180 daf (Fig.
+\ref{fig:m1}, a and b).
+These simulations are coincident with results presented by \cite{lopez2018}, where physiological maturity of coca pod varies from 140 to 162 daf.
+Our results showed how the harvest day can vary depending on the accumulate temperature during each especific crop cycle simulated (Fig.
+\ref{fig:dayH}).
+
+Biomass simulations use SIMPLE model presented similar predicted values (10000 Kg ha$^{-1}$) for coca drops in Costa Rica using SUCROScocoa \citep{zuidema2005}.
+Biomass simulations are a common evaluation in crop modells such as Sirius \citep{Crout20142}, SUBSTOR-potato \citep{Raymundo2017} and  DSSAT, CropSyst, STICS and WOFOST \citep{Confalonieri2016}.
+The approach of this research was focus  on the harvest date prediction, hence the leaves crop cycle was evaluated indirectly this study.
+The fraction of intercepted photosynthesis active radiation (fSolar) decreased (Fig.
+\ref{fig:m1}, b) when the senescence of the canopy \citep{zuidema2005}.
+In cocoa canopy senecence  refereed to  a group of leaves responsible  at the moment of the fruit formation to produce carbohydrates.
+These leaves eventually drop becoming on litter over the soil.
+Leaves life cycle has been simulated using crop models \citep{Crout2010, zuidema2005}, which can be the reference to improve our SIMPLE coca crop model in future studies.
+
+Yield prediction presented a RRMSE values 7.2\%, which were significant lower than those  presented for other crops using the SIMPLE model which reported and RRMSE of 24.4\% \citep{Zao2019simple}.
+Resulting in reliable approach for cocoa yield prediction.
+In general, these results may help to improve the quality of cocoa seed considering the moment to harvest can be variable depending on weather changes.
 
@@
 \section{Conclusions}
-Based on principles of crop physiology this  generic crop model , which was developed with relatively few equations and parameters was fitted for cocoa fruit growth.
-The SIMPLE model worked well for cocoa yield prediction and to demonstrate that the moment to harvest the cocoa pod is affected by the weather variations over the physiology of the trees.
+Cocoa fruit development for harvest in the right time depend on whether conditions and principles of crop physiology and flower phenology.
+This was common for the five regions.Thermal time characterization range from 1200 to 3000 days $^\circ$C, with a T$_{b}$ of 10 $^\circ$C for the fruit development.
+The SIMPLE model as a generic crop model allowed to predict the harvest date with better precision than only considering days by calendar.
+Thus, the crop cycle of cocoa for harvest  should be shorter than 180 days after flowering.
+These results confirm the potential of crop model for tropical crop in Latin America.
 
-\acknowledgments{The authors wish to thank to Grupo BIOS and Fedecacao for facilitated the data from cocoa fields.
-In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
-This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
+\section{Author contributions}
+ARV and ARC model calibration and data analsis. AMG field data collection. ARV were primarily responsible for writing the manuscript.
+
+\acknowledgments{The authors wish to thank to Grupo BIOS and Fedecacao for facilitated the data from cocoa fields. }
 

### 2021-11-14 paperModel/paperModel_2.tex
Corrections

--- 2021-11-05 paperModel/paperModel.tex
@@
 \usepackage[colorinlistoftodos]{todonotes}
+\usepackage[dvipsnames]{xcolor}
 \newcounter{mycomment}
@@
 %=================================================================
-%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
+%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, characterisation, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
 %% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
@@
 % Authors, for the paper (add full first names)
-\Author{Angela Romero V $^{1,\ddagger}$*\orcidA{}, Anyela V. Camargo R. $^{1, \ddagger}$ and Adriana M. Gallego$^{2}$}
+\Author{Angela P. Romero V. $^{1,\ddagger}$*\orcidA{}, Anyela V. Camargo R.$^{1, \ddagger}$, Oscar D. Ramirez$^{3}$, Paula A. Arenas V$^{3}$ and Adriana M. Gallego$^{2}$}
 
@@
 % MDPI internal command: Authors, for citation in the left column
-\AuthorCitation{Romero, A.; Gallego, A.; Camargo, A.}
+\AuthorCitation{Romero, A.P.V;  Camargo, A.; Arenas P.; Ramirez O. D.; Gallego, A.;}
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
@@
 % Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address{$^{1}$ \quad The John Bingham Laboratory, NIAB, 93 Lawrence Weaver Road, Cambridge CB3 0LE, UK 1\\$^{2}$ \quad Grupo BIOS, Centro de Bioinformática y Biología Computacional de Colombia - BIOS, Manizales, Caldas, Colombia, South America.
+\address{$^{1}$ \quad The John Bingham Laboratory, NIAB, 93 Lawrence Weaver Road, Cambridge CB3 0LE, UK 1\\$^{2}$ \quad Grupo BIOS, Centro de Bioinformática y Biología Computacional de Colombia - BIOS, Manizales, Caldas, Colombia, South America.\\ $^{3}$ \quad Federación Nacional de Cafeteros FEDECACAO, Colombia, South America.
 }%; e-mail@e-mail.com}
@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{Crop modelling can support agronomical decisions of crop production under a range of scenarios improving competitiveness.
-Cocoa production systems in Latin America has a high importance over social and economic development, facing  the fight against hunger and poverty.
-Although Colombian cocoa has the potential to be in the high value markets for fine flavour, it is still not widely produced as the lack of adoptions of technologies by the traditional farmers.
-They empirically harvest after 5 or 6 months after flowering date.
-However, cocoa fruits development can be considered as the result of a number of physiological and morphological processes that can be described by mathematical relationships even under uncontrolled environments.
-Thus, we parametrized the SIMPLE crop model \citep{Zao2019simple} to predict the best time for harvest cocoa fruits in Colombia.
-The results showed an RRMSE of 7.2\% for the yield prediction, while the simulated harvest date varied between +/- 2 to 20 days depending on the temperature variations of the year between regions.
-This crop model application contributed to understand and predict the phenology of cacao fruits the varieties ICS95 y CCN51, which is key to produce high quality cacao beans.
-The aim of this study was developed a practical tool for farmers on cocoa fields to predict the best moment to harvest, using easily available data for crop modelling such as flowering date and weather variables (solar radiation, rain, and temperature).
+\abstract{Cacao production systems in Latin America {\color{red}have } a high importance over social and economic development, facing  the fight against hunger and poverty.
+Although Colombian cacao has the potential to be in the high value markets for fine flavour, {\color{red} the lack of expert support as well as the use of traditional, and oftentimes suboptimal, technologies makes cocoa production negligibly.
+For example, traditionally cacao harvest takes place at 5 or 6 months after flowering, other environmental parameters that have more association with pod maturation speed are not taken into account.
+Cocoa fruits development can be considered as the result of a number of physiological and morphological processes that can be described by mathematical relationships even under uncontrolled environments.
+In this context, crop models are useful tools to simulate and predict crop development over time and under multiple environmental conditions.
+Since, harvesting at the right time can yield high quality cacao, we parametrised a crop model to predict the best time for harvest cocoa fruits in Colombia.
+Our aim was to develop a practical tool that supports cacao farmers in the production of high quality cacao.} When comparing simulated and observed data, our results showed an RRMSE of 7.2\% for the yield prediction, while the simulated harvest date varied between +/- 2 to 20 days depending on the temperature variations of the year between regions.
+{\color{red} This crop model contributed to understand and predict the phenology of cacao fruits for two key cacao varieties ICS95 y CCN51.}
 }
@@
 \section{Introduction}
+
 Cocoa ( \textit{Theobroma cacao }L.
@@
 
-Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}, it is still not widely produced as the lack of adoptions of technologies by the traditional farmers.
-They empirically harvest at 5 or 6 months after flowering date, hence   a mix of quality of cocoa beans can be fermented.
-This produce heterogeneous characteristics between each fermentation batch diminishing the quality of cocoa final product \citep{Escobar2021}.
+Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}, {\color{red} the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly.
+For example, the farmer practice of empirically harvesting from 5 to 6 months or {\color{olive} 180 days } after flowering date (DAF), produced a mix of quality of cocoa beans that are then be fermented.
+This practice produces heterogeneous characteristics between each fermentation batch which potentially diminish the quality of the end cocoa product} \citep{Escobar2021}.
 To identify the best moment to harvest is important to consider physiological responses affected by climate variables such as rain, solar radiation and wind.
@@
 
-In this paper, we present a physiological parametrization of SIMPLE crop model for cocoa to predict the best harvest time  and yield production.
-We used the SIMPLE crop model \citep{Zao2019simple} for three reasons: 1: That it is very comprehensively described in the original paper.
-2: That the code was available in R for initial trials and 3: That it had already been successfully fitted to other perennial crops in south America.
-Overall, the model simulates crop development, growth and yield, and predict the maturation day when the fruit is ready to harvest.
-It includes 13 parameters ( daily weather data, irrigation, and soil and key dates) to specify a crop type, with four of these for cultivar characteristics easily available from farmer.
-Thus, this could be used as  tool for small farmers with the aim to improve the quality of cocoa to become more competitive in Fine-cocoa market.
+In this paper, we present a physiological parametrisation of SIMPLE crop model for cocoa to predict best harvest time and overall yield.
+{\color{red} We parameterised the SIMPLE crop model \citep{Zao2019simple} because it had already been successfully fitted to other tree crops in south America.
+The model simulates crop development, growth and yield, and predicts the maturation day when the fruit is likely to be ready for harvest.
+Our aim was to develop a practical tool that supports cacao farmers in the production of high quality cacao.
+Thus, making Colombian cacao more competitive in the Fine-cocoa market.}
 
 \section{Materials and Methods}
-\subsection{Floral Phenology of Cocoa }
-Normally the phenological stages in cocoa are  divided in two main phases: vegetative and reproductive.
-In the SIMPLE model we simulate the reproductive phase (fig.
-\ref{fig:pheno}, a) described by the floral phenology from the date of inflorescence emergence (BBCH scale 5) (fig.
-\ref{fig:pheno},b ) to predict the date of ripening of fruit and seed (BBCH 8) (fig.
-\ref{fig:pheno},c ) \citep{Niemenak2010} .
-In the Andean region the reproductive phase is cyclically fulfilled during two annual cycles passing by the following phases: inflorescence emergence, flowering, pollination, fruit development and harvest.
-Therefore, for modelling parametrization the crop cycle of cocoa as perennial plant does not start at the plantation date such as annual crops systems.
-Instead  the start point of the cocoa crop cycle to model is the inflorescence emergence date (fig.
+
+\subsection{Floral Phenology of Cocoa}
+
+Usually, the phenological stages of a cacao tree are divided in two main phases: vegetative and reproductive.
+Defining the reproductive phase (fig.
+\ref{fig:pheno}, a) as described by the floral phenology from the date of inflorescence emergence (BBCH scale 5) (fig.
+\ref{fig:pheno},b ) to the date of ripening of fruit and seed (BBCH 8) (fig.
+\ref{fig:pheno},c ) \citep{Niemenak2010}.
+The reproductive phase in the Andean region of Colombia is cyclically fulfilled during two annual cycles which goes through the following phases: inflorescence emergence, flowering, pollination, fruit development and harvest.
+Therefore, for modelling parameterisation the crop cycle of cocoa as perennial plant does not start at the plantation date such as annual crops systems.
+Instead, the start point of the cocoa crop cycle is the inflorescence emergence date (fig.
 \ref{fig:pheno},b).
-Consequently , the growth period of the fruit can varied from  110 to 150  daa (days after anthesis) \citep{lopez2018} when cacao fruits reaches the physiological maturity, but it can be harvested at 170 days daa \citep{Niemenak2010} for quality purposes.
+Consequently , the growth period of the fruit can vary from  110 to 150  daa (days after anthesis) \citep{lopez2018} when cacao fruits reach the physiological maturity, but it can be harvested at 170 days daa \citep{Niemenak2010} for quality purposes.
 
-Cocoa is a cauliflorous plant, which means that flowers grow on the trunk and branches.
-In Colombia cocoa trees usually produce flowers throughout the year.
-Cocoa trees produces with up to 10000 flowers per tree each year, which the 50 \%  do not develop into ripe fruits according to Fedecacao reports.
-The flower takes 30 days passing by 12 micro-stages from meristem development (stages 1to 6) to the fully developed flower (Stages 7 to 12) \citep{swanson2005} when it is ready to be pollinated.
+{\color{red} Cocoa is a cauliflorous plant, which means that flowers grow on the trunk and branches.
+Cocoa trees usually produce up to 10000 flowers per tree each year, 50 \%  of them do not develop into ripe fruits (Personal communication with Fedecacao).
+Flower development takes approximately 30 days across 12 micro-stages, from} meristem development (stages 1 to 6) to the fully developed flower (Stages 7 to 12) \citep{swanson2005, lopez2018} when it is ready to be pollinated.
 The opening of flowers or anthesis  occurs over a 12-hour period during the night and it is synchronised between the groups of mature flowers \citep{Niemenak2010}.
 However, the live of a flower can last approximately 1 day after the opening falling form the trunk if it is unfertilised \citep{cheesman1927, Niemenak2010}.
-
-Subsequentially, after anthesis the fruit growths by approximately 150 days until the maturation, mucilage.
+{\color{red} Subsequentially, after anthesis, fruits growth for approximately 150 days until the maturation, mucilage}.
 Therefore, the complete maturation process of the fruit, from the pollination to fully mature fruit, takes 160- 210 days \citep{berry1994}.
-The accumulation of lipids, storage proteins and anthocyanin starts about 85 days after pollination when fruits have an active metabolism and seeds   moisture content decreases up to 30\% \citep{Lehrian1980, Niemenak2010}.
+The accumulation of lipids, storage proteins and anthocyanin start about 85 days after pollination when fruits have an active metabolism and seeds moisture content decreases up to 30\% \citep{Lehrian1980, Niemenak2010}.
 During this phase the quality of cocoa seeds is defined.
@@
 
-\subsection{Test site and yield production }
-Cacao is cultivated in 30 states out of 32 from the total Colombian territory with about 147,000 ha \citep{Meza2021}.
-Cocoa fields studied were located in Saravena (Arauca), Rionegro (Santander), Cali (Valle del Cauca), Apartado (Antioquia) and Manizalez (Caldas) (Fig.\ref{fig:yield}, b).
-We considered 112 parcels data from Fedecacao reports.
-Each parcel data contained age of crop, density of planting, yield, number of fruits harvested, flowering date, harvest date.
-Acording to personal comunication from framers the biggest flowering occurs in September and January to harvest in March and July.
-However, the data reported showed that this flowering and pod productions are not constant for all the regions.
-Cali and Caldas presented the lowest production.
-However, the pod harvest in Caldas increased in May and from October to December, and decreased from January to March.
-Meanwhile, Arauca and Apartado reported the highest yield in the months January, July, November and December.
-Santander has picks of production in March, May and September (Fig.\ref{fig:yield}, a).
+\subsection{Inputs and Data Acquisition }
+
+{\color{olive} Input variables required to run the cocoa model included the flowering date and daily weather of solar radiation (SRAD), maximum and minimum temperature (TMAX, TMIN) and rain.
+Therefore, weather data as csv file was  downloaded from the POWER Data Access Viewer \citep{nasapower} from January 01 of 2018 to December 31 of 2020 for the five locations in Colombia (Fig.\ref{fig:yield}, b).
+The csv file had to be transformed to .WHT file using R 1.4 version \citep{Rstudio2020}.
+
+Field data for this study was provide by Fedecacao  (National Cacao Producers Federation).
+Reports contained information of 23 flowering dates from 12-July-2019 to 23-June-2020.
+Each of these dates had  their corresponding date of harvest after at exactly 180 DAF,  the age of the trees, plant density (trees ha$^{-1}$), yield  (dry beans kg ha$^{-1}$) and number of fruits harvested per hectare.
+Thus, 23 flowering dates for five farms gave us 115 samples in total.}
+
+\subsection{Test Site and Yield Production }
+
+{\color{olive} Five farms in different regions were tested : Saravena (Arauca), Rionegro (Santander), Cali (Valle del Cauca), Apartado (Antioquia) and Manizalez (Caldas).
+These were considered because had the data available of production in the reports of  Fedecacao.
+Yield depend on the successfully development of flowers to form ripe pods.
+} {\color{red}  According to personal communication from farmers, the highest flowering season occurs in September and January suggesting that harvest occurs in March and July.
+However, the data collected from the farms suggested that flowering and pod production were not constant for all the locations.
+For example, pod harvest in Caldas farm increased in May and from October to December, and decreased from January to March.
+Meanwhile, the Arauca and Apartado farms reported the highest yield in the months January, July, November and December.
+Santander farm had picks of production in March, May and September (Fig.\ref{fig:yield}, a).} {\color{olive} However,  cacao is cultivated in 30 regions of Colombian which accounts for approximately about 147,000 ha \citep{Meza2021}.
+Therefore, the national production of cocoa beans in 2020 was  63.416 ton according  to  information  published by Fedecacao (\citep{Fede2021}).
+The highest yield was reposted by Santander region with 26,315,  following in descending order by Antioquia with 5.974 tons, Arauca with 5.082 tons, Caldas with 1.343 tons and Valle del Cauca with 339 tons.}
 
@@
 	\includegraphics[scale=0.3]{images/map.png}\\
-	\caption{\footnotesize {Cocoa production characterization of cocoa for five locations\\}}
+	\caption{\footnotesize {Cocoa production  of five farms in different regions.\\}}
 	\label{fig:yield}
+	{\footnotesize (a) Production per month (2019 - 2020) of five farms. (b) Map showing the regions where farms are located.}
 \end{figure}
@@
 
-\subsection{Weather conditions}
-\begin{figure}[h!]
-	\centering
-	\includegraphics[scale=0.4]{images/clima.png}
-	\caption{\footnotesize {Colombian weather conditions. \\}}	
-	\label{fig:temp}
-	{\footnotesize (a), Available photosynthetic solar radiation (PAR).
-(b), Daily average temperature  (c), Monthly average of precipitation (bars) and relative humidity (dotted line) from 2018 to 2021 }
-\end{figure}
-\newpage
+\subsection{Thermal Time for Pod Harvest Date Identification}
 
-Solar radiation (PAR), temperature, precipitation ant relative humidity was studied from 2018 to 2020 (Fig.\ref{fig:temp}).
-Santander and Caldas had the biggest variability and the maximum of solar radiation values over 20 MJ m$^{2}$day$^{-1}$.
-In contrast , Cali and Arauca presented the lowest values of PAR below 5 MJ m$^{2}$day$^{-1}$.
-(Fig.\ref{fig:temp}, a).
-Even though, Cali and Santander had contrasting PAR conditions, they regions presented  have similar temperature during 2020 (Fig.\ref{fig:temp}, b).
-The temperature ranges from 16 to 28 $^\circ$C and it is relatively constant for each region.
-However, Arauca presented the biggest variability with hotter months during the first half of the year 2019 and 2020.
-Apartado was found as the hottest region studied with 26 $^\circ$C and Caldas as the coldest site with 18 $^\circ$C.
-Precipitation in Colombia is presented in two seasons per year from February to April and from October to November, while the relative humidity remain constant over 80\% (Fig.\ref{fig:temp},c).
-In general, the coldest regions tested (Caldas and Santander) had the maximun values of solar radiation available for photosynthesis.
+{\color{olive} The cumulative sum of daily temperature from a reference day 0 is defined as \emph{Thermal time} and  its units of measurement are in days degrees (days $^\circ$C).
+That starter point of 0 days $^\circ$C generally is the planting date \citep{Ritchie1991} but for cocoa crop simulation model it is the flowering date.
+Thermal time of a cultivated plant may consider the base temperature (T$_{b}$), which is the minimum temperature required by cocoa plant to grow.
+T$_{b}$ an vary between cultivars \citep{Slafer1995, Daymond2008}.
+For cocoa the vegetative growth T$_{b}$ has been reported between 18.6 and 20.8 $^\circ$C \citep{Daymond2008}.
+Nevertheless, the pod growth  has a lower T$_{b}$ which range between 9 and 12.9 $^\circ$C \citep{Daymond2008, lahive2019}.
+We calculated the cocoa thermal time with a pod growth T$_{b}$ of 10 $^\circ$C because it is the absolute minimum temperature for cocoa growing in South America reported by \cite{Erneholm1948}, in \citep{lahive2019}.
 
-\subsection{Thermal time for pod harvest date identification}
-It is required to model crop growth  considering the cocoa base temperature (T$_{b}$), at which the plant development stops \citep{Ritchie1991}.
-This characterization was conducted for the five regions counted 180 days (6 months) after flowering, as farmers used to harvest by calendar days.
-Characterizing the thermal time the models predicts the maturation day to harvest the cocoa can vary depending on temperature variations.
-Thermal time (degree days) was calculated  from the flowering date to the pod harvest  using the ecuation \ref{equ:tem}.
-Where tt is the cumulative sum of the daily temperature (T$_{i}$) and T$_{b}$ for cocoa is 10$^\circ$C according to \cite{lahive2019}.
+The thermal time required for the crop model was characterised for each location starting from flowering date (0 days $^\circ$C) to harvest date (180 after flowering = 6 months), as farmers used to harvest by calendar days.
+}  Thus, the cocoa model predicts the maturation day to harvest pods.
+It can vary depending on temperature variations.
+Thermal time was calculated using the equation \ref{equ:tem}.
+Where tt is the cumulative sum of the daily temperature (T$_{i}$) and T$_{b}$ for cocoa is 10$^\circ$C.
 
@@
 Once the physiological parameters were calibrated to the  simulated yields for cocoa were reasonably close to the observed yield.
-23 flowering dates from 12-July-2019 to 23-June-2020 were introduced in the treatment file, running the program and saving the results.
+The 23 flowering from Fedecacao reports were introduced in the treatment file, running the program and saving the results.
 
-\begin{enumerate}
-	\item Input/Simulation Management.csv
-	\item Input/Species parameter.csv
-	\item Input/Cultivar.csv
-	\item Input/Treatment.csv	
-	\item Input/Soil.csv
-	\item Observation/Obsdummy crop Exp name.csv	
-	\item Weather/dummy weather.WTH
-\end{enumerate}
+\subsection{Parameters}
 
-\subsection{Inputs and parameters}
-Input variables required to run SIMPLE model for cocoa include the flowering date and daily weather of solar radiation (SRAD), maximum and minimum temperature (TMAX, TMIN) and rain.
-Weather data as csv file was  downloaded from the POWER Data Access Viewer \citep{nasapower} from January 01 2018 to December 31 2020 for four locations in Colombia (Cali, Rionegro -Santander, Apartado - Antioquia, Saravena Arauca).
-The csv file had to be transformed to .WHT file using R 1.4 version \citep{Rstudio2020}.
-
-There were three parameters varied by region (table \ref{tab:reparam}) thermal time required for pod harvest after the flowering date (Tsum) , the Radiation use efficiency (RUE) and yield observed on field.
+{\color{olive}This cocoa model has  were three parameters which vary by region (table \ref{tab:reparam}): The thermal time required for harvest after the flowering date (Tsum) , the Radiation Use Efficiency (RUE) and yield observed on field.
 Physiological parameters in table \ref{tab:Treaparam} are  common  for all the regions studied.
-These parameter were calibrate for cultivars ICS95 and CCN51 considering a range of time of 200 days (DAP) from flowering date to harvest day, even thought farmers collect the pod at 180 DAP.
+These parameters were calibrate for cultivars ICS95 and CCN51 considering a range of time of 200 DAF from flowering date to harvest day, even thought farmers collect the pod at 180 DAF.
 Heat and water stress parameters were not considered.
+}
 
@@
 
-\subsection{Evaluation of model performance}
+\subsection{Evaluation of Model Performance}
 
-The SIMPLE model performance was evaluated by comparing simulated values cocoa yield with those reported by Fedecacao from cocoa plantations, using the statistical indice of relative root mean square error (RRMSE) (Equ.
+The cocoa model performance was evaluated by comparing simulated values cocoa yield with those reported by Fedecacao from cocoa plantations, using the statistical indice of relative root mean square error (RRMSE) (Equ.
 \ref{equ:RMSE}) \citep{Zao2019simple, Bai2020}.
@@
 
-\section{Results}
+\section{RESULT}
 
-\subsection{Weather effects over flowering time }
+\subsection{Weather Conditions Over Flowering Time }
 
-Figure \ref{fig:heat} shows the pearson correlation to study the weather data of the flowering time over the months of flowering  (monthF), month of harvest (monthH) and their final yield (fruit\_kg) .
-The results showed that thermal time with T$_{b}$ (ttb), temperature minimum and maximum (TMIN and TMAX) and Dew Frost Point at 2 meters (T2MDEW) are positive related.
-The wind (WS2M) is clear correlated with ttb.
+{\color{olive} The weather data from NASA platform is showed in the figure \ref{fig:temp}.
+Data from 2018 to 2020 were analysed to show the tendencies by two years when the field data was reported.} Looking at the data, Santander and Caldas had the biggest variability and the maximum of solar radiation values over 20 MJ m$^{2}$day$^{-1}$.
+In contrast , Cali and Arauca presented the lowest values of PAR below 5 MJ m$^{2}$day$^{-1}$.
+(Fig.\ref{fig:temp}, a).
+Even though, Cali and Santander had contrasting PAR conditions, they regions presented  have similar temperature during 2020 (Fig.\ref{fig:temp}, b).
+The temperature ranges from 16 to 28 $^\circ$C and it is relatively constant for each region.
+However, Arauca presented the biggest variability with hotter months during the first half of the year 2019 and 2020.
+Apartado was found as the hottest region studied with 26 $^\circ$C and Caldas as the coldest site with 18 $^\circ$C.
+Precipitation in Colombia is presented in two seasons per year from February to April and from October to November, while the relative humidity remain constant over 80\% (Fig.\ref{fig:temp},c).
+In general, the coldest regions tested (Caldas and Santander) had the maximun values of solar radiation available for photosynthesis.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[scale=0.4]{images/clima.png}
+	\caption{\footnotesize {Colombian weather conditions. \\}}	
+	\label{fig:temp}
+	{\footnotesize (a), Available photosynthetic solar radiation (PAR).
+(b), Daily average temperature  (c), Monthly average of precipitation (bars) and relative humidity (dotted line) from 2018 to 2021 }
+\end{figure}
+\newpage
+
+Figure \ref{fig:heat} shows the pearson correlation to study the weather data of the flowering time over the months of flowering  (monthF), month of harvest (monthH) and their final yield (fruit\_kg).
+The results showed that thermal time  T$_{b}$ (ttb) is 0.52 correlated with daily average temperature and  maximum temperature (TMAX)  and  temperature minimum (TMIN) and Dew Frost Point at 2 meters (T2MDEW) with  a correlation coefficient of 0.60.
+The wind (WS2M) is correlated with ttb with 0.57.
 However, less clear correlations were found of monthF  with T2MDEW, relative humidity (RH), WS2M and rain.
@@
 	\centering
-	\includegraphics[scale=0.35]{images/heatm.png}\\
-	\caption{\footnotesize {Pearson correlation average weather variables and flowering dates for all regions. \\}}
+	\includegraphics[scale=0.4]{images/heatm.png}\\
+	\caption{\footnotesize {Pearson correlation average weather variables and flowering dates for five locations in Colombia. \\}}
 	\label{fig:heat} 
+	{\footnotesize Numbers in the squares are the correlation coefficients }
 \end{figure}
@@
 
-\subsection{Thermal time }
-Thermal time characterization  was made considering 180 days after flowering (daf) for each region.
+\subsection{Thermal Time }
+
+Thermal time characterisation was made considering 180 DAF for each location.
 The boxplot in the figure \ref{fig:ttbox} shows the data distribution where boxes indicate the range of the central 50\% of the total data per region, the central line in the box is marking the median value and lines draw out from each box mean the range of the remaining data.
@@
 Meanwhile, Cali and Santander presented similar thermal time around 2000 $^\circ$C (table \ref{tab:reparam}).
-The accumulated temperature during the pod development (fig.\ref{fig:ttbox}) depends on the region where cocoa is cultivated and the variety planted.
-Thermal time values are also proportional to the yield reported on field (table \ref{tab:reparam}).\\
+The accumulated temperature during the pod development (fig.\ref{fig:ttbox}) depends on the region where cocoa is cultivated, and the variety planted.
+Thermal time values are also proportional to the yield reported on field (table \ref{tab:reparam}).
+\\
 
@@
 	\includegraphics[scale=0.3]{images/ttbbox.png}\\
-	\caption{\footnotesize {Cocoa yield and thermal time characterization at 180 days after flowering.\\}}
+	\caption{\footnotesize {Cocoa yield and thermal time characterisation at 180 days after flowering.\\}}
 	\label{fig:ttbox}
@@
 
-\subsection{Model validation}
-Therefore, biomass production of the aerial part of the plant was simulated per region (fig.
-\ref{fig:m1}, a) as biomass-rate which is the daily biomass growth rate.
-As for cocoa simulation we did not consider water and heat stress variables, in this research biomass simulation is affected by the solar radiation, RUE, daily temperature, atmospheric CO$_{2}$ concentration (ppm) and the fraction of solar radiation intercepted by a tree of cocoa during the fruit development (fSolar) (fig.
-\ref{fig:m1}, b).
-Due to the lack of this kind of data from fields, the comparison was not possible with observed data of biomass production through the crop cycle.
-However, the biomass simulation also responded proportionally with the yield production.
-Biomass of cocoa fruits simulated include all the parts that compound  the pod when they are fresh.
-Therefore, these this information may be of interest to farmers to predict pod fresh mass.
+\subsection{Model Validation}
+{\color{olive}Biomass production of the aerial part of the cocoa plant was simulated which include every organ of the plant that is over the soil surface.
+It is important to calculated how much biomass  form the aerial part belongs to the the pods according to the harvest index (HI) (table \ref{tab:Treaparam}), hence cocoa seeds.
+The figure \ref{fig:m1} (a) shows the daily biomass growth rate.} Biomass simulation is affected by the solar radiation, RUE, daily temperature, atmospheric CO$_{2}$ concentration (ppm) and the fraction of solar radiation intercepted by a tree of cocoa during the fruit development (fSolar) (fig.
+\ref{fig:m1}, b).{\color{olive} There were not field data of absorption of solar radiation by the plants or biomass production then there is not observed data to compare the simulated data.}   {\color{blue}WHAT KIND OF DATA??
+I THOUGHT YOU DOWNLOADED DATA FROM NASA.
+AR: Just weather data is from NASA  } {\color{blue} WHAT DO YOU MEAN BIOMASS RESPONDED?
+AR: I meant that it is proportional to the yield.
+I DIDN'T UNDERSTAND HERE, DROP THE THE FROM BIOMASS BECAUSE WE ARE NOT TALKING ABOUT A SPECIFIC BIOMASS AR: it is biomass of the aerial part (every organ of the plant that is over the soil surface)}.
+{\color{olive} However, the daily biomass growth rate increased proportionally with the yield production (Fig.
+\ref{fig:m1}, c) .
+Thus, it was possible to calculated the final yield as the product of accumulated biomass at the harvest day and HI = 0.3 (Equ.\ref{equ:HI}).}
 
-The figure \ref{fig:m1}, b the fSolar was simulated thought the fruit development resulting that crops in  all regions reached the maximun fSolar at 0.94, except Caldas where crops reached 0.76.
+\begin{equation}
+Cocoa~yield = Accumulated biomass~X ~ HI \\
+\label{equ:HI}
+\end{equation}
+(HI) (Eq. (5))
+
+{\color{olive} The fraction of solar radiation intercepted cocoa trees during the fruit development (fSolar) was also simulated during the fruit development cycle.
+The results in the figure \ref{fig:m1} (b) showed that} all regions had the maximum fSolar at 0.94, except Caldas where crops reached 0.76.
 Apartado and Arauca reached faster this fSolar-max.
-These high values of solar radiation intercepted for photosynthesis, lasted differently depending on the region and their solar radiation (Fig.\ref{fig:temp}, a) : Apartado 66 days  from  69 to 135 days after flowering (daf),  Arauca 61 days from 72 to 133 daf, Cali 38 days from 83 to 121 daf,  Santander 24 days from 92 to 116 daf and Caldas 2 days at 125 daf.
+These high values of solar radiation intercepted for photosynthesis, lasted differently depending on the region and their solar radiation (Fig.\ref{fig:temp}, a) : Apartado 66 days  from  69 to 135 DAF,  Arauca 61 days from 72 to 133 DAF, Cali 38 days from 83 to 121 DAF,  Santander 24 days from 92 to 116 DAF and Caldas 2 days at 125 DAF.
 fSolar declined in the interception of solar radiation until the pod harvest day.
 
-The cocoa yield simulation (kg ha$^{-1}$ per year) was validated using observed data (table \ref{tab:reparam}) showing a final RRMSE of 7.2\% (fig.
+Cocoa yield simulation (kg ha$^{-1}$ per year) was validated using observed data (table \ref{tab:reparam}) showing a final RRMSE of 7.2\% (fig.
 \ref{fig:m1}, c).
@@
 (c) Yield.
-Crop cycle close to 180 daf (vertical red line) base on figure  \ref{fig:ttbox}.
+Crop cycle close to 180 DAF (vertical red line) base on figure  \ref{fig:ttbox}.
 \\ }}
@@
 \begin{table}[h!]	
-	\caption {\footnotesize {Summary of relative root mean square error RMSE for the predicted cocoa yield using SIMPLE model.}}
+	\caption {\footnotesize {Summary of relative root mean square error RMSE for yield prediction using The cocoa model.}}
 	\label{tab:error} 
@@
 
-\subsection{Prediction of the pod harvest day }
-These result demonstrated that the traditional way to harvest always at 180 daf is not having account physiological and environmental conditions that can be affecting the pod development and maturation.
-Figure \ref{fig:dayH} present that observe day of harvest or maturity day is 180 daf independently of the region.
-All the regions except Santander, presented the earliest harvest when the flowering date was between December and February.
-The fruit can be ready to harvest depending of the enviromental condition per region.
-Thus, when the fruit development was simulated the maturity day in Cali and Caldas were from 3 to 12 and 4 to 23 days before 180 daf (Fig.
+\subsection{Predicting Pod Harvest Day }
+
+{\color{olive} The cocoa model predict the day when pods can be harvested (predicted) and the results were compared with the 180 counting after flowering (observed) because it is the empirical date of harvest.
+The figure \ref{fig:dayH} present that observe day of harvest (180 DAF) independently of the region.
+All the regions except Santander, presented the earliest predicted harvest when the flowering date was between December and February.
+The fruit can be ready to harvest before or after 180 DAF depending on the environmental condition per region.
+The results demonstrated that the traditional way to harvest which is always at 180 DAF, it is not having account physiological and environmental conditions that can be affecting the pod maturation.} Thus, when the fruit development was simulated the maturity day in Cali and Caldas were from 3 to 12 and 4 to 23 days before 180 DAF (Fig.
 \ref{fig:dayH}, c and f respectively).
-Apartado presented the most similar predicted dates of harvest to 180 daf with 170 to 182 daf  (Fig.
+Apartado presented the most similar predicted dates of harvest to 180 DAF with 170 to 182 DAF  (Fig.
 \ref{fig:dayH}, d).
-The pod may be harvest in Arauca (Fig.\ref{fig:dayH}, b)  between 165 and 193 daf, Santander (Fig.\ref{fig:dayH}, a) between 165 and 183 daf, ten days less than in Arauca for the same months of flowering of July and August of 2019 .
+The pod may be harvest in Arauca (Fig.\ref{fig:dayH}, b)  between 165 and 193 DAF, Santander (Fig.\ref{fig:dayH}, a) between 165 and 183 DAF, ten days less than in Arauca for the same months of flowering of July and August of 2019 .
 Only Arauca presented longer crop cycles when the flowering was between during that period of time.
@@
 \ref{fig:temp}, b ).
+{\color{olive} To summarize, depending on which month of the year the trees are flowering,  the number of days to reach the harvest of ripe pods vary more or less 180 days (table \ref{tab:harvest}).}
 
@@
 
+\begin{table}[h!]	
+	\caption {\footnotesize {Average of days to harvest according to the month of flowering.}}
+	\label{tab:harvest} 
+	\centering
+	\begin{small}
+		{\def\arraystretch{2}\tabcolsep=10pt
+		\begin{tabular}{l c c c c c }
+			\hline
+			{\bf Month }&{\bf Santander }&{\bf Arauca}&{\bf Cali}&{\bf Apartado}&{\bf Caldas}\\
+			\hline
+			January    & 166.5 & 166.5 & 169.5& 171.5 & 158.5 \\
+			February   & 166 & 171 & 171 & 173 & 163  \\
+			March      & 165 & 171 & 172 & 176 & 167  \\
+			April      & 165.5 & 172.5 & 173& 178.5 & 170   \\
+			May       & 166.5 & 174.5 & 174.5&  181& 175   \\
+			June      & 169 & 173.5 & 177&  182& 175.5   \\
+			July      & 176 & 191 & 175& 179 & 173.5  \\
+			August    & 181 & 186 & 175& 178.5 & 171   \\
+			September   & 176 & 182.5 & 175& 176.5 & 168   \\
+			October    & 179 & 176 & 172.5& 173 & 163.5   \\
+			November   & 173.5 & 170 & 170.5&171.5 & 159.5   \\
+			December   & 169 & 166 & 168.5& 171 & 157  \\
+			\hline
+		\end{tabular} \\
+	}
+		{\footnotesize Days to harvest cocoa after flowering are approximate, as these are results from the cocoa model simulations. Calibration was based on FEDECACAO reports from 2018 to 2020. } 
+	\end{small}
+\end{table}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Moreover, the lack of water during dry seasons may reduce the nutrients uptake provoking the massive flower drops \citep{Vaughton2017}.
-For future studies, wind could be included in the SIMPLE model as an input to simulated this mechanical effects over number of flowers.
+For future studies, wind could be included in the SIMPLEcocoa model as an input to simulated this mechanical effects over number of flowers.
 In general, field data regarding counting flowers pollinated by month should be better reported for the region here studied.
@@
 \ref{fig:m1}).
-The thermal time calculation was defined base on  180 daf because farmers cut the pods by calendar days.
+The thermal time calculation was defined base on  180 DAF because farmers cut the pods by calendar days.
 Previous studies stated that evaluating the level of knowledge of growers regarding cocoa crop management, showed that the harvest was in the group of activities that presented the lowest level of information by the farmers \citep{Gutierrez2020}.
@@
 As the harvest day was predicted form the flowering date (Fig.
-\ref{fig:dayH}, consequently, biomass production and fSolar presented a crop cycle shorter than 180 daf (Fig.
+\ref{fig:dayH}, consequently, biomass production and fSolar presented a crop cycle shorter than 180 DAF (Fig.
 \ref{fig:m1}, a and b).
-These simulations are coincident with results presented by \cite{lopez2018}, where physiological maturity of coca pod varies from 140 to 162 daf.
+These simulations are coincident with results presented by \cite{lopez2018}, where physiological maturity of coca pod varies from 140 to 162 DAF.
 Our results showed how the harvest day can vary depending on the accumulate temperature during each especific crop cycle simulated (Fig.
@@
 
-Biomass simulations use SIMPLE model presented similar predicted values (10000 Kg ha$^{-1}$) for coca drops in Costa Rica using SUCROScocoa \citep{zuidema2005}.
+Biomass simulations use SIMPLEcocoa model presented similar predicted values (10000 Kg ha$^{-1}$) for coca drops in Costa Rica using SUCROScocoa \citep{zuidema2005}.
 Biomass simulations are a common evaluation in crop modells such as Sirius \citep{Crout20142}, SUBSTOR-potato \citep{Raymundo2017} and  DSSAT, CropSyst, STICS and WOFOST \citep{Confalonieri2016}.
@@
 These leaves eventually drop becoming on litter over the soil.
-Leaves life cycle has been simulated using crop models \citep{Crout2010, zuidema2005}, which can be the reference to improve our SIMPLE coca crop model in future studies.
+Leaves life cycle has been simulated using crop models \citep{Crout2010, zuidema2005}, which can be the reference to improve our SIMPLEcocoa crop model in future studies.
 
@@
 The app development is on charge of Grupo BIOS to be deliver to farmers in Caldas initially at the end of 2021.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
 This research presented and initial crop calibration that can be improved with further studies, including effects over the pod production by diseases, nutritional diffidences and abiotic stresses.
@@
 Moreover, It will be required the help of entrepreneurs, researchers, academics and non-specialized communities to transfers the knowledge to cocoa growers.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
+
 Cocoa fruit development for harvest in the right time depend on whether conditions and principles of crop physiology and flower phenology.
-This was common for the five regions.Thermal time characterization range from 1200 to 3000 days $^\circ$C, with a T$_{b}$ of 10 $^\circ$C for the fruit development.
-The SIMPLE model as a generic crop model allowed to predict the harvest date with better precision than only considering days by calendar.
+This was common for the five regions.
+Thermal time characterisation range from 1200 to 3000 days $^\circ$C, with a T$_{b}$ of 10 $^\circ$C for the fruit development.
+The SIMPLEcocoa model allowed to predict the harvest date with better precision than only considering days by calendar.
 Thus, the crop cycle of cocoa for harvest  should be shorter than 180 days after flowering.
-These results confirm the potential of crop model for tropical crop in Latin America.
+These results confirm the potential of the Crop Simulation Model approaches  for tropical crop in Latin America.
 
 \section{Author contributions}
-ARV and ARC model calibration and data analsis. AMG field data collection. ARV were primarily responsible for writing the manuscript.
+ARV and ARC model calibration and data analysis. PA and ODR field data collection. AMG, ARC, and ARV editing and results interpretations. ARV were primarily responsible for writing the manuscript.
 

### 2021-11-23 paperModel/paperModel_2.tex
submitted

@@
 \usepackage[colorinlistoftodos]{todonotes}
-\usepackage[dvipsnames]{xcolor}
+%\usepackage[dvipsnames]{xcolor}
 \newcounter{mycomment}
@@
 % Full title of the paper (Capitalized)
-\Title{A modelling strategy to improve cacao quality and productivity }
+\Title{A crop modelling strategy to improve cacao quality and productivity }
 
@@
 % Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address{$^{1}$ \quad The John Bingham Laboratory, NIAB, 93 Lawrence Weaver Road, Cambridge CB3 0LE, UK 1\\$^{2}$ \quad Grupo BIOS, Centro de Bioinformática y Biología Computacional de Colombia - BIOS, Manizales, Caldas, Colombia, South America.\\ $^{3}$ \quad Federación Nacional de Cafeteros FEDECACAO, Colombia, South America.
+\address{$^{1}$ \quad The John Bingham Laboratory, NIAB, 93 Lawrence Weaver Road, Cambridge CB3 0LE, UK \\$^{2}$ \quad Grupo BIOS, Centro de Bioinformática y Biología Computacional de Colombia - BIOS, Manizales, Caldas, Colombia, South America.\\ $^{3}$ \quad Federación Nacional de Cafeteros FEDECACAO, Colombia, South America.
 }%; e-mail@e-mail.com}
@@
 % Contact information of the corresponding author
-\corres{Correspondence: angela.romerovergel@niab.com} %; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}
+\corres{Correspondence: ing.aromerov@gmail.com; anyela.camargorodriguez@niab.com} %; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}
 
@@
 % Abstract (Do not insert blank lines, i.e. \\)
-\abstract{Cacao production systems in Latin America {\color{red}have } a high importance over social and economic development, facing  the fight against hunger and poverty.
-Although Colombian cacao has the potential to be in the high value markets for fine flavour, {\color{red} the lack of expert support as well as the use of traditional, and oftentimes suboptimal, technologies makes cocoa production negligibly.
+
+\abstract{Cacao production systems in Latin America have a high importance over social and economic development, facing  the fight against hunger and poverty.
+Although Colombian cacao has the potential to be in the high value markets for fine flavour, the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly.
 For example, traditionally cacao harvest takes place at 5 or 6 months after flowering, other environmental parameters that have more association with pod maturation speed are not taken into account.
-Cocoa fruits development can be considered as the result of a number of physiological and morphological processes that can be described by mathematical relationships even under uncontrolled environments.
+Cacao fruits development can be considered as the result of a number of physiological and morphological processes that can be described by mathematical relationships even under uncontrolled environments.
 In this context, crop models are useful tools to simulate and predict crop development over time and under multiple environmental conditions.
 Since, harvesting at the right time can yield high quality cacao, we parametrised a crop model to predict the best time for harvest cocoa fruits in Colombia.
-Our aim was to develop a practical tool that supports cacao farmers in the production of high quality cacao.} When comparing simulated and observed data, our results showed an RRMSE of 7.2\% for the yield prediction, while the simulated harvest date varied between +/- 2 to 20 days depending on the temperature variations of the year between regions.
-{\color{red} This crop model contributed to understand and predict the phenology of cacao fruits for two key cacao varieties ICS95 y CCN51.}
-}
+Our aim was to develop a practical tool that supports cacao farmers in the production of high quality cacao.
+When comparing simulated and observed data, our results showed an RRMSE of 7.2\% for the yield prediction, while the simulated harvest date varied between +/- 2 to 20 days depending on the temperature variations of the year between regions.
+This crop model contributed to understand and predict the phenology of cacao fruits for two key cultivars ICS95 y CCN51.}
 
 % Keywords
-\keyword{ICS95; CCN51; thermal time, flowering date }
+\keyword{thermal time, flowering date (FD), harvest prediction, biomass, light interception }
 
@@
 
-Cocoa ( \textit{Theobroma cacao }L.
+Cacao ( \textit{Theobroma cacao }L.
 ) is an important worldwide perennial tropical crop endemic to the South American rainforests \citep{zuidema2005, motamayor2002, argout2011, Rodriguez2019}.
 Cacao plant is a member of the Malvaceae (formerly Sterculiaceae)  botanical family such as  cotton \textit{ Gossypium hirstium} \citep{Nix2017cotton}.
-Coton has been is modeled in SIMPLE model \citep{Zao2019simple}.
-Cocoa is grown for its fruits, known as cacao pods \citep{ Niemenak2010, suarez2021}.
+Cotton has been is modeled in SIMPLE model \citep{Zao2019simple}.
+Cacao is grown for its fruits, known as cacao pods \citep{ Niemenak2010, suarez2021}.
 Only the 5\% of the world cocoa yield is destined for Fine-cocoa production due to the low productivity through  the traditional crop management \citep{argout2011}.
@@
 
-Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}, {\color{red} the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly.
-For example, the farmer practice of empirically harvesting from 5 to 6 months or {\color{olive} 180 days } after flowering date (DAF), produced a mix of quality of cocoa beans that are then be fermented.
-This practice produces heterogeneous characteristics between each fermentation batch which potentially diminish the quality of the end cocoa product} \citep{Escobar2021}.
+Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}, the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly.
+For example, the farmer practice of empirically harvesting from 5 to 6 months or 180 Days After Flowering  (DAF) date, produced a mix of quality of cacao beans that are then fermented.
+This practice produces heterogeneous characteristics between each fermentation batch which potentially diminish the quality of the final cocoa product \citep{Escobar2021}.
 To identify the best moment to harvest is important to consider physiological responses affected by climate variables such as rain, solar radiation and wind.
-Thus,  for cocoa in Colombia, physiological simulation models may be valuable to identify the best moment to harvest cocoa considering variable weather conditions, soil types and cultivar specifications.
+Therefore,  for cacao production in Colombia, physiological simulation models may be valuable to identify the best moment to harvest cocoa fruits considering variable weather conditions, soil types and cultivar specifications.
 
@@
 However, the information reported  is be less than for annual crops due to the lack of field data available , relatively high research costs and the difficulties of accumulated errors in long-term simulations \citep{zuidema2005}.
-For cacao there  approaches  to predict yield mainly  using algorithms of machine learning \citep{lamos2020} and just one mechanistic model simulates physiological cocoa performance  "SUCROS-cocoa" \citep{zuidema2005}.
+For cacao there are  approaches  to predict yield mainly  using algorithms of machine learning \citep{lamos2020} and just one mechanistic model simulates physiological cocoa performance  "SUCROS-cocoa" \citep{zuidema2005}.
 This crop model calculates light interception, photosynthesis, maintenance respiration,
@@
 
-In this paper, we present a physiological parametrisation of SIMPLE crop model for cocoa to predict best harvest time and overall yield.
-{\color{red} We parameterised the SIMPLE crop model \citep{Zao2019simple} because it had already been successfully fitted to other tree crops in south America.
+In this paper, we present a physiological parametrisation of the SIMPLE crop model \citep{Zao2019simple} for cocoa to predict best harvest time and overall yield in the framework of KOCOLATL project, which is a collaborative work between Colombian partners (researches, farmers and organizations such as Fedecacao and BIOS) and researchers in NIAB from United Kingdom.
+We parametrised this crop model  because it had already been successfully fitted to other tree crops in south America.
 The model simulates crop development, growth and yield, and predicts the maturation day when the fruit is likely to be ready for harvest.
-Our aim was to develop a practical tool that supports cacao farmers in the production of high quality cacao.
-Thus, making Colombian cacao more competitive in the Fine-cocoa market.}
+Our aim was to develop a practical tool that supports cacao farmers in the production of high quality cocoa.
+Thus, making Colombian cacao more competitive in the Fine-cocoa market.
 
@@
 
-Usually, the phenological stages of a cacao tree are divided in two main phases: vegetative and reproductive.
-Defining the reproductive phase (fig.
-\ref{fig:pheno}, a) as described by the floral phenology from the date of inflorescence emergence (BBCH scale 5) (fig.
+The phenological stages of a cacao tree are divided in two main phases, vegetative and reproductive.
+The reproductive phase (fig.
+\ref{fig:pheno}, a) is represented by the floral phenology which goes from the date of inflorescence emergence (BBCH scale 5) (fig.
 \ref{fig:pheno},b ) to the date of ripening of fruit and seed (BBCH 8) (fig.
 \ref{fig:pheno},c ) \citep{Niemenak2010}.
-The reproductive phase in the Andean region of Colombia is cyclically fulfilled during two annual cycles which goes through the following phases: inflorescence emergence, flowering, pollination, fruit development and harvest.
-Therefore, for modelling parameterisation the crop cycle of cocoa as perennial plant does not start at the plantation date such as annual crops systems.
-Instead, the start point of the cocoa crop cycle is the inflorescence emergence date (fig.
+The reproductive phase is cyclically across two annual cycles which go through the following phases: inflorescence emergence, flowering, pollination, fruit development and harvest.
+Therefore, to model the cacao crop cycle as perennial plant, we regarded the the start point of the cocoa crop cycle as the inflorescence emergence date, as opposed to the sowing/emergence date usually used to model annual crops (fig.
 \ref{fig:pheno},b).
-Consequently , the growth period of the fruit can vary from  110 to 150  daa (days after anthesis) \citep{lopez2018} when cacao fruits reach the physiological maturity, but it can be harvested at 170 days daa \citep{Niemenak2010} for quality purposes.
+Therefore, the growth period of the fruit could vary from 110 to 150 daa (days after anthesis) \citep{lopez2018} when cacao fruits reach the physiological maturity, but it can be harvested at 170 days daa \citep{Niemenak2010} for quality purposes.
 
-{\color{red} Cocoa is a cauliflorous plant, which means that flowers grow on the trunk and branches.
-Cocoa trees usually produce up to 10000 flowers per tree each year, 50 \%  of them do not develop into ripe fruits (Personal communication with Fedecacao).
-Flower development takes approximately 30 days across 12 micro-stages, from} meristem development (stages 1 to 6) to the fully developed flower (Stages 7 to 12) \citep{swanson2005, lopez2018} when it is ready to be pollinated.
+Since cacao is a cauliflorous plant, flowers grow on trunk and branches.
+Cacao trees usually produce up to 10000 flowers per tree each year, 50 \%  of which do not develop into ripe fruits (Personal communication with Fedecacao).
+Flower development takes approximately 30 days across 12 micro-stages, from meristem growth (stages 1 to 6) to the fully developed flower (Stages 7 to 12) \citep{swanson2005, lopez2018} when it is ready to be pollinated.
 The opening of flowers or anthesis  occurs over a 12-hour period during the night and it is synchronised between the groups of mature flowers \citep{Niemenak2010}.
-However, the live of a flower can last approximately 1 day after the opening falling form the trunk if it is unfertilised \citep{cheesman1927, Niemenak2010}.
-{\color{red} Subsequentially, after anthesis, fruits growth for approximately 150 days until the maturation, mucilage}.
+However, the life of a flower can last approximately 1 day after the opening, when it falls form the trunk if it is unfertilised \citep{cheesman1927, Niemenak2010}.
+Subsequentially, after anthesis, fruits growth for approximately 150 days until the maturation, mucilage.
 Therefore, the complete maturation process of the fruit, from the pollination to fully mature fruit, takes 160- 210 days \citep{berry1994}.
@@
 	\centering
-	\includegraphics[scale=0.4]{images/phenology.png}\\
+	\includegraphics[scale=0.15]{images/phenology.png}\\
 	\caption{\footnotesize {Phenology of cocoa in Colombia for crop modelling.\\}} 
-	\footnotesize{Source: Taken from from Dreamstime.com, phys.org \citep{toledo2021} and \cite{lopez2018}}
+	\footnotesize{Source: Taken from Dreamstime.com, phys.org \citep{toledo2021} and \cite{lopez2018}}
 	\label{fig:pheno}
@@
 
+\subsection{Test Site and Yield Production }
+
+To calibrate and evaluate the cacao model, 23 field data samples were collected by Fedecacao, from five farms located in the geographic regions of Saravena (Arauca), Rionegro (Santander), Cali (Valle del Cauca), Apartado (Antioquia) and Manizalez (Caldas) (Fig.\ref{fig:yield}, a).
+Each field sample reported an observed flowering date and corresponding yield, for a given tree and month of the year.
+Field samples reported sights of 23 flowering dates from 12-July-2019 to 23-June-2020.
+Each of these dates had their corresponding date of harvest after at exactly 180 DAF,  the age of the trees, plant density (trees ha$^{-1}$), yield  (dry beans kg ha$^{-1}$) and number of fruits harvested per hectare.
+Thus, 23 flowering dates for five farms gave us 115 samples in total.
+
 \subsection{Inputs and Data Acquisition }
 
-{\color{olive} Input variables required to run the cocoa model included the flowering date and daily weather of solar radiation (SRAD), maximum and minimum temperature (TMAX, TMIN) and rain.
-Therefore, weather data as csv file was  downloaded from the POWER Data Access Viewer \citep{nasapower} from January 01 of 2018 to December 31 of 2020 for the five locations in Colombia (Fig.\ref{fig:yield}, b).
-The csv file had to be transformed to .WHT file using R 1.4 version \citep{Rstudio2020}.
+In order to model the phenology of the cacao fruit, linked to collected field data, the cacao model required specific input variables such as Flowering Date (FD), Daily Solar Radiation (SRAD), Daily Maximum and Minimum Temperature (TMAX, TMIN) and Daily Precipitation, specific to a particular agroclimatic region.
+FD was extracted from the field data, corresponding weather data linked to geographical region was sourced from the POWER Data Access Viewer \citep{nasapower}, from January 01 of 2018 to December 31 of 2020 (Fig.\ref{fig:yield}, b).
+Weather data was processed using R (R 1.4 version) \citep{Rstudio2020}.
 
-Field data for this study was provide by Fedecacao  (National Cacao Producers Federation).
-Reports contained information of 23 flowering dates from 12-July-2019 to 23-June-2020.
-Each of these dates had  their corresponding date of harvest after at exactly 180 DAF,  the age of the trees, plant density (trees ha$^{-1}$), yield  (dry beans kg ha$^{-1}$) and number of fruits harvested per hectare.
-Thus, 23 flowering dates for five farms gave us 115 samples in total.}
-
-\subsection{Test Site and Yield Production }
-
-{\color{olive} Five farms in different regions were tested : Saravena (Arauca), Rionegro (Santander), Cali (Valle del Cauca), Apartado (Antioquia) and Manizalez (Caldas).
-These were considered because had the data available of production in the reports of  Fedecacao.
-Yield depend on the successfully development of flowers to form ripe pods.
-} {\color{red}  According to personal communication from farmers, the highest flowering season occurs in September and January suggesting that harvest occurs in March and July.
-However, the data collected from the farms suggested that flowering and pod production were not constant for all the locations.
-For example, pod harvest in Caldas farm increased in May and from October to December, and decreased from January to March.
-Meanwhile, the Arauca and Apartado farms reported the highest yield in the months January, July, November and December.
-Santander farm had picks of production in March, May and September (Fig.\ref{fig:yield}, a).} {\color{olive} However,  cacao is cultivated in 30 regions of Colombian which accounts for approximately about 147,000 ha \citep{Meza2021}.
-Therefore, the national production of cocoa beans in 2020 was  63.416 ton according  to  information  published by Fedecacao (\citep{Fede2021}).
-The highest yield was reposted by Santander region with 26,315,  following in descending order by Antioquia with 5.974 tons, Arauca with 5.082 tons, Caldas with 1.343 tons and Valle del Cauca with 339 tons.}
+Since cacao yield depends on the successful development of flowers to form ripe pods, the current view from Colombian farmers (personal communication with Fedecacao)  is that the highest flowering season occurs in September and January which suggests that harvest occurs in March and July.
+These communication were given through weekly online meetings of KOCOLATL project.
+However, the field data collected from Fedecacao suggested that flowering and pod production were not constant for all the locations.
+For example, pod harvest in the Caldas farm increased in May, and from October to December, and decreased from January to March.
+The Arauca and Apartado farms reported highest yield in the months of January, July, November and December.
+The Santander farm had picks of production in March, May and September (Fig.\ref{fig:yield}, b).
 
@@
 	\label{fig:yield}
-	{\footnotesize (a) Production per month (2019 - 2020) of five farms. (b) Map showing the regions where farms are located.}
+	{\footnotesize (a) Map showing the regions where farms are located.  (b) Production per month (2019 - 2020) of five farms.}\\
 \end{figure}
-\newpage
 
@@
 
-{\color{olive} The cumulative sum of daily temperature from a reference day 0 is defined as \emph{Thermal time} and  its units of measurement are in days degrees (days $^\circ$C).
-That starter point of 0 days $^\circ$C generally is the planting date \citep{Ritchie1991} but for cocoa crop simulation model it is the flowering date.
-Thermal time of a cultivated plant may consider the base temperature (T$_{b}$), which is the minimum temperature required by cocoa plant to grow.
-T$_{b}$ an vary between cultivars \citep{Slafer1995, Daymond2008}.
-For cocoa the vegetative growth T$_{b}$ has been reported between 18.6 and 20.8 $^\circ$C \citep{Daymond2008}.
+The cumulative sum of daily temperature from a reference day 0 is defined as \emph{Thermal time} and  its units of measurement are in days degrees (days $^\circ$C).
+That starter point of 0 days $^\circ$C generally is the planting date \citep{Ritchie1991} but as indicated earlier, our cacao model used FD as the starting point.
+Thermal time for the development of a cultivated plant may consider the base temperature (T$_{b}$), which is the minimum temperature required by cacao plants to grow.
+In addition, T$_{b}$ can vary between cultivars \citep{Slafer1995, Daymond2008}.
+For cacao the vegetative growth T$_{b}$ has been reported between 18.6 and 20.8 $^\circ$C \citep{Daymond2008}.
 Nevertheless, the pod growth  has a lower T$_{b}$ which range between 9 and 12.9 $^\circ$C \citep{Daymond2008, lahive2019}.
-We calculated the cocoa thermal time with a pod growth T$_{b}$ of 10 $^\circ$C because it is the absolute minimum temperature for cocoa growing in South America reported by \cite{Erneholm1948}, in \citep{lahive2019}.
+We calculated the cacao thermal time with a pod growth T$_{b}$ of 10 $^\circ$C because it is the absolute minimum temperature for cocoa growing in South America reported by \cite{Erneholm1948}, in \citep{lahive2019}.
 
-The thermal time required for the crop model was characterised for each location starting from flowering date (0 days $^\circ$C) to harvest date (180 after flowering = 6 months), as farmers used to harvest by calendar days.
-}  Thus, the cocoa model predicts the maturation day to harvest pods.
+The thermal time required for the cacao crop model was characterised for each location starting from FD (0 days $^\circ$C) to harvest date (180 after flowering = 6 months), as farmers used to harvest by calendar days.
+Thus, the cacao model predicts the maturation day to harvest pods.
 It can vary depending on temperature variations.
@@
 
-Calibration of crop models are conducted typically for particular cultivars and require site specific inputs of weather \citep{Crout20142}.
-The procedure for the SIMPLE model \citep{Zao2019simple} calibration  was a sequential process of modifying physiological variables specific for cocoa in the inputs files and  adding the appropriate files of weather for each region tested.
-In the SIMPLE model has seven input files where new cocoa crop data should be provided.
+Calibration of any crop model is conducted typically for a particular cultivar and agroclimatic region \citep{Crout20142}.
+The cacao model was calibrated, for five particular regions and two cultivars, by sequentially modifying the physiological variables and then comparing the degree of similarity between observed and predicted data \citep{Crout2008chapter, Zao2019simple,Camargo2019aquacropr, Bai2020}.
+The original code of SIMPLE model from \citep{Zao2019simple} was modified into the cocoa model.
+The process of calibration stops when the distance between observed and simulated data doesn't improved any longer.
+The cacao model has seven input files where new cocoa crop data are provided.
 Files in the list below can be edited to define the features of the new cultivars or experiments.
-Cocoa crop has not been simulated with SIMPLE model, hence consider previous cocoa studies, values such as leaf area index (LAI) \citep{Agele2016} and Harvest index (HI) \citep{Quintana2015} where modified.
-Radiation Use Efficiency (RUE) was calibrate according to the perennial crops  (banana and cotton) that had been calibrated previously in SIMPLE model \citep{Zao2019simple} with a RUE of 0.8 and 0.85 respectively.
-As cocoa trees are under shadow the RUE was lower with values between 0.7 and 0.5 g MJ$^{\mathbf{-1}}$ m$^{2}$ (table \ref{tab:reparam}).
-Once the physiological parameters were calibrated to the  simulated yields for cocoa were reasonably close to the observed yield.
-The 23 flowering from Fedecacao reports were introduced in the treatment file, running the program and saving the results.
+Since cacao phenology has not been simulated with the SIMPLE model, cacao physiological information such as Leaf Area Index (LAI) for shade plants \citep{Agele2016, Soltani2012, zuidema2005,Baracaldo2014} and Harvest Index (HI) \citep{Quintana2015} were extracted from the current literature and verified with Fedecacao cacao experts.
+Other variables such as Radiation Use Efficiency (RUE) \citep{Fletcher2013RUE,Bonhomme2000} were based on the perennial crops banana and cotton which have been calibrated previously in the SIMPLE model \citep{Zao2019simple} using RUEs of 0.8 and 0.85 for banana and cotton, respectively.
+As cocoa trees usually grow under shadow \citep{lahive2019}, we estimated by trial and error that lower RUE values (between 0.7 and 0.5 g MJ$^{\mathbf{-1}}$ m$^{2}$) than those utilized by \cite{Zao2019simple}, was a most ideal range  (table \ref{tab:reparam}).
 
@@
 
-{\color{olive}This cocoa model has  were three parameters which vary by region (table \ref{tab:reparam}): The thermal time required for harvest after the flowering date (Tsum) , the Radiation Use Efficiency (RUE) and yield observed on field.
+This cocoa model has three parameters which vary by region (table \ref{tab:reparam}): The thermal time required for harvest after the FD (Tsum) , the Radiation Use Efficiency (RUE) and yield observed on field.
 Physiological parameters in table \ref{tab:Treaparam} are  common  for all the regions studied.
-These parameters were calibrate for cultivars ICS95 and CCN51 considering a range of time of 200 DAF from flowering date to harvest day, even thought farmers collect the pod at 180 DAF.
-Heat and water stress parameters were not considered.
-}
+These parameters were calibrate for cultivars ICS95 and CCN51 considering a range of time of 200 DAF to harvest day, even thought farmers collect the pod at 180 DAF.
+Heat and water stress parameters were not considered as this study was not assessing biotic o abiotic stresses.
 
@@
 \begin{table}[h!]	
-	\caption {\footnotesize {Parameter values used to run SIMPLEcocoa model.}}
+	\caption {\footnotesize {Parameter values used to run the cacao model.}}
 	\centering
@@
 			&CO$_{2}$ & 400 ppm\\
-			&SowingDate &Flowering date\\
+			&SowingDate &Flowering Date (FD)\\
 			\hline
@@
 
-The cocoa model performance was evaluated by comparing simulated values cocoa yield with those reported by Fedecacao from cocoa plantations, using the statistical indice of relative root mean square error (RRMSE) (Equ.
-\ref{equ:RMSE}) \citep{Zao2019simple, Bai2020}.
+The cocoa model performance was evaluated by comparing simulated values cocoa yield with those reported by Fedecacao from cocoa plantations, using the statistical indice of Relative Root Mean Square Error (RRMSE)  in the equation \ref{equ:RRMSE}, where n is the total number of observations, Y$_{i}$ corresponds to observed value from field and X$_{i}$ is the predict value from the model.\\
 
 \begin{equation}
-RMSE= \sqrt{\frac{1}{n}  \sum_{i=1}^{n} (Y_{i}-X_{i})^{2} } 
-\label{equ:RMSE}
+RRMSE= \sqrt{\frac{\frac{1}{n}  \sum_{i=1}^{n} (Y_{i}-X_{i})^{2}}{\sum_{i=1}^{n} X_{i}^{2} } } \times 100\% 
+\label{equ:RRMSE}
 \end{equation}
 
-\section{RESULT}
+\section{Results}
 
 \subsection{Weather Conditions Over Flowering Time }
-
-{\color{olive} The weather data from NASA platform is showed in the figure \ref{fig:temp}.
-Data from 2018 to 2020 were analysed to show the tendencies by two years when the field data was reported.} Looking at the data, Santander and Caldas had the biggest variability and the maximum of solar radiation values over 20 MJ m$^{2}$day$^{-1}$.
-In contrast , Cali and Arauca presented the lowest values of PAR below 5 MJ m$^{2}$day$^{-1}$.
+We analysed and compared weather patterns linked our five geographical regions (Fig.\ref{fig:yield}, a) and across 2018 to 2020 .
+Looking at the data, the regions of Santander and Caldas had the biggest variability and the maximum of solar radiation values over 20 MJ m$^{2}$day$^{-1}$ .
+In contrast , the regions of Cali and Arauca presented the lowest values of PAR below 5 MJ m$^{2}$day$^{-1}$.
 (Fig.\ref{fig:temp}, a).
-Even though, Cali and Santander had contrasting PAR conditions, they regions presented  have similar temperature during 2020 (Fig.\ref{fig:temp}, b).
+Even though, Cali and Santander had contrasting PAR conditions, these regions presented  have similar temperature during 2020 (Fig.\ref{fig:temp}, b).
 The temperature ranges from 16 to 28 $^\circ$C and it is relatively constant for each region.
 However, Arauca presented the biggest variability with hotter months during the first half of the year 2019 and 2020.
-Apartado was found as the hottest region studied with 26 $^\circ$C and Caldas as the coldest site with 18 $^\circ$C.
+Apartado was found as the hottest region studied with a relative constant temperature of 26 $^\circ$C and Caldas as the coldest site with 16 $^\circ$C.
 Precipitation in Colombia is presented in two seasons per year from February to April and from October to November, while the relative humidity remain constant over 80\% (Fig.\ref{fig:temp},c).
-In general, the coldest regions tested (Caldas and Santander) had the maximun values of solar radiation available for photosynthesis.
+In general, the coldest regions tested (Caldas and Santander) had the maximum values of solar radiation available for photosynthesis.
 
@@
 Figure \ref{fig:heat} shows the pearson correlation to study the weather data of the flowering time over the months of flowering  (monthF), month of harvest (monthH) and their final yield (fruit\_kg).
-The results showed that thermal time  T$_{b}$ (ttb) is 0.52 correlated with daily average temperature and  maximum temperature (TMAX)  and  temperature minimum (TMIN) and Dew Frost Point at 2 meters (T2MDEW) with  a correlation coefficient of 0.60.
+The results showed that thermal time T$_{b}$ (ttb) is correlated (P = 0.52) with daily average temperature and maximum temperature (TMAX) and  temperature minimum (TMIN) and Dew Frost Point at 2 meters (T2MDEW) with  a correlation coefficient of 0.60.
 The wind (WS2M) is correlated with ttb with 0.57.
 However, less clear correlations were found of monthF  with T2MDEW, relative humidity (RH), WS2M and rain.
+Although, farmers stated that the number of flowers pollinated decrease by months where wind and rain are high (personal communication with Fedecacao), our analysis  (Fig.
+\ref{fig:heat}) could not show a correlation between flower shedding and high wind or rain because information on number of  flowers at FD was not part of the field data collected.
 
@@
 	\centering
-	\includegraphics[scale=0.4]{images/heatm.png}\\
-	\caption{\footnotesize {Pearson correlation average weather variables and flowering dates for five locations in Colombia. \\}}
+	\includegraphics[scale=0.2]{images/heatm.png}\\
+	\caption{\footnotesize {Pearson correlation average weather variables and FD for five locations in Colombia. \\}}
 	\label{fig:heat} 
@@
 The boxplot in the figure \ref{fig:ttbox} shows the data distribution where boxes indicate the range of the central 50\% of the total data per region, the central line in the box is marking the median value and lines draw out from each box mean the range of the remaining data.
-Therefore, this boxplot shows differences between locations as was expected following the tendencies of the temperature  (fig.
-\ref{fig:temp},c).
-Apartado and Arauca had the highest temperatures hence, the highest thermal time values with 2909 and 2764 days$^\circ$C  respectively.
+Therefore, this boxplot shows differences between locations as was expected following the tendencies of the temperature of the figure \ref{fig:temp}, (b), where Aparatado, Arauca and Cali had greater temperatures than Santander and Caldas.
+
+Therefore, Apartado and Arauca had the highest temperatures and consequently the highest thermal time values with 2909 and 2764 days$^\circ$C  respectively.
 Caldas had the lowest values with 1173 days$^\circ$C.
@@
 	\centering
-	\includegraphics[scale=0.3]{images/ttbbox.png}\\
-	\caption{\footnotesize {Cocoa yield and thermal time characterisation at 180 days after flowering.\\}}
+	\includegraphics[scale=0.2]{images/ttbbox.png}\\
+	\caption{\footnotesize {Cocoa yield and thermal time characterisation at 180 DAF.\\}}
 	\label{fig:ttbox}
@@
 \subsection{Model Validation}
-{\color{olive}Biomass production of the aerial part of the cocoa plant was simulated which include every organ of the plant that is over the soil surface.
-It is important to calculated how much biomass  form the aerial part belongs to the the pods according to the harvest index (HI) (table \ref{tab:Treaparam}), hence cocoa seeds.
-The figure \ref{fig:m1} (a) shows the daily biomass growth rate.} Biomass simulation is affected by the solar radiation, RUE, daily temperature, atmospheric CO$_{2}$ concentration (ppm) and the fraction of solar radiation intercepted by a tree of cocoa during the fruit development (fSolar) (fig.
-\ref{fig:m1}, b).{\color{olive} There were not field data of absorption of solar radiation by the plants or biomass production then there is not observed data to compare the simulated data.}   {\color{blue}WHAT KIND OF DATA??
-I THOUGHT YOU DOWNLOADED DATA FROM NASA.
-AR: Just weather data is from NASA  } {\color{blue} WHAT DO YOU MEAN BIOMASS RESPONDED?
-AR: I meant that it is proportional to the yield.
-I DIDN'T UNDERSTAND HERE, DROP THE THE FROM BIOMASS BECAUSE WE ARE NOT TALKING ABOUT A SPECIFIC BIOMASS AR: it is biomass of the aerial part (every organ of the plant that is over the soil surface)}.
-{\color{olive} However, the daily biomass growth rate increased proportionally with the yield production (Fig.
-\ref{fig:m1}, c) .
-Thus, it was possible to calculated the final yield as the product of accumulated biomass at the harvest day and HI = 0.3 (Equ.\ref{equ:HI}).}
+
+Biomass production of the aerial part of the cacao tree was simulated which include every organ of the plant that is over the soil surface.
+It is important to calculate how much biomass from the crop aerial is partitioned to the cacao pods according to the harvest index (HI) (table \ref{tab:Treaparam}) as this quantity will correspond to the weight of cacao seeds.
+
+Figure \ref{fig:m1} (a) shows the daily biomass growth rate for five agroclimatic regions.
+Biomass simulation is affected by the solar radiation, RUE \citep{Fletcher2013RUE,Bonhomme2000}, daily temperature, atmospheric CO$_{2}$ concentration (ppm) and the fraction of solar radiation intercepted by a tree of cocoa during the fruit development (fSolar) (fig.
+\ref{fig:m1}, b).
+We did not compare predicted and observed Biomass as there were not field data corresponding to absorption of solar radiation by the plants or of biomass production.
+Figure \ref{fig:m1} (c) also shows that the daily biomass growth rate increased proportionally with the yield production, suggesting that it was possible to calculate the final yield as the product of accumulated biomass at the harvest day and HI = 0.3 (Equ.\ref{equ:HI}).
 
 \begin{equation}
-Cocoa~yield = Accumulated biomass~X ~ HI \\
+Cocoa~yield = Accumulated~ biomass~\times~ HI \\
 \label{equ:HI}
 \end{equation}
-(HI) (Eq. (5))
 
-{\color{olive} The fraction of solar radiation intercepted cocoa trees during the fruit development (fSolar) was also simulated during the fruit development cycle.
-The results in the figure \ref{fig:m1} (b) showed that} all regions had the maximum fSolar at 0.94, except Caldas where crops reached 0.76.
-Apartado and Arauca reached faster this fSolar-max.
-These high values of solar radiation intercepted for photosynthesis, lasted differently depending on the region and their solar radiation (Fig.\ref{fig:temp}, a) : Apartado 66 days  from  69 to 135 DAF,  Arauca 61 days from 72 to 133 DAF, Cali 38 days from 83 to 121 DAF,  Santander 24 days from 92 to 116 DAF and Caldas 2 days at 125 DAF.
-fSolar declined in the interception of solar radiation until the pod harvest day.
+The fraction of solar radiation intercepted by cacao trees (fSolar) was also simulated for the fruit development cycle.
+The relation of fsolar and is strongly related with RUE \citep{Fletcher2013RUE,Bonhomme2000}, LAI and hence the senescence of the canopy leaves \citep{lahive2019, Danner2015, Soltani2012, Romero2017, Vina2011, Zao2019simple}.
+Results in the figure \ref{fig:m1} (b) showed that all regions had a maximum fSolar of 0.94, except from Caldas whose peak fSolar was at 0.76.
+fSolar-max was faster reached in the regions of Apartado and Arauca.
+These fSolar-max values of solar radiation intercepted for photosynthesis, lasted differently depending on the region and their solar radiation (Fig.\ref{fig:temp}, a) : Apartado 66 days  from  69 to 135 DAF,  Arauca 61 days from 72 to 133 DAF, Cali 38 days from 83 to 121 DAF,  Santander 24 days from 92 to 116 DAF and Caldas 2 days at 125 DAF.
+fSolar declined and hence the interception of solar radiation until the pod harvest day.
 
-Cocoa yield simulation (kg ha$^{-1}$ per year) was validated using observed data (table \ref{tab:reparam}) showing a final RRMSE of 7.2\% (fig.
+The RRSME has been used before to evaluated other crop model simulations \citep{Zao2019simple, Bai2020}.
+Our results of cocoa yield simulation (kg ha$^{-1}$ per year) and validation (table \ref{tab:reparam}) achieved a final RRMSE of 7.2\%  fit between simulated and observed data (fig.
 \ref{fig:m1}, c).
@@
 Therefore, the highest  yield values simulated were obtained for Arauca over 4000 kg ha$^{-1}$, followed by Apartado Santander with yields over 2000 kg ha$^{-1}$.
-The lowest yield was simulated for Caldas region with less of 1000 kg ha$^{-1}$ .
-Final yield in the model was is calculated as the product of biomass of aerial part and harvest index (HI) \citep{Zao2019simple, Amir1991}, where the HI is similar to the CropSyst \citep{STOCKLE2003} and AquaCrop \cite{Steduto2009}.
+The lowest yield was simulated for the Caldas region with less of 1000 kg ha$^{-1}$ .
+Final yield in the model was calculated as the product of biomass of aerial part and harvest index (HI) \citep{Zao2019simple, Amir1991}, where the HI is similar to the CropSyst \citep{STOCKLE2003} and AquaCrop \cite{Steduto2009}.
 
@@
 
-\subsection{Predicting Pod Harvest Day }
+\subsection{Predicting Optimal Pod Harvest Day }
 
-{\color{olive} The cocoa model predict the day when pods can be harvested (predicted) and the results were compared with the 180 counting after flowering (observed) because it is the empirical date of harvest.
-The figure \ref{fig:dayH} present that observe day of harvest (180 DAF) independently of the region.
-All the regions except Santander, presented the earliest predicted harvest when the flowering date was between December and February.
+Farmers practice of empirically harvesting is about counting  180 DAF (6 months) by calendar without taking into account weather changes or physiological features of cacao cultivars.
+Since our cacao model was fitted to predict optimal pods harvest date, we compared model predicted dates against the usual 180 days that growers and experts count after the FD.
+Figure \ref{fig:dayH} shows observed day of harvest (180 DAF) independently of the region.
+All the regions except Santander, presented the earliest predicted harvest when the FD was between December and February.
 The fruit can be ready to harvest before or after 180 DAF depending on the environmental condition per region.
-The results demonstrated that the traditional way to harvest which is always at 180 DAF, it is not having account physiological and environmental conditions that can be affecting the pod maturation.} Thus, when the fruit development was simulated the maturity day in Cali and Caldas were from 3 to 12 and 4 to 23 days before 180 DAF (Fig.
-\ref{fig:dayH}, c and f respectively).
-Apartado presented the most similar predicted dates of harvest to 180 DAF with 170 to 182 DAF  (Fig.
-\ref{fig:dayH}, d).
-The pod may be harvest in Arauca (Fig.\ref{fig:dayH}, b)  between 165 and 193 DAF, Santander (Fig.\ref{fig:dayH}, a) between 165 and 183 DAF, ten days less than in Arauca for the same months of flowering of July and August of 2019 .
-Only Arauca presented longer crop cycles when the flowering was between during that period of time.
-This means that Arauca had bigger variation of temperature between months (Fig.
-\ref{fig:temp}, b ).
-{\color{olive} To summarize, depending on which month of the year the trees are flowering,  the number of days to reach the harvest of ripe pods vary more or less 180 days (table \ref{tab:harvest}).}
+The results in the figure \ref{fig:dayH} demonstrated that the traditional way to harvest which is always at 180 DAF, it is not having account physiological and environmental conditions that can be affecting the pod maturation.
+Thus, when the fruit development was simulated the maturity day in Cali and Caldas were from 3 to 12 and 4 to 23 days before 180 DAF.
+Apartado presented the most similar predicted dates of harvest to 180 DAF with 170 to 182 DAF.
+The pod may be harvest in Arauca from 165 to 193 DAF, Santander from 165 to 183 DAF, ten days less than in Arauca for the same months of flowering of July and August of 2019 .
+Only Arauca presented longer crop cycles when the FD was between during that period of time.
+This means that Arauca had bigger variation of temperature between months.
+To summarize, depending on which month of the year cacao trees are flowering,  the number of days to reach the harvest of ripe pods vary more or less 180 days.
+For example, if cacao trees in Caldas are flowering in January, pods will be ripe to harvest in about 158 days proximately  (table \ref{tab:harvest}).
 
@@
 	\includegraphics[scale=0.4]{images/RegionHarvest2.png}
-	\caption{\footnotesize {Harvest day prediction from flowering date for (a) Santander, (b) Arauca, (c) Cali, (d) Apartado,(f) Caldas. \\ }} 
+	\caption{\footnotesize {Cacao harvest day prediction from FD for Apartado , Arauca, Caldas, Cali  and  Santander. DAF = Days After Flowering. FD = Flowering Date. \\ }} 
 	\label{fig:dayH}
@@
 
+Cacao is one of the most economically important crops in Colombia.
+In 2016 Colombia produced a historical of 56K tons, up from 35K ten years ago, and over the same period, imports decreased from 12.8K to 488 tons (FAOStat).
+Cacao has been fundamental in the transformation of the rural economy dependent on illicit crops.
+As such, Cacao is called the "peace crop", nominated by  Colombia's President and 2016 nobel peace prize winner, Juan Manuel Santos, and by the President of the Association of Cacao Growers (Fedecacao).
+Although Colombian cacao has the potential to be in the high value markets for fine flavour, the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly.
+This study reports the cacao model which was calibrated for five agroclimatic regions in Colombia.
+The purpose of this model was to help farmers achive  higher quality cacao beans by providing them with a tool that not only predicts optimal harvest date but also estimated yield and biomass.
+
 \subsection{Weather Effects over Flower Stability and Pollination}
-This study presents a new approach of SIMPLE model calibration for cocoa as tropical crop in South America for five environments.
-The analysis of the weather conditions (Fig.
-\ref{fig:heat}) did not confirmed clearly that flower are affected by the wind and rain by mechanical damage.
+
+Our analysis of weather variables (Fig.
+\ref{fig:heat}) could not show a correlation between flower shedding and high wind or rain because information on flowers was not part of the field data collected.
 However, farmers stated that the number of flowers pollinated decrease by months where wind and rain are high.
-Moreover, wind can affect the availability of tiny flies pollinators from Diptera order and from the families of of the biting midges \textit{Ceratopogonidae},  genus  \textit{Forcipomyia} \citep{Saunders1959, kaufmann1975, sotomayor2020} to reach the cocoa flowers.
-However, the stability of cocoa flowers is influenced by seasonal wheather conditions (abiotic) and pollination (biotic) \citep{Frimpong2014}.
-Therefore, pollinator population shoud be coincidence with the phenology of the flowering cocoa trees \citep{Young1983, Young2012}.
+High winds can affect the availability of tiny flies pollinators from Diptera order and from the families of of the biting midges \textit{Ceratopogonidae},  genus  \textit{Forcipomyia} \citep{Saunders1959, kaufmann1975, sotomayor2020} to reach the cacao flowers.
+However, the stability of cacao flowers is influenced by seasonal wheather conditions (abiotic) and pollination (biotic) \citep{Frimpong2014}.
+Therefore, pollinator population should be coincidence with the phenology of the flowering cacao trees \citep{Young1983, Young2012}.
 Flower opening is very well synchronised between the cohorts of mature flowers opening each night \citep{Niemenak2010}.
@@
 
-In our preliminary analysis, Santanter data, showed that the number of successful flowers pollinated to produce final yield can be affected negatively mostly by the rain, TMAX and wind.
+When analysing the data from the Santanter region, we showed that the number of successful flowers pollinated to produce final yield could be affected by rain, TMAX and wind.
 Nevertheless, a better field data tracing flowers development is essential to understand if it is a mechanical or physiological effect.
-In contrast,  \cite{Frimpong2009, Frimpong2011} indicated that the numbers of cocoa pollinators were reduced during the dry season, but increased in the wet season.
+\cite{Frimpong2009, Frimpong2011} indicated that the numbers of cocoa pollinators were reduced during the dry season, but increased in the wet season.
 This, could be due to midges need a moist environment to develop \citep{Frimpong2014}, which is difficult during the dry season as cocoa leaves create a dried ground mat \citep{Frimpong2009, Frimpong2014}.
 Moreover, the lack of water during dry seasons may reduce the nutrients uptake provoking the massive flower drops \citep{Vaughton2017}.
-For future studies, wind could be included in the SIMPLEcocoa model as an input to simulated this mechanical effects over number of flowers.
+For future studies, wind could be included in the cocoa model as an input to simulate this mechanical effects over number of flowers.
 In general, field data regarding counting flowers pollinated by month should be better reported for the region here studied.
 
-\subsection{Thermal Time for Harvest day predictions}
+\subsection{Thermal Time for Harvest Day Predictions}
 We define the thermal time required to harvest cocoa pods for five Colombian regions as maturation of the fruit is related with temperature during the growth cycle \citep{lopez2018}.
@@
 In contrast, the SIMPLE model has an open code in R, which we could adapt such a model would be very useful to compare yields and predict harvest date in different climates.
-As the harvest day was predicted form the flowering date (Fig.
+As the harvest day was predicted from the FD (Fig.
 \ref{fig:dayH}, consequently, biomass production and fSolar presented a crop cycle shorter than 180 DAF (Fig.
@@
 
-Biomass simulations use SIMPLEcocoa model presented similar predicted values (10000 Kg ha$^{-1}$) for coca drops in Costa Rica using SUCROScocoa \citep{zuidema2005}.
+Biomass simulations use cocoa model presented similar predicted values (10000 Kg ha$^{-1}$) for coca drops in Costa Rica using SUCROScocoa \citep{zuidema2005}.
 Biomass simulations are a common evaluation in crop modells such as Sirius \citep{Crout20142}, SUBSTOR-potato \citep{Raymundo2017} and  DSSAT, CropSyst, STICS and WOFOST \citep{Confalonieri2016}.
@@
 \ref{fig:m1}, b) when the senescence of the canopy \citep{zuidema2005}.
-In cocoa canopy senecence  refereed to  a group of leaves responsible  at the moment of the fruit formation to produce carbohydrates.
+The relation of fsolar and LAI \citep{lahive2019, Danner2015, Soltani2012, Romero2017, Vina2011, Baracaldo2014} and RUE \citep{Fletcher2013RUE,Bonhomme2000} have widely been reviewed in literature .
+The  production of photo-assimilates, dry matter and yield can be affected by LAI and RUE reduction \citep{Deblonde2001, Fletcher2013RUE,Bonhomme2000, Romero2017}.
+Therefore, LAI values are utilized to predict primary photosynthetic production and crop growth \cite{Romero2017, Soltani2012, zuidema2005,Baracaldo2014}.
+In cocoa canopy senescence  refereed to  a group of leaves responsible  at the moment of the fruit formation to produce carbohydrates.
 These leaves eventually drop becoming on litter over the soil.
-Leaves life cycle has been simulated using crop models \citep{Crout2010, zuidema2005}, which can be the reference to improve our SIMPLEcocoa crop model in future studies.
+Leaves life cycle has been simulated using crop models \citep{Crout2010, zuidema2005}, which can be the reference to improve our cacao crop model in future studies.
 
@@
 
-\subsection{App development and future challenges }
-The original code of SIMPLE model of \citep{Zao2019simple} was modified to make easy the implementation of this cocoa model as and app to be used in smartphones and desktops by farmers in Colombia.
+\subsection{App Development}
+The cocoa model code was adapted to make easy the implementation of this cocoa model as and app to be used in smartphones and desktops by farmers in Colombia.
 Therefore, the new version for cocoa crops simulation will be used to predict yield, date of harvest and biomass production, inserting only the date of flowering and region.
@@
 \section{Conclusions}
-This research presented and initial crop calibration that can be improved with further studies, including effects over the pod production by diseases, nutritional diffidences and abiotic stresses.
+
+Cocoa fruit development for harvest in the right time depend on whether conditions and principles of crop physiology and flower phenology.
+This was common for the five regions.
+Thermal time characterisation range from 1200 to 3000 days $^\circ$C, with a T$_{b}$ of 10 $^\circ$C for the fruit development.
+The cocoa model allowed to predict the harvest date with better precision than only considering days by calendar.
+Thus, the crop cycle of cocoa for harvest  should be shorter than 180 days after flowering.
+These results confirm the potential of the Crop Simulation Model approaches  for tropical crop in Latin America.
+
+This research presented and initial crop calibration that can be improved with further studies, to include effects over the pod production by diseases, nutritional deficiencies and abiotic stresses.
 The future challenge will be that traditional farmers start to harvest more aware of the environmental effects over their crops.
@@
 
-Cocoa fruit development for harvest in the right time depend on whether conditions and principles of crop physiology and flower phenology.
-This was common for the five regions.
-Thermal time characterisation range from 1200 to 3000 days $^\circ$C, with a T$_{b}$ of 10 $^\circ$C for the fruit development.
-The SIMPLEcocoa model allowed to predict the harvest date with better precision than only considering days by calendar.
-Thus, the crop cycle of cocoa for harvest  should be shorter than 180 days after flowering.
-These results confirm the potential of the Crop Simulation Model approaches  for tropical crop in Latin America.
+\section{Availability of Source Code}
+Software will be provided under user request.
+The data used in this study can be found as follows: kocolatl available in the  research data repository.\\ https://github.com/anyelacamargo/kocolatl.git .
 
-\section{Author contributions}
-ARV and ARC model calibration and data analysis. PA and ODR field data collection. AMG, ARC, and ARV editing and results interpretations. ARV were primarily responsible for writing the manuscript.
+\section{Author Contributions}
+ARV and AVCR model calibration and data analysis.
+PA and ODR field data collection.
+AMG, ARC, and ARV editing and results interpretations.
+ARV and AVCR were primarily responsible for writing the manuscript.
 
-\acknowledgments{The authors wish to thank to Grupo BIOS and Fedecacao for facilitated the data from cocoa fields. }
+\acknowledgments{The authors wish to thank to Grupo BIOS and Fedecacao for facilitated the data from cocoa fields. Thanks to Grupo BIOS for manage and founding this research.}
 

### 2021-12-15 paperModel/paperCacaoModel_round1.tex
round1 corrections

--- 2021-11-23 paperModel/paperModel_2.tex
@@
 \usepackage[utf8]{inputenc}
+\usepackage[spanish,english]{babel}
 %=================================================================
@@
 Although Colombian cacao has the potential to be in the high value markets for fine flavour, the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly.
-For example, traditionally cacao harvest takes place at 5 or 6 months after flowering, other environmental parameters that have more association with pod maturation speed are not taken into account.
+\textcolor{blue}{Traditionally cacao harvest takes place at 5 or 6 months  by calendar after flowering, the problem is that  the seeds are often  unripe or  over ripe germinating inside the pod afecting the quality of the final cocoa products}.
+Other environmental parameters that have more association with pod maturation speed are not taken into account.
 Cacao fruits development can be considered as the result of a number of physiological and morphological processes that can be described by mathematical relationships even under uncontrolled environments.
@@
 Since, harvesting at the right time can yield high quality cacao, we parametrised a crop model to predict the best time for harvest cocoa fruits in Colombia.
-Our aim was to develop a practical tool that supports cacao farmers in the production of high quality cacao.
+.
+Results can be a practical tool that supports cacao farmers in the production of high quality cocoa.
 When comparing simulated and observed data, our results showed an RRMSE of 7.2\% for the yield prediction, while the simulated harvest date varied between +/- 2 to 20 days depending on the temperature variations of the year between regions.
-This crop model contributed to understand and predict the phenology of cacao fruits for two key cultivars ICS95 y CCN51.}
+This crop model contributed to understand and predict the phenology of cacao fruits for two key cultivars ICS95 y CCN51.
+\textcolor{blue}{The aim of this study was to calibrate a crop model that simulates cacao crop development and yield, and predicts the maturation day when the fruit is likely to be ready for harvest.}}
 
@@
 Cacao plant is a member of the Malvaceae (formerly Sterculiaceae)  botanical family such as  cotton \textit{ Gossypium hirstium} \citep{Nix2017cotton}.
-Cotton has been is modeled in SIMPLE model \citep{Zao2019simple}.
+Cotton has been is modelled in SIMPLE model \citep{Zao2019simple}.
 Cacao is grown for its fruits, known as cacao pods \citep{ Niemenak2010, suarez2021}.
@@
 Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}, the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly.
-For example, the farmer practice of empirically harvesting from 5 to 6 months or 180 Days After Flowering  (DAF) date, produced a mix of quality of cacao beans that are then fermented.
+The farmer practice of empirically harvesting from 5 to 6 months or 180 Days After Flowering  (DAF) date, produced a mix of quality of cacao beans that are then fermented.
+\textcolor{blue}{The problem is that  the seeds are often  unripe or  over ripe germinating inside the pod affecting the cocoa quality}.
 This practice produces heterogeneous characteristics between each fermentation batch which potentially diminish the quality of the final cocoa product \citep{Escobar2021}.
-To identify the best moment to harvest is important to consider physiological responses affected by climate variables such as rain, solar radiation and wind.
-Therefore,  for cacao production in Colombia, physiological simulation models may be valuable to identify the best moment to harvest cocoa fruits considering variable weather conditions, soil types and cultivar specifications.
+\textcolor{blue}{Therefore,  It is important to identify the best moment to harvest  considering physiological responses affected by  soil types , cultivar specifications and weather variables such as rain, solar radiation and wind.
+Thus,  for cacao production in Colombia, physiological simulation models may be valuable to predict the right moment to harvest cocoa fruits.}
 
@@
 Recently crop modelling studies are focusing on  perennial crops  production \citep{zuidema2005, Zao2019simple, Bai2020, Romero2021}.
-However, the information reported  is be less than for annual crops due to the lack of field data available , relatively high research costs and the difficulties of accumulated errors in long-term simulations \citep{zuidema2005}.
+However, the information reported  is less than for annual crops due to the lack of field data available , relatively high research costs and the difficulties of accumulated errors in long-term simulations \citep{zuidema2005}.
 For cacao there are  approaches  to predict yield mainly  using algorithms of machine learning \citep{lamos2020} and just one mechanistic model simulates physiological cocoa performance  "SUCROS-cocoa" \citep{zuidema2005}.
@@
 We parametrised this crop model  because it had already been successfully fitted to other tree crops in south America.
-The model simulates crop development, growth and yield, and predicts the maturation day when the fruit is likely to be ready for harvest.
-Our aim was to develop a practical tool that supports cacao farmers in the production of high quality cocoa.
+Our aim was to calibrate a crop model that simulates crop development, growth and yield, and predicts the maturation day when the fruit is likely to be ready for harvest.
+Results can be a practical tool that supports cacao farmers in the production of high quality cocoa.
 Thus, making Colombian cacao more competitive in the Fine-cocoa market.
 
-\section{Materials and Methods}
-
-\subsection{Floral Phenology of Cocoa}
+\subsection{Floral Phenology of Cacao}
 
@@
 
+\section{Materials and Methods}
+
 \subsection{Test Site and Yield Production }
@@
 Thus, 23 flowering dates for five farms gave us 115 samples in total.
+
+\subsection{Weather Conditions }
+We analysed and compared weather patterns linked our five geographical regions (Fig.\ref{fig:yield}, a) and across 2018 to 2020 .
+Looking at the data, the regions of Santander and Caldas had the biggest variability and the maximum of solar radiation values over 20 MJ m$^{2}$day$^{-1}$ .
+In contrast , the regions of Cali and Arauca presented the lowest values of PAR below 5 MJ m$^{2}$day$^{-1}$.
+(Fig.\ref{fig:temp}, a).
+Even though, Cali and Santander had contrasting PAR conditions, these regions presented  have similar temperature during 2020 (Fig.\ref{fig:temp}, b).
+The temperature ranges from 16 to 28 $^\circ$C and it is relatively constant for each region.
+However, Arauca presented the biggest variability with hotter months during the first half of the year 2019 and 2020.
+Apartado was found as the hottest region studied with a relative constant temperature of 26 $^\circ$C and Caldas as the coldest site with 16 $^\circ$C.
+Precipitation in Colombia is presented in two seasons per year from February to April and from October to November, while the relative humidity remain constant over 80\% (Fig.\ref{fig:temp},c).
+In general, the coldest regions tested (Caldas and Santander) had the maximum values of solar radiation available for photosynthesis.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[scale=0.4]{images/clima.png}
+	\caption{\footnotesize {Colombian weather conditions. \\}}	
+	\label{fig:temp}
+	{\footnotesize (a), Available photosynthetic solar radiation (PAR).
+(b), Daily average temperature  (c), Monthly average of precipitation (bars) and relative humidity (dotted line) from 2018 to 2021 }
+\end{figure}
+\newpage
 
@@
 
-The cocoa model performance was evaluated by comparing simulated values cocoa yield with those reported by Fedecacao from cocoa plantations, using the statistical indice of Relative Root Mean Square Error (RRMSE)  in the equation \ref{equ:RRMSE}, where n is the total number of observations, Y$_{i}$ corresponds to observed value from field and X$_{i}$ is the predict value from the model.\\
+The cocoa model performance was evaluated by comparing simulated values cocoa yield with those reported by Fedecacao from cocoa plantations, using the statistical index of Relative Root Mean Square Error (RRMSE)  in the equation \ref{equ:RRMSE}, where n is the total number of observations, Y$_{i}$ corresponds to observed value from field and X$_{i}$ is the predict value from the model.\\
 
@@
 \subsection{Weather Conditions Over Flowering Time }
-We analysed and compared weather patterns linked our five geographical regions (Fig.\ref{fig:yield}, a) and across 2018 to 2020 .
-Looking at the data, the regions of Santander and Caldas had the biggest variability and the maximum of solar radiation values over 20 MJ m$^{2}$day$^{-1}$ .
-In contrast , the regions of Cali and Arauca presented the lowest values of PAR below 5 MJ m$^{2}$day$^{-1}$.
-(Fig.\ref{fig:temp}, a).
-Even though, Cali and Santander had contrasting PAR conditions, these regions presented  have similar temperature during 2020 (Fig.\ref{fig:temp}, b).
-The temperature ranges from 16 to 28 $^\circ$C and it is relatively constant for each region.
-However, Arauca presented the biggest variability with hotter months during the first half of the year 2019 and 2020.
-Apartado was found as the hottest region studied with a relative constant temperature of 26 $^\circ$C and Caldas as the coldest site with 16 $^\circ$C.
-Precipitation in Colombia is presented in two seasons per year from February to April and from October to November, while the relative humidity remain constant over 80\% (Fig.\ref{fig:temp},c).
-In general, the coldest regions tested (Caldas and Santander) had the maximum values of solar radiation available for photosynthesis.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[scale=0.4]{images/clima.png}
-	\caption{\footnotesize {Colombian weather conditions. \\}}	
-	\label{fig:temp}
-	{\footnotesize (a), Available photosynthetic solar radiation (PAR).
-(b), Daily average temperature  (c), Monthly average of precipitation (bars) and relative humidity (dotted line) from 2018 to 2021 }
-\end{figure}
-\newpage
-
-Figure \ref{fig:heat} shows the pearson correlation to study the weather data of the flowering time over the months of flowering  (monthF), month of harvest (monthH) and their final yield (fruit\_kg).
+Figure \ref{fig:heat} shows the Pearson correlation to study the weather data of the flowering time over the months of flowering  (monthF), month of harvest (monthH) and their final yield (fruit\_kg).
 The results showed that thermal time T$_{b}$ (ttb) is correlated (P = 0.52) with daily average temperature and maximum temperature (TMAX) and  temperature minimum (TMIN) and Dew Frost Point at 2 meters (T2MDEW) with  a correlation coefficient of 0.60.
@@
 Thermal time characterisation was made considering 180 DAF for each location.
-The boxplot in the figure \ref{fig:ttbox} shows the data distribution where boxes indicate the range of the central 50\% of the total data per region, the central line in the box is marking the median value and lines draw out from each box mean the range of the remaining data.
+The box-plot in the figure \ref{fig:ttbox} shows the data distribution where boxes indicate the range of the central 50\% of the total data per region, the central line in the box is marking the median value and lines draw out from each box mean the range of the remaining data.
 Therefore, this boxplot shows differences between locations as was expected following the tendencies of the temperature of the figure \ref{fig:temp}, (b), where Aparatado, Arauca and Cali had greater temperatures than Santander and Caldas.
@@
 \ref{fig:m1}, c).
+\textcolor{blue}{ The low RRMSE  indicates the simplicity of the Cacao model to simulated cacao  yield.
+Although, the original SIMPLE model can be used to asses heat and water stress, our cacao model calibration  did  not accounted biotic (pest and diseases) and abiotic (heat or water stress) factors that could have significant effects on the cacao production.}
+
 Individual errors per region are presented in table \ref{tab:error}, where the best fit of the calibration model for yield prediction  was for crops in  Apartado and the highest error was calculated for Caldas crops.
@@
 Cacao has been fundamental in the transformation of the rural economy dependent on illicit crops.
-As such, Cacao is called the "peace crop", nominated by  Colombia's President and 2016 nobel peace prize winner, Juan Manuel Santos, and by the President of the Association of Cacao Growers (Fedecacao).
+As such, Cacao is called the "peace crop", nominated by  Colombia's President and 2016 Nobel peace prize winner, Juan Manuel Santos, and by the President of the Association of Cacao Growers (Fedecacao).
 Although Colombian cacao has the potential to be in the high value markets for fine flavour, the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly.
-This study reports the cacao model which was calibrated for five agroclimatic regions in Colombia.
-The purpose of this model was to help farmers achive  higher quality cacao beans by providing them with a tool that not only predicts optimal harvest date but also estimated yield and biomass.
+
+\textcolor{blue}{This study reports the cacao model which was calibrated for five agroclimatic regions in Colombia.
+The model simulates crop development, growth and yield, and predicts the maturation day when the fruit is likely to be ready for harvest.
+The purpose of this model was to help farmers achieve  higher quality cacao beans by providing them with a tool that not only predicts optimal harvest date but also estimated a potential yield and biomass .
+The low RRMSE for yield prediction indicates the simplicity of the Cacao model to simulated cacao  yield.
+Although, the original SIMPLE model can be used to asses heat and water stress, our cacao model calibration  did  not accounted biotic (pest and diseases) and abiotic (heat or water stress) factors that could have significant effects on the cacao production.
+Santander and Caldas have the  highest solar radiation but the lowest temperature.
+In contrast , Arauca and Apartado have the  highest temperature but the lowest solar radiation
+Weather analysis could not show a correlation between flower shedding and high wind or rain because information on flowers was not part of the field data collected.
+	The growth cycle was simulated in terms of thermal time which was defined for the five regions tested, given the dependency to the crop model to predict the harvest day according to the weather variations while fruit are growing.
+The thermal time calculation was defined base on  180 DAF because farmers cut the pods by calendar days.
+Our results showed how the harvest day can vary depending on the accumulate temperature during each specific crop cycle simulated (Fig.
+\ref{fig:dayH}).
+}
 
@@
 Abscission processed of  flowers are mainly controlled  by  three  hormones: auxin, ethylene, and abscisic acid (ABA) \citep{Aneja1999}.
-Ethylene generally promotes abscission because it may  inhibit the transport of auxin from the leaf blade, which allow the action of ABA to promove the fall of flowers \citep{Beyer1975}.
+Ethylene generally promotes abscission because it may  inhibit the transport of auxin from the leaf blade, which allow the action of ABA to promote the fall of flowers \citep{Beyer1975}.
 In general, environmental conditions can also  stimulate   to  a  decrease  the  auxin/ethylene relationship \citep{Aneja1999}.
 
-When analysing the data from the Santanter region, we showed that the number of successful flowers pollinated to produce final yield could be affected by rain, TMAX and wind.
+When analysing the data from  Santanter region, we showed that the number of successful flowers pollinated to produce final yield could be affected by rain, TMAX and wind.
 Nevertheless, a better field data tracing flowers development is essential to understand if it is a mechanical or physiological effect.
@@
 the effects of temperature and solar radiation on fruit growth and development was previously studied by \cite{Daymond2008}, showing that crop under higher temperatures thought the crop cycle induce greater fruit losses because of physiological maturation (cherelle wilt).
-When the fruit is mature, seeds are able to germinate inside the pod before the harvest day \citep{lopez2018}.
+When the fruit is mature, seeds are able to germinate  \citep{lopez2018}.
+\textcolor{blue}{However, when ripe fruits stay for a longer time on the tree without be harvested  on the right moment, seeds can germinate inside the pod damaging  cacao production for high-quality taste and affecting the final yield.
+}
 
@@
 These simulations are coincident with results presented by \cite{lopez2018}, where physiological maturity of coca pod varies from 140 to 162 DAF.
-Our results showed how the harvest day can vary depending on the accumulate temperature during each especific crop cycle simulated (Fig.
+Our results showed how the harvest day can vary depending on the accumulate temperature during each specific crop cycle simulated (Fig.
 \ref{fig:dayH}).
+\textcolor{blue} {Harvesting cocoa fruits at the right moment the quality of the seed inside the pod can improve avoiding the germination before collect the pot from  cacao trees.}\textcolor{olive} {An research made by Gallego (2021), cacao clones CCN51 and ICS95 were handed-pollinated and pods were collected at 6, 7, and 8 months of ripeness after pollination.
+Cacao fruits were harvested from the municipality of Belalcazar, Caldas, Colombia to analyse metabolites related with cocoa quality were more expressed when fruits are collected at 6, 7, and 8 months after flowering.
+Results showed that the metabolites that mostly explain the differences between the classes ( metabolites with the highest VIP) are highly overexpressed at month 6 and then decline or under-expressed at the next months.
+No significant differences were found in their metabolic profile Between month 7 and 8.
+Quality of cocoa is defined by  some metabolites that are associated with flavonoids or colouration.
+When the expression of flavonoids decrease , indicates  that  metabolites  associated with flavonoids  are being oxidized and metabolites are converted by cells into other less oxidizing metabolites.
+The colouration does not show a greater difference after month 6.
+If there is no good  accumulation of flavonoids which are also related with metabolites associated with colouration, the quality of cocoa seed decrease.
+Ripening is associated with changes in colouration and these in turn with flavonoids so month 6 was ideal for harvesting assuring the high quality of cocoa.
+} \textcolor{blue} { Therefore, the cacao model can predict with more detail the optimum day to harvest the pod ensuring the quality of the cacao bean  destined for Fine-cocoa products.
+}
 
@@
 Resulting in reliable approach for cocoa yield prediction.
-In general, these results may help to improve the quality of cocoa seed considering the moment to harvest can be variable depending on weather changes.
+\textcolor{blue}{The low RRMSE for yield prediction indicates the simplicity of the Cacao model to simulated cacao  yield.
+Although, the original SIMPLE model can be used to asses heat and water stress, our cacao model calibration  did  not accounted biotic (pest and diseases) and abiotic (heat or water stress) factors that could have significant effects on the cacao production.} In general, these results may help to improve the quality of cocoa seed considering the moment to harvest can be variable depending on weather changes.
 
@@
 Thus, the crop cycle of cocoa for harvest  should be shorter than 180 days after flowering.
-These results confirm the potential of the Crop Simulation Model approaches  for tropical crop in Latin America.
+
+\textcolor{blue}{Results are only valid for the regions tested and they can be used for farmers in Colombia for improve the cacao quality.
+However, the model calibration and values of parameters used may be applied by the international community to test cacao crops in other regions calculating the specific thermal time using the SIMPLE model.
+Moreover, other cacao varieties and regions to test can be in included.
+As for most of the crop modelling studies field data is extremely important .} These results confirm the potential of the Crop Simulation Model approaches  for tropical crops in Colombia and other regions in Latin America.
 

### 2021-12-15 paperModel/paperCacaoModel_round1.tex
round1 corrections

@@
 This crop model contributed to understand and predict the phenology of cacao fruits for two key cultivars ICS95 y CCN51.
-\textcolor{blue}{The aim of this study was to calibrate a crop model that simulates cacao crop development and yield, and predicts the maturation day when the fruit is likely to be ready for harvest.}}
+\textcolor{blue}{The aim of this study was to calibrate a crop model that simulates cacao crop development and yield, and predicts the maturation day when the fruit is likely to be ready for harvest which will improve the quality of the seeds.}}
 
@@
 \textcolor{blue}{Therefore,  It is important to identify the best moment to harvest  considering physiological responses affected by  soil types , cultivar specifications and weather variables such as rain, solar radiation and wind.
-Thus,  for cacao production in Colombia, physiological simulation models may be valuable to predict the right moment to harvest cocoa fruits.}
+Thus,  for cacao production in Colombia, physiological simulation models may be valuable to predict the right moment to harvest cocoa fruits, hence it will improve the quality of seeds.}
 

### 2021-12-16 paperModel/paperCacaoModel_round1.tex
correction round 1

@@
 \textcolor{blue}{Therefore,  It is important to identify the best moment to harvest  considering physiological responses affected by  soil types , cultivar specifications and weather variables such as rain, solar radiation and wind.
-Thus,  for cacao production in Colombia, physiological simulation models may be valuable to predict the right moment to harvest cocoa fruits, hence it will improve the quality of seeds.}
+Thus,  for cacao production in Colombia, physiological simulation models may be valuable to predict the right moment to harvest cocoa fruits, hence it will improve the quality of seeds.
+}
+
+\textcolor{blue}{Climate change could severely impact cacao production as cacao crop is often described as being sensitive to climate change  producing negatively impact its growth, productivity and yield quality  \citep{zuidema2005, Ceccarelli2021}.
+However, there are some   	 adaptive responses to mitigate negative climate  effects, for example elevated CO$_{2}$ and tolerant varieties \citep{lahive2019, Ceccarelli2021}.
+The role of genetic diversity in producing more resilient cacao  to drought and temperature tolerance, or increased CO2 response,, are presented in \citep{Medina2017} .Thus crop models that include weather data can provide a tool to study the impact of climate change over the quality of cacao.
+}
 
@@
 \subsection{Cocoa Crop Model Simulations}
-Although cocoa is a relevant crop and there is an extensive agronomic literature, there is only one physiological crop model specific for cocoa so far.
-The (SUCROS-Cocoa) developed by \cite{zuidema2005}.
+Although cocoa is a relevant crop and there is an extensive agronomic literature, there is only one physiological crop model specific for cocoa so far is The (SUCROS-Cocoa) developed by \cite{zuidema2005}.
 However, the code was not easy available for adaptations.
@@
 \ref{fig:dayH}).
-\textcolor{blue} {Harvesting cocoa fruits at the right moment the quality of the seed inside the pod can improve avoiding the germination before collect the pot from  cacao trees.}\textcolor{olive} {An research made by Gallego (2021), cacao clones CCN51 and ICS95 were handed-pollinated and pods were collected at 6, 7, and 8 months of ripeness after pollination.
-Cacao fruits were harvested from the municipality of Belalcazar, Caldas, Colombia to analyse metabolites related with cocoa quality were more expressed when fruits are collected at 6, 7, and 8 months after flowering.
-Results showed that the metabolites that mostly explain the differences between the classes ( metabolites with the highest VIP) are highly overexpressed at month 6 and then decline or under-expressed at the next months.
-No significant differences were found in their metabolic profile Between month 7 and 8.
-Quality of cocoa is defined by  some metabolites that are associated with flavonoids or colouration.
-When the expression of flavonoids decrease , indicates  that  metabolites  associated with flavonoids  are being oxidized and metabolites are converted by cells into other less oxidizing metabolites.
-The colouration does not show a greater difference after month 6.
-If there is no good  accumulation of flavonoids which are also related with metabolites associated with colouration, the quality of cocoa seed decrease.
-Ripening is associated with changes in colouration and these in turn with flavonoids so month 6 was ideal for harvesting assuring the high quality of cocoa.
-} \textcolor{blue} { Therefore, the cacao model can predict with more detail the optimum day to harvest the pod ensuring the quality of the cacao bean  destined for Fine-cocoa products.
-}
+\textcolor{blue} {Harvesting cocoa fruits at the right moment the quality of the seed inside the pod can improve avoiding the germination before collect the pod from  cacao trees.
+Therefore, the cacao model can predict with more detail the optimum day to harvest the pod ensuring the quality of the cacao bean  destined for Fine-cocoa products.}
+\textcolor{olive} {Fruit ripening is a highly coordinated developmental process that coincides with seed maturation \citep{Osorio2013} .
+As cacao is a non climacteric fruit, the pod needs be connected to the mother plant until maturation.
+Therefore, an indirect way to infer the seed quality is checking the pod quality.
+Pod quality can be determined as the physico-chemical changes occurring at the peel level, (including colouration changes, and chemical dynamical changes in metabolites production such as sugars, phenolics, fatty acids) which give information about the seed status.
+For instance, clear evidence of above occurs when cacao ripe fruits stay for a longer time on the tree without be harvested on the right moment, causing that seeds germinate  inside the pod and damaging the cacao production for high-quality taste and affecting the final yield.
+An ongoing project conducted by BIOS is showing us the expression patterns for metabolites associated with cacao ripening.
+In this case, we found sugars, phenolic, fatty acids and flavonoids which according to the ripening stage change their expression.
+Furthermore, we are identifying biochemical markers of fruit ripening.
+We are also connecting these chemical changes with seed quality.
+It let us understand how the ripen occur in cacao and suggest the best moment to harvest.}
 
